### 2024-03-15 src/tex/ms.tex
[showyourwork] first commit

% Define document class
\documentclass[twocolumn]{aastex631}
\usepackage{showyourwork}

% Begin!
\begin{document}

% Title
\title{An open source scientific article}

% Author list
\author{@thomasckng}

% Abstract with filler text
\begin{abstract}
    Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
    Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae, felis.
    Curabitur dictum gravida mauris, consectetuer id, vulputate a, magna.
    Donec vehicula augue eu neque, morbi tristique senectus et netus et.
    Mauris ut leo, cras viverra metus rhoncus sem, nulla et lectus vestibulum.
    Phasellus eu tellus sit amet tortor gravida placerat.
    Integer sapien est, iaculis in, pretium quis, viverra ac, nunc.
    Praesent eget sem vel leo ultrices bibendum.
    Aenean faucibus, morbi dolor nulla, malesuada eu, pulvinar at, mollis ac.
    Curabitur auctor semper nulla donec varius orci eget risus.
    Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
    Duis eget orci sit amet orci dignissim rutrum.
\end{abstract}

% Main body with filler text
\section{Introduction}
\label{sec:intro}

Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae, felis.
Curabitur dictum gravida mauris, consectetuer id, vulputate a, magna.
Donec vehicula augue eu neque, morbi tristique senectus et netus et.
Mauris ut leo, cras viverra metus rhoncus sem, nulla et lectus vestibulum.
Phasellus eu tellus sit amet tortor gravida placerat.
Integer sapien est, iaculis in, pretium quis, viverra ac, nunc.
Praesent eget sem vel leo ultrices bibendum.
Aenean faucibus, morbi dolor nulla, malesuada eu, pulvinar at, mollis ac.
Curabitur auctor semper nulla donec varius orci eget risus.
Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
Duis eget orci sit amet orci dignissim rutrum.

Nam dui ligula, fringilla a, euismod sodales, sollici- tudin vel, wisi.
Morbi auctor lorem non justo, nam lacus libero, pretium at, lobortis vitae.
Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna.
Morbi ac orci et nisl hendrerit mollis, suspendisse ut massa, cras nec ante.
Pellentesque a nulla cum sociis natoque penatibus et magnis dis parturient.
Aliquam tincidunt urna, nulla ullamcorper vestibulum turpis.
Pellentesque cursus luctus mauris \citep{Luger2021}.

\bibliography{bib}

\end{document}

### 2024-03-15 src/tex/ms.tex
Initial setup

@@
 % Define document class
-\documentclass[twocolumn]{aastex631}
+\documentclass[aps,prd,twocolumn,superscriptaddress,preprintnumbers,nofootinbib]{revtex4-2}
+
 \usepackage{showyourwork}
+\usepackage{amsfonts,amssymb,amsmath}
+\usepackage[nolist,nohyperlinks]{acronym}
+\usepackage{bookmark}
+\usepackage{orcidlink}
+\usepackage{mathtools}
 
-% Begin!
 \begin{document}
 
-% Title
-\title{An open source scientific article}
+\title{Detector-frame cosmology with non-parametric methods}
 
-% Author list
-\author{@thomasckng}
+\author{Thomas C. K. Ng\,\orcidlink{0000-0002-9491-1598}}
+\email{thomas.ng@link.cuhk.edu.hk}
+\affiliation{Department of Physics, The Chinese University of Hong Kong, Shatin, Hong Kong}
 
-% Abstract with filler text
+\author{Stefano Rinaldi\,\orcidlink{0000-0001-5799-4155}}
+\email{}
+\affiliation{}
+
+\author{Otto A. Hannuksela\,\orcidlink{0000-0002-3887-7137}}
+\email{hannuksela@phy.cuhk.edu.hk}
+\affiliation{Department of Physics, The Chinese University of Hong Kong, Shatin, Hong Kong}
+
 \begin{abstract}
-    Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
-    Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae, felis.
-    Curabitur dictum gravida mauris, consectetuer id, vulputate a, magna.
-    Donec vehicula augue eu neque, morbi tristique senectus et netus et.
-    Mauris ut leo, cras viverra metus rhoncus sem, nulla et lectus vestibulum.
-    Phasellus eu tellus sit amet tortor gravida placerat.
-    Integer sapien est, iaculis in, pretium quis, viverra ac, nunc.
-    Praesent eget sem vel leo ultrices bibendum.
-    Aenean faucibus, morbi dolor nulla, malesuada eu, pulvinar at, mollis ac.
-    Curabitur auctor semper nulla donec varius orci eget risus.
-    Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
-    Duis eget orci sit amet orci dignissim rutrum.
+    \textsf{\color{red} [TODO: Abstract]}
 \end{abstract}
 
-% Main body with filler text
+\maketitle
+
+\begin{acronym}
+    \acro{GW}{gravitational wave}
+    \acro{LVK}{LIGO-Virgo-KAGRA Collaboration}
+    \acro{PE}{parameter estimation}
+\end{acronym}
+
 \section{Introduction}
-\label{sec:intro}
+\label{sec:introduction}
 
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit.
-Ut purus elit, vestibulum ut, placerat ac, adipiscing vitae, felis.
-Curabitur dictum gravida mauris, consectetuer id, vulputate a, magna.
-Donec vehicula augue eu neque, morbi tristique senectus et netus et.
-Mauris ut leo, cras viverra metus rhoncus sem, nulla et lectus vestibulum.
-Phasellus eu tellus sit amet tortor gravida placerat.
-Integer sapien est, iaculis in, pretium quis, viverra ac, nunc.
-Praesent eget sem vel leo ultrices bibendum.
-Aenean faucibus, morbi dolor nulla, malesuada eu, pulvinar at, mollis ac.
-Curabitur auctor semper nulla donec varius orci eget risus.
-Duis nibh mi, congue eu, accumsan eleifend, sagittis quis, diam.
-Duis eget orci sit amet orci dignissim rutrum.
+\section{Background}
+\label{sec:background}
 
-Nam dui ligula, fringilla a, euismod sodales, sollici- tudin vel, wisi.
-Morbi auctor lorem non justo, nam lacus libero, pretium at, lobortis vitae.
-Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet magna.
-Morbi ac orci et nisl hendrerit mollis, suspendisse ut massa, cras nec ante.
-Pellentesque a nulla cum sociis natoque penatibus et magnis dis parturient.
-Aliquam tincidunt urna, nulla ullamcorper vestibulum turpis.
-Pellentesque cursus luctus mauris \citep{Luger2021}.
+\section{Method}
+\label{sec:method}
+
+\section{Results}
+\label{sec:results}
+
+\section{Discussion}
+\label{sec:discussion}
+
+\section{Conclusion}
+\label{sec:conclusion}
+
+\begin{acknowledgments}
+
+This paper was compiled using \textsc{showyourwork} \cite{Luger2021} to facilitate reproducibility.
 

### 2024-03-19 src/tex/ms.tex
Update ms.tex

@@
 % Define document class
-\documentclass[aps,prd,twocolumn,superscriptaddress,preprintnumbers,nofootinbib]{revtex4-2}
+\documentclass[aps,prd,twocolumn,superscriptaddress,preprintnumbers,nofootinbib,hidelinks]{revtex4-2}
 
@@
 \usepackage{mathtools}
-
 \begin{document}
@@
 
-\author{Thomas C. K. Ng\,\orcidlink{0000-0002-9491-1598}}
+\author{Thomas~C.~K.~Ng\,\orcidlink{0000-0002-9491-1598}}
 \email{thomas.ng@link.cuhk.edu.hk}
@@
 
-\author{Stefano Rinaldi\,\orcidlink{0000-0001-5799-4155}}
-\email{}
-\affiliation{}
+\author{Stefano~Rinaldi\,\orcidlink{0000-0001-5799-4155}}
+\email{stefano.rinaldi@uni-heidelberg.de}
+\affiliation{Institut~für~Theoretische~Astrophysik, ZAH, Universität~Heidelberg, Albert-Ueberle-Stra{\ss}e~2, 69120 Heidelberg, Germany}
+\affiliation{Dipartimento di Fisica e Astronomia ``G. Galilei'', Università di Padova, Via Marzolo 8, 35122 Padova, Italy}
 
-\author{Otto A. Hannuksela\,\orcidlink{0000-0002-3887-7137}}
+\author{Otto~A.~Hannuksela\,\orcidlink{0000-0002-3887-7137}}
 \email{hannuksela@phy.cuhk.edu.hk}
@@
 
+S.R.~acknowledges financial support from the European Research Council for the ERC Consolidator grant DEMOBLACK, under contract no.~770017.
 This paper was compiled using \textsc{showyourwork} \cite{Luger2021} to facilitate reproducibility.
+
+\end{acknowledgments}
 

### 2024-04-03 src/tex/ms.tex
Update ms.tex
Add outline (WIP)

@@
-% Define document class
 \documentclass[aps,prd,twocolumn,superscriptaddress,preprintnumbers,nofootinbib,hidelinks]{revtex4-2}
@@
 \usepackage{mathtools}
+
+\newcommand{\todo}[1]{\textcolor{red}{[TODO: #1]}}
+
 \begin{document}
@@
 \begin{abstract}
-    \textsf{\color{red} [TODO: Abstract]}
+    \todo{Abstract}
 \end{abstract}
@@
 
-\section{Background}
-\label{sec:background}
+% Motivation
+    % Hubble tension
+    % GW can provide independent measurements of Hubble constant
+    % With EM counterpart, can measure redshift
+    % Without EM counterpart, can measure redshift from population statistics
+
+    % The challenge of understanding the Universe's dynamics, particularly the Hubble tension, requires precise measurements of the Hubble constant.
+    % Building upon the existing spectral siren method, which capitalizes on population information from gravitational wave sources, this talk explores an alternative way to analyse the population data to obtain the cosmological parameters.
+    % We'll focus on how non-parametric methods, which are flexible models that can be used to agnostically reconstruct arbitrary probability densities, can be incorporate in this framework and leverage on the detector-frame mass distribution to infer the cosmological parameters.
+
+% Previous work
+    % Spectral siren
+    % Other population studies
+
+% What's new?
+    % Non-parametric methods
+    % Detector-frame
+
+% Section outline
+In Sec.~\ref{sec:method},
+In Sec.~\ref{sec:results},
+In Sec.~\ref{sec:discussion},
+In Sec.~\ref{sec:conclusion},
 
@@
 \label{sec:method}
+
+In this section, we describe the framework of our method,
+
+\subsection{Framework}
+\label{sec:framework}
+
+\subsubsection{Non-parametric reconstruction of detector-frame observed population}
+\label{sec:reconstruction}
+
+\subsubsection{Transformation of source-frame population model}
+\label{sec:transformation}
+
+\subsubsection{Inference}
+\label{sec:inference}
+
+\subsection{Injection test}
+\label{sec:injection}
 

### 2024-04-03 src/tex/ms.tex
Add text

@@
 \begin{abstract}
-    \todo{Abstract}
+    The challenge of understanding the Universe's dynamics, particularly the Hubble tension, requires precise measurements of the Hubble constant.
+    Building upon the existing spectral siren method, which capitalizes on population information from gravitational wave sources, this study explores an alternative way to analyze the population data to obtain the cosmological parameters in $\Lambda$CDM.
+    We will focus on how non-parametric methods, which are flexible models that can be used to agnostically reconstruct arbitrary probability densities, can be incorporated into this framework and leverage the detector-frame mass distribution to infer the cosmological parameters.
+    \todo{Rewrite abstract}
 \end{abstract}
@@
     \acro{PE}{parameter estimation}
+    \acro{CMB}{cosmic microwave background}
+    \acro{EM}{electromagnetic}
 \end{acronym}
@@
 % Motivation
-    % Hubble tension
-    % GW can provide independent measurements of Hubble constant
-    % With EM counterpart, can measure redshift
+In recent years, the precision of cosmological measurements has improved significantly, leading to the discovery of the Hubble tension, a discrepancy between the value of the Hubble constant $H_0$ inferred from the \ac{CMB} \citep{Planck:2018vyg} and local measurements \citep{Riess:2021jrx}.
+This tension has motivated the search for new methods to measure $H_0$ with high precision, and \ac{GW} sources have emerged as a promising tool for this purpose.
+To measure $H_0$ from \ac{GW} sources, one needs not only the luminosity distance which can be inferred from the signal, but also the redshift of the source.
+Some \ac{GW} sources can be associated with \ac{EM} counterparts, allowing for the measurement of the redshift. \citep{LIGOScientific:2017adf}
     % Without EM counterpart, can measure redshift from population statistics
-
-    % The challenge of understanding the Universe's dynamics, particularly the Hubble tension, requires precise measurements of the Hubble constant.
-    % Building upon the existing spectral siren method, which capitalizes on population information from gravitational wave sources, this talk explores an alternative way to analyse the population data to obtain the cosmological parameters.
-    % We'll focus on how non-parametric methods, which are flexible models that can be used to agnostically reconstruct arbitrary probability densities, can be incorporate in this framework and leverage on the detector-frame mass distribution to infer the cosmological parameters.
 

### 2024-04-03 src/tex/ms.tex
Add text

@@
 This tension has motivated the search for new methods to measure $H_0$ with high precision, and \ac{GW} sources have emerged as a promising tool for this purpose.
-To measure $H_0$ from \ac{GW} sources, one needs not only the luminosity distance which can be inferred from the signal, but also the redshift of the source.
-Some \ac{GW} sources can be associated with \ac{EM} counterparts, allowing for the measurement of the redshift. \citep{LIGOScientific:2017adf}
-    % Without EM counterpart, can measure redshift from population statistics
+To infer $H_0$ from \ac{GW} sources, one needs not only the luminosity distance which can be inferred from the signal, but also the redshift of the source.
+Some \ac{GW} sources can be associated with \ac{EM} counterparts, allowing for the measurement of the redshift, e.g., GW170817 \citep{LIGOScientific:2017adf, Guidorzi:2017ogy}.
+However, most \ac{GW} sources do not have \ac{EM} counterparts.
+In this case, one can use population statistics to infer $H_0$ by marginalizing over the redshift distribution of the sources.
+This method is known as the spectral siren \citep{You:2020wju, Mastrogiovanni:2021wsd, LIGOScientific:2021aug, Ezquiaga:2022zkx}.
 
 % Previous work
-    % Spectral siren
-    % Other population studies
+The fundamental idea of the spectral siren method is to assume a population model for the mass distribution of the sources and marginalize over the redshift distribution to infer the cosmological parameters.
+Many studies have explored different variations of the spectral siren method.
+\todo{Summarize previous work}
 
 % What's new?
-    % Non-parametric methods
-    % Detector-frame
+In this study, we explore an alternative method based on the spectral siren framework.
 

### 2024-04-03 src/tex/ms.tex
Add introduction

@@
 Many studies have explored different variations of the spectral siren method.
-\todo{Summarize previous work}
+\todo{Summarize previous work and add references}
 
@@
 In this study, we explore an alternative method based on the spectral siren framework.
+We propose two major modifications to the standard spectral siren method.
+First, instead of transforming the luminosity distance to redshift during the inference process, we transform the population model to the detector frame.
+Second, instead of computing the likelihood of the observed data given the population model, we reconstruct the observed population directly from the data using non-parametric methods.
+With these modifications, we aim to separate the spectral siren method into two distinct parts.
+The first part is the reconstruction of the observed population from the data, which does not require any assumptions about the population model, and the intermediate result is completely data-driven.
+All the assumptions including the population model, cosmological model and the selection function are then incorporated in the second half.
+This separation allows us to explore the impact of different assumptions easily.
+\todo{Add references}
 
 % Section outline
-In Sec.~\ref{sec:method},
-In Sec.~\ref{sec:results},
-In Sec.~\ref{sec:discussion},
-In Sec.~\ref{sec:conclusion},
+In Sec.~\ref{sec:method}, we describe our modified method and the testing procedure.
+In Sec.~\ref{sec:results}, we present the results of our method.
+In Sec.~\ref{sec:discussion}, we compare our result with other methods and discuss future directions.
+Finally, we conclude in Sec.~\ref{sec:conclusion} with a summary.
+\todo{Rewrite outline}
 

### 2024-04-08 src/tex/ms.tex
Update ms.tex

@@
 We propose two major modifications to the standard spectral siren method.
-First, instead of transforming the luminosity distance to redshift during the inference process, we transform the population model to the detector frame.
+First, instead of transforming the luminosity distance to redshift during the \ac{PE} process, we transform the population model to the detector frame.
 Second, instead of computing the likelihood of the observed data given the population model, we reconstruct the observed population directly from the data using non-parametric methods.
@@
 
-In this section, we describe the framework of our method,
+In this section, we describe the framework of our method, followed by a description of the testing setup.
 
@@
 \label{sec:framework}
+
+Our method consists of three main steps: the non-parametric reconstruction of the observed population, the transformation of the source-frame population model to the detector frame, and the \ac{PE}.
+We show the flowchart of our method in Fig.~XXX. \todo{Add flowchart}
 
@@
 
-\subsubsection{Inference}
-\label{sec:inference}
+\subsubsection{Parameter estimation}
+\label{sec:pe}
 

### 2024-04-12 src/tex/ms.tex
Add method framework

@@
     \acro{EM}{electromagnetic}
+    \acro{DPGMM}{Dirichlet process Gaussian mixture model}
+    \acro{BBH}{binary black hole}
+    \acro{GWTC-3}{the third Gravitational-Wave Transient Catalog}
 \end{acronym}
@@
 
+The first step of our method is to reconstruct the observed population.
+we choose to consider the primary mass $m_1$ of the \ac{BBH} sources.
+From the posterior samples obtained from the \ac{PE} of each \ac{GW} event, we can reconstruct the observed detector-frame mass distribution $p(m_1|\{d\})$, where $\{d\}$ is the set of all the observed data.
+
+To perform the reconstruction, we use a non-parametric reconstruction package \textsc{FIGARO} \citep{Rinaldi:2022kyg}.
+\textsc{FIGARO} can reconstruct an arbitrary probability density function from a set of samples based on the \ac{DPGMM}.
+It can also perform a hierarchy of \ac{DPGMM} to reconstruct an polulation distribution with posterior samples from multiple events \citep{Rinaldi:2021bhm}.
+The output of \textsc{FIGARO} is a set of distributions, where each distribution is a realization of the observed population that is consistent with the data.
+Therefore, the set of distributions represents the uncertainty in the reconstruction procedure.
+In Fig.~XXX \todo{Add figure}, we show an example of the reconstruction obtained with \textsc{FIGARO}.
+
+This part of the method is completely data-driven and does not require any assumptions about the population model or the cosmological model.
+Therefore the intermediate result is independent of the assumptions, and we can reuse the result for different population models, cosmological models, or selection functions.
+This reduces the computational cost of the method, as we only need to perform the reconstruction of the observed population once.
+
 \subsubsection{Transformation of source-frame population model}
 \label{sec:transformation}
+
+On the other hand, we need a source-frame population model $p(m_1|\Lambda)$ to represent the underlying population of the sources, where $\Lambda$ is the polulation model.
+We can either use a parametric model to only assume the form of the population distribution, or we can use a fixed population model to assume the exact population distribution.
+With $p(m_1|\Lambda)$, we need to transform it to the detector frame with a given cosmological model to obtain the detector-frame population model $p(m_1|\Lambda, \Omega)$, where $\Omega$ is the cosmological model.
+In this study, we use the $\Lambda$CDM model with the Planck 2018 cosmological parameters \citep{Planck:2018vyg}, except for $H_0$ which we choose to be a free parameter.
+With different $\Omega$, $p(m_1|\Lambda)$ will be transformed differently, and the resulting $p(m_1|\Lambda, \Omega)$ will be different.
+In Fig.~XXX \todo{Add figure}, we show an example of the transformation without any selection effects.
+
+After the cosmological transformation, we need to consider the selection effects.
+The selection effects can be easily incorporated by applying the selection function to the transformed population.
+\todo{Add more details. Parametric or fixed?}
+
+Opposite to the first part of the method, which is completely data-driven, this part includes all the assumptions about the population model, the cosmological model, and the selection function.
 
@@
 
+With $p(m_1|\{d\})$ and $p(m_1|\Lambda, \Omega)$ obtained from Sec.~\ref{sec:reconstruction} and Sec.~\ref{sec:transformation} respectively, we can perform the \ac{PE} to infer the free parameters in the parametric models assumed in Sec.~\ref{sec:transformation}.
+The \ac{PE} is performed by comparing the observed population $p(m_1|\{d\})$ with the transformed population $p(m_1|\Lambda, \Omega)$.
+To compare the two populations, we use the Jensen–Shannon distance $d_{JS}$ as the distance measure.
+$d_{JS}$ is calculated as
+\begin{equation}
+    d_{JS}(p, q) = \sqrt{\frac{D_{KL}(p||m) + D_{KL}(q||m)}{2}},
+\end{equation}
+where $p$ and $q$ are the two distributions to be compared, $m = (p + q) / 2$, and $D_{KL}$ is the Kullback–Leibler divergence.
+$d_{JS}$ is a symmetric and smoothed version of $D_{KL}$, and it is bounded between 0 and 1.
+For each reconstructed population $p(m_1|\{d\})$, we find the best-fit parameters by minimizing $d_{JS}$ with respect to the free parameters.
+This procedure gives us a set of best-fit parameters for each reconstructed population.
+Note that the resulting distribution of the best-fit parameters is not a posterior distribution in the Bayesian sense, as we do not consider the likelihood of the observed data given the population model.
+Instead, it is a distribution of the best-fit parameters that are consistent with the observed data, distributed according to the uncertainty in the reconstruction of the observed population.
+In Fig.~XXX \todo{Add figure}, we show an example of the fitting procedure.
+
 \subsection{Injection test}
 \label{sec:injection}
+
+\subsection{GWTC-3 test}
+\label{sec:gwtc3}
 
@@
 \label{sec:discussion}
+
+% Comparison with other methods
+
+% Future directions
+    % Likelihood
+    % luminosity distance/redshift polulation
 

### 2024-05-02 src/tex/ms.tex
Modify SF text
Move selection function to the reconstruction process in the text

--- 2024-04-03 src/tex/ms.tex
@@
     \acro{EM}{electromagnetic}
+    \acro{(H)DPGMM}{hierarchy of Dirichlet process Gaussian mixture model}
+    \acro{BBH}{binary black hole}
+    \acro{GWTC-3}{the third Gravitational-Wave Transient Catalog}
 \end{acronym}
@@
 We propose two major modifications to the standard spectral siren method.
-First, instead of transforming the luminosity distance to redshift during the inference process, we transform the population model to the detector frame.
-Second, instead of computing the likelihood of the observed data given the population model, we reconstruct the observed population directly from the data using non-parametric methods.
+First, instead of transforming the individual posterior samples from detector frame to source frame during \ac{PE} process, we transform the population model to the detector frame before performing \ac{PE}.
+Second, instead of computing the likelihood of the observed data given the population model, we reconstruct the detector-frame intrinsic population directly from the data using non-parametric method, where detector-frame intrinsic population is the detector-frame polulation after removing the selection effects.
+
 With these modifications, we aim to separate the spectral siren method into two distinct parts.
-The first part is the reconstruction of the observed population from the data, which does not require any assumptions about the population model, and the intermediate result is completely data-driven.
-All the assumptions including the population model, cosmological model and the selection function are then incorporated in the second half.
+The first part is the reconstruction of the observed intrinsic population from the data, which does not require any assumptions about the population model, and the intermediate result only depends on the data and the assumption on the selection function.
+other assumptions including the population model and cosmological model are then incorporated in the second half.
 This separation allows us to explore the impact of different assumptions easily.
+The intermediate result of the first part can be reused as long as the data and the assumption on the selection function are the same.
+The assumption on the selection function usually does not change in population studies, as the selection function is determined by the detector sensitivity.
 \todo{Add references}
@@
 
-In this section, we describe the framework of our method,
+Our method consists of three main steps: the non-parametric reconstruction of the detector-frame intrinsic population, the transformation of the source-frame population model to the detector frame, and the \ac{PE}.
+We show the flowchart of our method in Fig.~XXX. \todo{Add flowchart}
 
-\subsection{Framework}
-\label{sec:framework}
-
-\subsubsection{Non-parametric reconstruction of detector-frame observed population}
+\subsection{Non-parametric reconstruction of detector-frame intrinsic population}
 \label{sec:reconstruction}
 
-\subsubsection{Transformation of source-frame population model}
+The first step of our method is to reconstruct the detector-frame intrinsic population.
+we choose to consider the primary mass $m_1$ of the \ac{BBH} sources, in principle, we can consider more parameters at once.
+From the posterior samples obtained from the \ac{PE} of each \ac{GW} event, we reconstruct the detector-frame intrinsic mass distribution $p(m^z_1|\mathbf{d})$, where $\mathbf{d}$ is the observed data of the \ac{GW} events.
+
+To perform the reconstruction, we use a non-parametric reconstruction package \textsc{FIGARO} \citep{Rinaldi:2022kyg}.
+Using \textsc{FIGARO}, we perform a \ac{(H)DPGMM} to reconstruct $p(m^z_1|\Xi_i(\mathbf{d}))$ with samples from single-event \ac{PE} posterior $p(m^z_1|d_i)$, where $d_i$ is the observed data of a single \ac{GW} event \citep{Rinaldi:2021bhm}, and $\Xi_i$ is the hyperparameters of the \ac{(H)DPGMM} for a single reconstruction.
+In the process of the reconstruction, selection effects are taken into account by \todo{Add equation} \citep{Mandel:2018mve}.
+
+Performing the reconstruction once will give a distribution $p(m^z_1|\Xi_i(\mathbf{d}))$ that is consistent with $\mathbf{d}$ with a given set of hyperparameters $\Xi_i$.
+To account for the uncertainty in the reconstruction, we need to perform the reconstruction multiple times to obtain a set of distributions $\{p(m^z_1|\Xi_i(\mathbf{d}))\}$.
+The uncertainty of the reconstruction can then be quantified by the spread of the distributions.
+In Fig.~XXX \todo{Add figure}, we show an example of the reconstruction obtained with \textsc{FIGARO}.
+
+This part of the method is mostly data-driven and only requires the assumption on the selection function.
+Therefore the intermediate result is independent of the population and cosmological model, and we can reuse the result for testing different models.
+This reduces the computational cost of the method, as we only need to perform the reconstruction once.
+
+\subsection{Transformation of source-frame population model}
 \label{sec:transformation}
 
-\subsubsection{Inference}
-\label{sec:inference}
+On the other hand, we need a source-frame population model $p(m_1|\Lambda)$ to represent the underlying population of the sources, where $\Lambda$ is the polulation model.
+We can either use a parametric model to only assume the form of the population distribution, or we can use a fixed population model to assume the exact population distribution.
+With $p(m_1|\Lambda)$, we need to transform it to the detector frame with a given cosmological model to obtain the detector-frame population model $p(m^z_1|\Lambda, \Omega)$, where $\Omega$ is the cosmological model.
+In this study, we use the $\Lambda$CDM model with the Planck 2018 cosmological parameters \citep{Planck:2018vyg}, except for $H_0$ which we choose to be a free parameter.
+We choose only $H_0$ as a free parameter because the current detection of \ac{GW} sources is local and the effect of $H_0$ dominates other cosmological parameters.
+With different $H_0$, $p(m_1|\Lambda)$ will be transformed differently, and the resulting $p(m^z_1|\Lambda, H_0)$ will be different.
+By considering a continuous set of $H_{0,j}$ and $\Lambda_j$, we can obtain a set of $p(m^z_1|\Lambda_j, H_{0,j})$, which is a set of continuous detector-frame intrinsic population model for each combination of $H_{0,j}$ and $\Lambda_j$.\footnote{In practice, we use a discrete set of $H_{0,j}$ and $\Lambda_j$ to approximate the continuous set.}
+In Fig.~XXX \todo{Add figure}, we show an example of the transformation.
 
-\subsection{Injection test}
-\label{sec:injection}
+Opposite to the first part of the method, this part includes the assumptions about the population and cosmological model.
+
+\subsection{Parameter estimation}
+\label{sec:pe}
+
+With $\{p(m^z_1|\Xi_i(\mathbf{d}))\}$ and $\{p(m^z_1|\Lambda_j, H_{0,j})\}$ obtained from Sec.~\ref{sec:reconstruction} and Sec.~\ref{sec:transformation} respectively, we can perform the \ac{PE} to infer the free parameters in the parametric models assumed in Sec.~\ref{sec:transformation}.
+The \ac{PE} is performed by comparing $\{p(m^z_1|\Xi_i(\mathbf{d}))\}$ with $\{p(m^z_1|\Lambda_j, H_{0,j})\}$.
+To compare the two populations, we use the Jensen–Shannon distance $d_{JS}$ as the distance measure.
+$d_{JS}$ is calculated as
+\begin{equation}
+    d_{JS}(p, q) = \sqrt{\frac{D_{KL}(p||m) + D_{KL}(q||m)}{2}},
+\end{equation}
+where $p$ and $q$ are the two distributions to be compared, $m = (p + q) / 2$, and $D_{KL}$ is the Kullback–Leibler divergence.
+$d_{JS}$ is a symmetric and smoothed version of $D_{KL}$, and it is bounded between 0 and 1.
+
+For each reconstructed distribution $p(m^z_1|\Xi_i(\mathbf{d}))$, we find the best-fit $\Lambda_j$ and $H_{0,j}$ by minimizing $d_{JS}$.
+This procedure gives us a set of best-fit $H_{0,i}$ and $\Lambda_i$ for each $\Xi_i$.
+Note that the resulting distribution of the best-fit parameters is not a posterior distribution in the Bayesian sense, as we do not consider the likelihood of the observed data given the population model.
+Instead, it is a distribution of the best-fit parameters that are consistent with the observed data, distributed according to the uncertainty in the reconstruction of the population.
+That is, the uncertainty in the comparison of the two populations is not considered in the resulting distribution.
+By considering this optimization procedure only, we can estimate the best-fit parameters without performing computationally expensive likelihood calculations.
+In Fig.~XXX \todo{Add figure}, we show an example of the fitting procedure.
+
+\subsection{Minimization of $d_{JS}$}
+\label{sec:minimization}
+
+\section{Analysis}
+\label{sec:analysis}
+
+\subsection{Injection}
+\label{sec:injection-setup}
+
+\subsection{Real data}
+\label{sec:real-data-setup}
 
@@
 
+\subsection{Injection}
+\label{sec:injection-results}
+
+% $H_0$ only
+
+% $H_0$ and $\Lambda$
+
+\subsection{Real data}
+\label{sec:real-data-results}
+
+% Power law + peak
+
+% Population model from simulation
+
 \section{Discussion}
 \label{sec:discussion}
+
+% Comparison with other methods
+
+% Future directions
+    % Likelihood
+    % luminosity distance/redshift polulation
 

### 2024-05-02 src/tex/ms.tex
Update ms.tex

@@
 This tension has motivated the search for new methods to measure $H_0$ with high precision, and \ac{GW} sources have emerged as a promising tool for this purpose.
-To infer $H_0$ from \ac{GW} sources, one needs not only the luminosity distance which can be inferred from the signal, but also the redshift of the source.
+To infer $H_0$ from \ac{GW} sources, one needs not only the luminosity distance which can be inferred from the signal but also the redshift of the source.
 Some \ac{GW} sources can be associated with \ac{EM} counterparts, allowing for the measurement of the redshift, e.g., GW170817 \citep{LIGOScientific:2017adf, Guidorzi:2017ogy}.
 However, most \ac{GW} sources do not have \ac{EM} counterparts.
-In this case, one can use population statistics to infer $H_0$ by marginalizing over the redshift distribution of the sources.
+In this case, one can use population statistics to infer $H_0$ by marginalizing the redshift distribution of the sources.
 This method is known as the spectral siren \citep{You:2020wju, Mastrogiovanni:2021wsd, LIGOScientific:2021aug, Ezquiaga:2022zkx}.
@@
 We propose two major modifications to the standard spectral siren method.
-First, instead of transforming the individual posterior samples from detector frame to source frame during \ac{PE} process, we transform the population model to the detector frame before performing \ac{PE}.
-Second, instead of computing the likelihood of the observed data given the population model, we reconstruct the detector-frame intrinsic population directly from the data using non-parametric method, where detector-frame intrinsic population is the detector-frame polulation after removing the selection effects.
+First, instead of transforming the individual posterior samples from the detector frame to the source frame during the \ac{PE} process, we transform the population model to the detector frame before performing \ac{PE}.
+Second, instead of computing the likelihood of the observed data given the population model, we reconstruct the detector-frame intrinsic population directly from the data using a non-parametric method, where the detector-frame intrinsic population is the detector-frame population after removing the selection effects.
 
 With these modifications, we aim to separate the spectral siren method into two distinct parts.
-The first part is the reconstruction of the observed intrinsic population from the data, which does not require any assumptions about the population model, and the intermediate result only depends on the data and the assumption on the selection function.
+The first part is the reconstruction of the observed intrinsic population from the data, which does not require any assumptions about the population model, and the intermediate result only depends on the data and the assumption of the selection function.
 other assumptions including the population model and cosmological model are then incorporated in the second half.
@@
 
-This part of the method is mostly data-driven and only requires the assumption on the selection function.
+This part of the method is mostly data-driven and only requires the assumption of the selection function.
 Therefore the intermediate result is independent of the population and cosmological model, and we can reuse the result for testing different models.
@@
 
-On the other hand, we need a source-frame population model $p(m_1|\Lambda)$ to represent the underlying population of the sources, where $\Lambda$ is the polulation model.
+On the other hand, we need a source-frame population model $p(m_1|\Lambda)$ to represent the underlying population of the sources, where $\Lambda$ is the population model.
 We can either use a parametric model to only assume the form of the population distribution, or we can use a fixed population model to assume the exact population distribution.
@@
 With different $H_0$, $p(m_1|\Lambda)$ will be transformed differently, and the resulting $p(m^z_1|\Lambda, H_0)$ will be different.
-By considering a continuous set of $H_{0,j}$ and $\Lambda_j$, we can obtain a set of $p(m^z_1|\Lambda_j, H_{0,j})$, which is a set of continuous detector-frame intrinsic population model for each combination of $H_{0,j}$ and $\Lambda_j$.\footnote{In practice, we use a discrete set of $H_{0,j}$ and $\Lambda_j$ to approximate the continuous set.}
+By considering a continuous set of $H_{0,j}$ and hyperparameters $\lambda_j$ of the population model, we can obtain a set of $p(m^z_1|\Lambda(\lambda_j), H_{0,j})$, which is a set of continuous detector-frame intrinsic population models for each combination of $H_{0,j}$ and $\Lambda(\lambda_j)$.\footnote{In practice, we use a discrete set of $H_{0,j}$ and $\Lambda(\lambda_j)$ with a small step size to approximate the continuous set.}
 In Fig.~XXX \todo{Add figure}, we show an example of the transformation.
@@
     % Likelihood
-    % luminosity distance/redshift polulation
+    % luminosity distance/redshift population
 

### 2024-05-14 src/tex/ms.tex
[showyourwork] sync

@@
 Some \ac{GW} sources can be associated with \ac{EM} counterparts, allowing for the measurement of the redshift, e.g., GW170817 \citep{LIGOScientific:2017adf, Guidorzi:2017ogy}.
-However, most \ac{GW} sources do not have \ac{EM} counterparts.
+However, most \ac{GW} sources do not have \ac{EM} counterparts. \todo{Add galaxy catalog method}
 In this case, one can use population statistics to infer $H_0$ by marginalizing the redshift distribution of the sources.
@@
 \begin{acknowledgments}
+T.N.~ and S.R.~ acknowledge financial support from the German Excellence Strategy via the Heidelberg Cluster of Excellence (EXC 2181 - 390900948) STRUCTURES.
+S.R.~ acknowledges financial support from the European Research Council for the ERC Consolidator grant DEMOBLACK, under contract no. 770017.
 
-S.R.~acknowledges financial support from the European Research Council for the ERC Consolidator grant DEMOBLACK, under contract no.~770017.
+This research made use of the bwForCluster Helix: the authors acknowledge support by the state of Baden-Württemberg through bwHPC and the German Research Foundation (DFG) through grant INST 35/1597-1 FUGG.
+
 This paper was compiled using \textsc{showyourwork} \cite{Luger2021} to facilitate reproducibility.

### 2024-05-16 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
-\title{Detector-frame cosmology with non-parametric methods}
+\title{Inferring cosmology from gravitational waves using non-parametric detector-frame mass distribution}
 
@@
 First, instead of transforming the individual posterior samples from the detector frame to the source frame during the \ac{PE} process, we transform the population model to the detector frame before performing \ac{PE}.
-Second, instead of computing the likelihood of the observed data given the population model, we reconstruct the detector-frame intrinsic population directly from the data using a non-parametric method, where the detector-frame intrinsic population is the detector-frame population after removing the selection effects.
+Second, instead of computing the likelihood of the observed data given the population model, we reconstruct the detector-frame observed population directly from the data using a non-parametric method.
 
 With these modifications, we aim to separate the spectral siren method into two distinct parts.
-The first part is the reconstruction of the observed intrinsic population from the data, which does not require any assumptions about the population model, and the intermediate result only depends on the data and the assumption of the selection function.
-other assumptions including the population model and cosmological model are then incorporated in the second half.
+The first part is the reconstruction of the observed population from the data, which does not require any assumptions about the population model, and the intermediate result is completely data-driven.
+Other assumptions including the population model, cosmological model and selection function are then incorporated in the second half.
 This separation allows us to explore the impact of different assumptions easily.
-The intermediate result of the first part can be reused as long as the data and the assumption on the selection function are the same.
-The assumption on the selection function usually does not change in population studies, as the selection function is determined by the detector sensitivity.
+The intermediate result of the first part can be reused as long as the data does not change, reducing the computational cost of the method.
 \todo{Add references}
@@
 
-Our method consists of three main steps: the non-parametric reconstruction of the detector-frame intrinsic population, the transformation of the source-frame population model to the detector frame, and the \ac{PE}.
+Our method consists of three main steps: the non-parametric reconstruction of the detector-frame observed population, the transformation of the source-frame population model to the detector frame, and the \ac{PE}.
 We show the flowchart of our method in Fig.~XXX. \todo{Add flowchart}
 
-\subsection{Non-parametric reconstruction of detector-frame intrinsic population}
+\subsection{Non-parametric reconstruction of detector-frame observed population}
 \label{sec:reconstruction}
 
-The first step of our method is to reconstruct the detector-frame intrinsic population.
+The first step of our method is to reconstruct the detector-frame observed population.
 we choose to consider the primary mass $m_1$ of the \ac{BBH} sources, in principle, we can consider more parameters at once.
-From the posterior samples obtained from the \ac{PE} of each \ac{GW} event, we reconstruct the detector-frame intrinsic mass distribution $p(m^z_1|\mathbf{d})$, where $\mathbf{d}$ is the observed data of the \ac{GW} events.
+From the posterior samples obtained from the \ac{PE} of each \ac{GW} event, we reconstruct the detector-frame observed mass distribution $p(m^z_1|\mathbf{Y})$, where $\mathbf{Y}$ is the set of posterior samples of all \ac{GW} events.
 
 To perform the reconstruction, we use a non-parametric reconstruction package \textsc{FIGARO} \citep{Rinaldi:2022kyg}.
-Using \textsc{FIGARO}, we perform a \ac{(H)DPGMM} to reconstruct $p(m^z_1|\Xi_i(\mathbf{d}))$ with samples from single-event \ac{PE} posterior $p(m^z_1|d_i)$, where $d_i$ is the observed data of a single \ac{GW} event \citep{Rinaldi:2021bhm}, and $\Xi_i$ is the hyperparameters of the \ac{(H)DPGMM} for a single reconstruction.
-In the process of the reconstruction, selection effects are taken into account by \todo{Add equation} \citep{Mandel:2018mve}.
+Using \textsc{FIGARO}, we perform a \ac{(H)DPGMM} to reconstruct $p(m^z_1|\Theta_i(\mathbf{Y}))$ with samples from all single-event \ac{PE} posterior \citep{Rinaldi:2021bhm}, and $\Theta_i$ is the hyperparameters of the \ac{(H)DPGMM} for a single reconstruction.
 
-Performing the reconstruction once will give a distribution $p(m^z_1|\Xi_i(\mathbf{d}))$ that is consistent with $\mathbf{d}$ with a given set of hyperparameters $\Xi_i$.
-To account for the uncertainty in the reconstruction, we need to perform the reconstruction multiple times to obtain a set of distributions $\{p(m^z_1|\Xi_i(\mathbf{d}))\}$.
+Performing the reconstruction once will give a distribution $p(m^z_1|\Theta_i(\mathbf{Y}))$ that is consistent with $\mathbf{Y}$ with a given set of hyperparameters $\Theta_i$.
+To account for the uncertainty in the reconstruction, we need to perform the reconstruction multiple times to obtain a set of distributions $p(m^z_1|\mathbf{\Theta}(\mathbf{Y}))$, where $\mathbf{\Theta}$ is a set of $\Theta_i$.
 The uncertainty of the reconstruction can then be quantified by the spread of the distributions.
@@
 
-This part of the method is mostly data-driven and only requires the assumption of the selection function.
-Therefore the intermediate result is independent of the population and cosmological model, and we can reuse the result for testing different models.
-This reduces the computational cost of the method, as we only need to perform the reconstruction once.
+This part of the method is completely data-driven.
+Therefore the intermediate result is independent of the population, cosmological model and selection function, and we can reuse the result for testing different models.
+This reduces the computational cost of the method, as we only need to perform the reconstruction once to analyze different models.
 
@@
 With $p(m_1|\Lambda)$, we need to transform it to the detector frame with a given cosmological model to obtain the detector-frame population model $p(m^z_1|\Lambda, \Omega)$, where $\Omega$ is the cosmological model.
-In this study, we use the $\Lambda$CDM model with the Planck 2018 cosmological parameters \citep{Planck:2018vyg}, except for $H_0$ which we choose to be a free parameter.
-We choose only $H_0$ as a free parameter because the current detection of \ac{GW} sources is local and the effect of $H_0$ dominates other cosmological parameters.
-With different $H_0$, $p(m_1|\Lambda)$ will be transformed differently, and the resulting $p(m^z_1|\Lambda, H_0)$ will be different.
-By considering a continuous set of $H_{0,j}$ and hyperparameters $\lambda_j$ of the population model, we can obtain a set of $p(m^z_1|\Lambda(\lambda_j), H_{0,j})$, which is a set of continuous detector-frame intrinsic population models for each combination of $H_{0,j}$ and $\Lambda(\lambda_j)$.\footnote{In practice, we use a discrete set of $H_{0,j}$ and $\Lambda(\lambda_j)$ with a small step size to approximate the continuous set.}
+With different $\Omega$, $p(m_1|\Lambda)$ will be transformed differently, and the resulting $p(m^z_1|\Lambda, \Omega)$ will be different.
+By considering a continuous set of $\Omega$ and  $\Lambda$ of the population model, we can obtain a set of $p(m^z_1|\mathbf{\Lambda}, \mathbf{\Omega})$, which is a set of continuous detector-frame intrinsic population models for each combination of $\Omega_j$ and $\Lambda_j$, where $\mathbf{\Lambda} = \{\Lambda_j\}$ and $\mathbf{\Omega} = \{\Omega_j\}$.\footnote{In practice, we use a discrete set of $\Omega_j$ and $\Lambda_j$ with a small step size in their hyperparameters to approximate the continuous set.}
+After obtaining $p(m^z_1|\mathbf{\Lambda}, \mathbf{\Omega})$, we can perform the \ac{PE} to infer the best-fit $\Omega_j$ and $\Lambda_j$.
+\todo{Add selection function}
 In Fig.~XXX \todo{Add figure}, we show an example of the transformation.
 
-Opposite to the first part of the method, this part includes the assumptions about the population and cosmological model.
+Opposite to the first part of the method, this part includes the assumptions about the population and cosmological model, as well as the selection function.
 
@@
 
-With $\{p(m^z_1|\Xi_i(\mathbf{d}))\}$ and $\{p(m^z_1|\Lambda_j, H_{0,j})\}$ obtained from Sec.~\ref{sec:reconstruction} and Sec.~\ref{sec:transformation} respectively, we can perform the \ac{PE} to infer the free parameters in the parametric models assumed in Sec.~\ref{sec:transformation}.
-The \ac{PE} is performed by comparing $\{p(m^z_1|\Xi_i(\mathbf{d}))\}$ with $\{p(m^z_1|\Lambda_j, H_{0,j})\}$.
+With $p(m^z_1|\mathbf{\Theta}(\mathbf{Y}))$ and $p(m^z_1|\mathbf{\Lambda}, \mathbf{\Omega})$ obtained from Sec.~\ref{sec:reconstruction} and Sec.~\ref{sec:transformation} respectively, we can perform the \ac{PE} to infer the free parameters in the parametric models assumed in Sec.~\ref{sec:transformation}.
+The \ac{PE} is performed by comparing $p(m^z_1|\Theta_i(\mathbf{Y}))$ with $p(m^z_1|\Lambda_j, \Omega_j)$.
 To compare the two populations, we use the Jensen–Shannon distance $d_{JS}$ as the distance measure.
@@
 
-For each reconstructed distribution $p(m^z_1|\Xi_i(\mathbf{d}))$, we find the best-fit $\Lambda_j$ and $H_{0,j}$ by minimizing $d_{JS}$.
-This procedure gives us a set of best-fit $H_{0,i}$ and $\Lambda_i$ for each $\Xi_i$.
+For each reconstructed distribution $p(m^z_1|\Theta_i(\mathbf{Y}))$, we find the best-fit $\Lambda_j$ and $\Omega_j$ by minimizing $d_{JS}$.
+This procedure gives us a set of best-fit $\Lambda_i$ and $\Omega_j$ for each $\Theta_i$.
 Note that the resulting distribution of the best-fit parameters is not a posterior distribution in the Bayesian sense, as we do not consider the likelihood of the observed data given the population model.

### 2024-05-16 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
-\subsection{Injection}
+\subsection{Simulation}
 \label{sec:injection-setup}
@@
 \label{sec:real-data-setup}
+Data retrived from \citet{LIGOScientific:2019lzm, KAGRA:2023pio}
 
@@
 
-\subsection{Injection}
+\subsection{Simulation}
 \label{sec:injection-results}
@@
 
+This research has made use of data or software obtained from the Gravitational Wave Open Science Center (gwosc.org), a service of the LIGO Scientific Collaboration, the Virgo Collaboration, and KAGRA.
+This material is based upon work supported by NSF's LIGO Laboratory which is a major facility fully funded by the National Science Foundation, as well as the Science and Technology Facilities Council (STFC) of the United Kingdom, the Max-Planck-Society (MPS), and the State of Niedersachsen/Germany for support of the construction of Advanced LIGO and construction and operation of the GEO600 detector.
+Additional support for Advanced LIGO was provided by the Australian Research Council.
+Virgo is funded, through the European Gravitational Observatory (EGO), by the French Centre National de Recherche Scientifique (CNRS), the Italian Istituto Nazionale di Fisica Nucleare (INFN) and the Dutch Nikhef, with contributions by institutions from Belgium, Germany, Greece, Hungary, Ireland, Japan, Monaco, Poland, Portugal, Spain.
+KAGRA is supported by Ministry of Education, Culture, Sports, Science and Technology (MEXT), Japan Society for the Promotion of Science (JSPS) in Japan; National Research Foundation (NRF) and Ministry of Science and ICT (MSIT) in Korea; Academia Sinica (AS) and National Science and Technology Council (NSTC) in Taiwan.
+
 This paper was compiled using \textsc{showyourwork} \cite{Luger2021} to facilitate reproducibility.

### 2024-05-17 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \newcommand{\todo}[1]{\textcolor{red}{[TODO: #1]}}
+\newcommand{\ste}[1]{\textcolor{blue}{[Ste: #1]}}
 
@@
 In this case, one can use population statistics to infer $H_0$ by marginalizing the redshift distribution of the sources.
-This method is known as the spectral siren \citep{You:2020wju, Mastrogiovanni:2021wsd, LIGOScientific:2021aug, Ezquiaga:2022zkx}.
+This method is known as the spectral siren method\citep{You:2020wju, Mastrogiovanni:2021wsd, LIGOScientific:2021aug, Ezquiaga:2022zkx}.
+\ste{You should mention that the galaxy catalog method also exists and cite some more relevant papers (e.g.
+Schutz 1986, Del Pozzo 2012, Farr et al.
+2019).
+References \texttt{\href{ https://onlinelibrary.wiley.com/doi/pdf/10.1002/andp.202200180}{here}}}
 
@@
 % What's new?
+\ste{Instead of immediately describing what we will do in the paper, I think you should add a brief description of non-parametric methods and then guide the reader to the same idea that we had (nonpar methods can reconstruct whatever distribution we want $\rightarrow$ cosmology must be involved to reconstruct astro dist $\rightarrow$ detector-frame dist `freezes' the cosmology and can be predicted using astro models + cosmological parameters $\rightarrow$ comparison with observations).
+Then you can briefly sketch what we present in the paper and summarise its content.
+It's more appealing and easy to understand if instead of presenting the reader with a bunch of facts you make them think with yow, somewhat like a lecture.}
 In this study, we explore an alternative method based on the spectral siren framework.
@@
 
+\ste{I see what you're doing here, and it's ok, but I'd present this as `benefits' of this approach, rather than our aims.}
 With these modifications, we aim to separate the spectral siren method into two distinct parts.
@@
 \todo{Add references}
+\todo{More events in the future}
 
@@
 \label{sec:method}
-
+\ste{In this section, we present the framework we developed to infer the cosmological parameters leveraging on the detector-frame mass distribution.
+In the remainder of this Section, we will make use of the following notation, mutuated from Rinaldi et al.
+(2024?
+2022?): ...}
 Our method consists of three main steps: the non-parametric reconstruction of the detector-frame observed population, the transformation of the source-frame population model to the detector frame, and the \ac{PE}.
@@
 
-To perform the reconstruction, we use a non-parametric reconstruction package \textsc{FIGARO} \citep{Rinaldi:2022kyg}.
-Using \textsc{FIGARO}, we perform a \ac{(H)DPGMM} to reconstruct $p(m^z_1|\Theta_i(\mathbf{Y}))$ with samples from all single-event \ac{PE} posterior \citep{Rinaldi:2021bhm}, and $\Theta_i$ is the hyperparameters of the \ac{(H)DPGMM} for a single reconstruction.
+To perform the reconstruction, we use a non-parametric reconstruction package \textsc{figaro}\footnote{\textsc{figaro} is publicly available at \url{https://github.com/sterinaldi/FIGARO} and via \texttt{pip}.} \citep{Rinaldi:2022kyg}.
+Using \textsc{figaro}, we perform a \ac{(H)DPGMM} to reconstruct $p(m^z_1|\Theta_i(\mathbf{Y}))$ with samples from all single-event \ac{PE} posterior \citep{Rinaldi:2021bhm}, and $\Theta_i$ are the hyperparameters of the \ac{(H)DPGMM} for a single reconstruction.
+\ste{Here it should be the other way around.
+We use (H)DPGMM as non-parametric model to reconstruct the observed detector-frame mass distribution -- you may want to extremely briefly summarise it here -- and then you mention the fact that the numerical implementation we use is \textsc{figaro}.}
 
@@
 The uncertainty of the reconstruction can then be quantified by the spread of the distributions.
-In Fig.~XXX \todo{Add figure}, we show an example of the reconstruction obtained with \textsc{FIGARO}.
+\todo{Mention the reconstruction holds only in the region with samples}
+In Fig.~XXX \todo{Add figure}, we show an example of the reconstruction obtained with \textsc{figaro}.
 
@@
 On the other hand, we need a source-frame population model $p(m_1|\Lambda)$ to represent the underlying population of the sources, where $\Lambda$ is the population model.
+\todo{Change to function of mass and redshift}
 We can either use a parametric model to only assume the form of the population distribution, or we can use a fixed population model to assume the exact population distribution.
@@
 By considering a continuous set of $\Omega$ and  $\Lambda$ of the population model, we can obtain a set of $p(m^z_1|\mathbf{\Lambda}, \mathbf{\Omega})$, which is a set of continuous detector-frame intrinsic population models for each combination of $\Omega_j$ and $\Lambda_j$, where $\mathbf{\Lambda} = \{\Lambda_j\}$ and $\mathbf{\Omega} = \{\Omega_j\}$.\footnote{In practice, we use a discrete set of $\Omega_j$ and $\Lambda_j$ with a small step size in their hyperparameters to approximate the continuous set.}
+\ste{This following sentence should be the very last sentence of this subsection, since it opens for the following PE description.} 
 After obtaining $p(m^z_1|\mathbf{\Lambda}, \mathbf{\Omega})$, we can perform the \ac{PE} to infer the best-fit $\Omega_j$ and $\Lambda_j$.
-\todo{Add selection function}
+\todo{PE->Inference of Cosmological Parameters}
+\todo{Add selection function} \ste{In the selection function part, you may also want to specify why we opted for reconstructing the observed distribution and not the intrinsic one with (H)DPGMM (namely the censoring of a specific portion of the parameter space).}
 In Fig.~XXX \todo{Add figure}, we show an example of the transformation.
@@
 For each reconstructed distribution $p(m^z_1|\Theta_i(\mathbf{Y}))$, we find the best-fit $\Lambda_j$ and $\Omega_j$ by minimizing $d_{JS}$.
+\todo{add the cutting of the low and high mass}
 This procedure gives us a set of best-fit $\Lambda_i$ and $\Omega_j$ for each $\Theta_i$.
@@
 In Fig.~XXX \todo{Add figure}, we show an example of the fitting procedure.
-
-\subsection{Minimization of $d_{JS}$}
-\label{sec:minimization}
+\todo{Add minimization process}
 
@@
 \label{sec:real-data-setup}
-Data retrived from \citet{LIGOScientific:2019lzm, KAGRA:2023pio}
+Data from \citet{LIGOScientific:2019lzm, KAGRA:2023pio}
 

### 2024-07-15 src/tex/ms.tex
[showyourwork] overleaf sync

--- 2024-05-16 src/tex/ms.tex
@@
 \usepackage{mathtools}
+\usepackage{array}
 
 \newcommand{\todo}[1]{\textcolor{red}{[TODO: #1]}}
+\newcommand{\ste}[1]{\textcolor{blue}{[Ste: #1]}}
 
@@
 \begin{acronym}
-    \acro{GW}{gravitational wave}
+    \acro{GW}{gravitational-wave}
     \acro{LVK}{LIGO-Virgo-KAGRA Collaboration}
@@
     \acro{EM}{electromagnetic}
-    \acro{(H)DPGMM}{hierarchy of Dirichlet process Gaussian mixture model}
+    \acro{(H)DPGMM}{hierarchy of Dirichlet process Gaussian mixture models}
     \acro{BBH}{binary black hole}
@@
 Some \ac{GW} sources can be associated with \ac{EM} counterparts, allowing for the measurement of the redshift, e.g., GW170817 \citep{LIGOScientific:2017adf, Guidorzi:2017ogy}.
-However, most \ac{GW} sources do not have \ac{EM} counterparts. \todo{Add galaxy catalog method}
-In this case, one can use population statistics to infer $H_0$ by marginalizing the redshift distribution of the sources.
-This method is known as the spectral siren \citep{You:2020wju, Mastrogiovanni:2021wsd, LIGOScientific:2021aug, Ezquiaga:2022zkx}.
+However, most \ac{GW} sources do not have \ac{EM} counterparts.
+In this case, one can assume \ac{GW} sources are in galaxies and associate the redshift of the galaxy to the \ac{GW} source. \citep{Schutz:1986gp, DelPozzo:2011vcw, Gray:2019ksv, Gray:2023wgj}
+Another alternative method is to use population statistics to infer $H_0$ by marginalizing the redshift distribution of the sources.
+This method is known as the spectral siren method\citep{You:2020wju, Mastrogiovanni:2021wsd, LIGOScientific:2021aug, Ezquiaga:2022zkx}.
 
 % Previous work
-The fundamental idea of the spectral siren method is to assume a population model for the mass distribution of the sources and marginalize over the redshift distribution to infer the cosmological parameters.
-Many studies have explored different variations of the spectral siren method.
+The fundamental idea of the spectral siren method is to infer cosmology by comparing the observed mass population with the intrinsic mass population since the observed mass population is redshifted due to the expansion of the Universe.
+Assumptions about the intrinsic population model and cosmological model are required to perform the comparison.
+Previous studies often infer either the intrinsic population model or the cosmological model, but not both simultaneously, due to the expensive computational cost.
 \todo{Summarize previous work and add references}
+However, the intrinsic population model and the cosmological model are degenerate in general, and it is necessary to infer both simultaneously.
 
-% What's new?
-In this study, we explore an alternative method based on the spectral siren framework.
-We propose two major modifications to the standard spectral siren method.
-First, instead of transforming the individual posterior samples from the detector frame to the source frame during the \ac{PE} process, we transform the population model to the detector frame before performing \ac{PE}.
-Second, instead of computing the likelihood of the observed data given the population model, we reconstruct the detector-frame observed population directly from the data using a non-parametric method.
+% Non-parametric methods
+To address this issue, we study an alternative method making use of non-parametric methods.
+Non-parametric methods allow us to reconstruct arbitrary probability densities without making assumptions about the form of the distribution.
+In our context, we can use non-parametric methods to reconstruct the observed mass population distribution directly from the data \todo{better wording?}.
+This provides an intermediate observation-driven result that encodes both the intrinsic population information and the cosmological information.
+By separating the reconstruction of the observed population from the inference of the intrinsic population and cosmological model, we can explore different models easily.
+Furthermore, the intermediate result represents the information contained in the observed data without any assumptions about the population model, cosmological model, or selection function, which allows us to study the features of the observed population directly.
 
-With these modifications, we aim to separate the spectral siren method into two distinct parts.
-The first part is the reconstruction of the observed population from the data, which does not require any assumptions about the population model, and the intermediate result is completely data-driven.
-Other assumptions including the population model, cosmological model and selection function are then incorporated in the second half.
-This separation allows us to explore the impact of different assumptions easily.
-The intermediate result of the first part can be reused as long as the data does not change, reducing the computational cost of the method.
-\todo{Add references}
+Some other studies have made use of non-parametric methods with the spectral siren framework.
+\todo{Summarize previous work and add references}
+\todo{Compare with our work}
+
+\todo{More events in the future}
 
@@
 In Sec.~\ref{sec:method}, we describe our modified method and the testing procedure.
-In Sec.~\ref{sec:results}, we present the results of our method.
-In Sec.~\ref{sec:discussion}, we compare our result with other methods and discuss future directions.
+In Sec.~\ref{sec:simulation}, we present the analysis setup and results of the simulation study.
+In Sec.~\ref{sec:real_data}, we apply our method to the real data from the \ac{LVK}.
+In Sec.~\ref{sec:discussion}, we compare our results with other methods and discuss future directions.
 Finally, we conclude in Sec.~\ref{sec:conclusion} with a summary.
@@
 \label{sec:method}
+In this section, we present the framework we developed to infer the cosmological parameters leveraging the detector-frame mass distribution.
+In the remainder of this Section, we will make use of the notation summarized in Table~\ref{tab:notation}, mutuated from \citet{Rinaldi:2021bhm, Rinaldi:2022kyg}.
+\todo{Renew notations}
+
+\begin{table}
+    \label{tab:notation}
+    \caption{Notation}
+    \begin{ruledtabular}
+        \begin{tabular}{c>{\raggedright\arraybackslash}p{0.8\linewidth}}
+            Notation & Description \\
+            \hline
+            $m_1$ & Primary mass of the \ac{GW} sources \\
+            $z$ & Redshift of the \ac{GW} sources \\
+            $d_L$ & Luminosity distance of the \ac{GW} sources \\
+            $m^z_1$ & Primary mass of the \ac{BBH} sources in the detector frame \\
+            $\mathbf{Y}$ & Set of posterior samples of all \ac{GW} events \\
+            $\mathbf{\Theta}$ & Set of hyperparameters of the \ac{(H)DPGMM} \\
+            $p(m^z_1|\mathbf{\Theta})$ & Detector-frame observed mass population distribution reconstructed with \ac{(H)DPGMM} \\
+            $\Lambda$ & Source-frame population model \\
+            $\Omega$ & Cosmological model \\
+            $p(m_1|\Lambda)$ & Source-frame mass population distribution assuming the population model $\Lambda$ \\
+            $p(m^z_1|\Lambda, \Omega)$ & Detector-frame mass population distribution assuming the population model $\Lambda$ and cosmological model $\Omega$ \\
+        \end{tabular}
+        \end{ruledtabular}
+\end{table}
+
+Specifically, we propose two major modifications to the standard spectral siren method.
+First, instead of transforming the individual posterior samples from the detector frame to the source frame during the \ac{PE} process, we transform the population model to the detector frame before performing \ac{PE}.
+Second, instead of computing the likelihood of the observed data given the population model, we reconstruct the detector-frame observed population directly from the data using a non-parametric method.
+
+With these modifications, we separate the spectral siren method into two distinct parts.
+The first part is the reconstruction of the observed population from the data, which does not require any assumptions about the population model, and the intermediate result is completely data-driven.
+We make use of a \ac{(H)DPGMM} developed in \citet{Rinaldi:2021bhm} as a non-parametric model to reconstruct the observed population distribution.
+Other assumptions including the population model, cosmological model and selection function are then incorporated in the second half.
+This separation allows us to explore the impact of different assumptions easily.
+The intermediate result of the first part can be reused as long as the data does not change, reducing the computational cost of the method.
 
@@
 we choose to consider the primary mass $m_1$ of the \ac{BBH} sources, in principle, we can consider more parameters at once.
-From the posterior samples obtained from the \ac{PE} of each \ac{GW} event, we reconstruct the detector-frame observed mass distribution $p(m^z_1|\mathbf{Y})$, where $\mathbf{Y}$ is the set of posterior samples of all \ac{GW} events.
+From the posterior samples obtained from the \ac{PE} of each \ac{GW} event, we reconstruct the detector-frame observed mass distribution.
 
-To perform the reconstruction, we use a non-parametric reconstruction package \textsc{FIGARO} \citep{Rinaldi:2022kyg}.
-Using \textsc{FIGARO}, we perform a \ac{(H)DPGMM} to reconstruct $p(m^z_1|\Theta_i(\mathbf{Y}))$ with samples from all single-event \ac{PE} posterior \citep{Rinaldi:2021bhm}, and $\Theta_i$ is the hyperparameters of the \ac{(H)DPGMM} for a single reconstruction.
-
-Performing the reconstruction once will give a distribution $p(m^z_1|\Theta_i(\mathbf{Y}))$ that is consistent with $\mathbf{Y}$ with a given set of hyperparameters $\Theta_i$.
-To account for the uncertainty in the reconstruction, we need to perform the reconstruction multiple times to obtain a set of distributions $p(m^z_1|\mathbf{\Theta}(\mathbf{Y}))$, where $\mathbf{\Theta}$ is a set of $\Theta_i$.
+We use a \ac{(H)DPGMM} developed in \citet{Rinaldi:2021bhm} as a non-parametric model to reconstruct the detector-frame observed mass population distribution $p(m^z_1|\mathbf{\Theta})$.
+The \ac{(H)DPGMM} is a hierarchy of Gaussian mixture models with a Dirichlet process prior, which allows the model to be flexible to reconstruct arbitrary population distribution from posterior samples.
+To perform the reconstruction, we use the package \textsc{figaro}\footnote{\textsc{figaro} is publicly available at \url{https://github.com/sterinaldi/FIGARO} and via \texttt{pip}.} \citep{Rinaldi:2024eep}.
+Performing the reconstruction once will give a distribution $p(m^z_1|\Theta_i)$ that is consistent with $\mathbf{Y}$ with a given set of hyperparameters $\Theta_i$.
+To account for the uncertainty in the reconstruction, we need to perform the reconstruction multiple times to obtain a set of distributions $p(m^z_1|\mathbf{\Theta})$, where $\mathbf{\Theta}$ is a set of $\Theta_i$.
 The uncertainty of the reconstruction can then be quantified by the spread of the distributions.
-In Fig.~XXX \todo{Add figure}, we show an example of the reconstruction obtained with \textsc{FIGARO}.
+It is important to note that the reconstruction is only valid in the region where the samples are present.
+In Fig.~XXX \todo{Add figure}, we show an example of the reconstruction obtained with \textsc{figaro}.
 
@@
 
-On the other hand, we need a source-frame population model $p(m_1|\Lambda)$ to represent the underlying population of the sources, where $\Lambda$ is the population model.
-We can either use a parametric model to only assume the form of the population distribution, or we can use a fixed population model to assume the exact population distribution.
-With $p(m_1|\Lambda)$, we need to transform it to the detector frame with a given cosmological model to obtain the detector-frame population model $p(m^z_1|\Lambda, \Omega)$, where $\Omega$ is the cosmological model.
-With different $\Omega$, $p(m_1|\Lambda)$ will be transformed differently, and the resulting $p(m^z_1|\Lambda, \Omega)$ will be different.
-By considering a continuous set of $\Omega$ and  $\Lambda$ of the population model, we can obtain a set of $p(m^z_1|\mathbf{\Lambda}, \mathbf{\Omega})$, which is a set of continuous detector-frame intrinsic population models for each combination of $\Omega_j$ and $\Lambda_j$, where $\mathbf{\Lambda} = \{\Lambda_j\}$ and $\mathbf{\Omega} = \{\Omega_j\}$.\footnote{In practice, we use a discrete set of $\Omega_j$ and $\Lambda_j$ with a small step size in their hyperparameters to approximate the continuous set.}
-After obtaining $p(m^z_1|\mathbf{\Lambda}, \mathbf{\Omega})$, we can perform the \ac{PE} to infer the best-fit $\Omega_j$ and $\Lambda_j$.
-\todo{Add selection function}
+On the other hand, we assume a population model $\Lambda$ to obtain a source-frame mass population $p(m_1|\Lambda)$, which represents the underlying population of the sources.
+Note that, either a parametric model can be used to only assume the form of the population distribution, or a fixed population model can be used to assume an exact population distribution.
+We then transform $p(m_1|\Lambda)$ to the detector frame with a given cosmological model $\Omega$ to obtain the detector-frame mass population $p(m^z_1|\Lambda, \Omega)$ by
+\begin{widetext}
+    \begin{equation}
+        \label{eq:transformation}
+        \begin{split}
+            &p(m^z_1|\Lambda, \Omega, \mathrm{det}) \\
+            &= \int p(m^z_1|m_1, z, \Lambda, \Omega, \mathrm{det}) p(m_1, z|\Lambda, \Omega, \mathrm{det}) \mathrm{d}m_1 \mathrm{d}z \\
+            &= \int \frac{p(\mathrm{det}|m^z_1, m_1, z, \Omega)p(m^z_1|m_1, z, \Omega)}{p(\mathrm{det}|m_1, z, \Omega)} \frac{p(\mathrm{det}|m_1, z, \Lambda, \Omega)p(m_1, z|\Lambda, \Omega)}{p(\mathrm{det}|\Lambda, \Omega)} \mathrm{d}m_1 \mathrm{d}z \\
+            &= \frac{1}{p(\mathrm{det}|\Lambda, \Omega)}\int \frac{p(\mathrm{det}|m_1, z, \Omega)p(m^z_1|m_1, z)p(\mathrm{det}|m_1, z, \Omega)p(m_1, z|\Lambda)}{p(\mathrm{det}|m_1, z, \Omega)}  \mathrm{d}m_1 \mathrm{d}z \\
+            &\propto \int \delta(m^z_1-m_1(1+z))p(\mathrm{det}|m_1, z, \Omega)p(m_1|\Lambda)p(z|\Lambda) \mathrm{d}m_1 \mathrm{d}z \\
+        \end{split}
+    \end{equation}
+\end{widetext}
+where we assume the source-frame mass distribution does not evolve with redshift, and $\mathrm{det}$ represents whether the event is detectable or not.
+We include the selection effect in Eq.~\eqref{eq:transformation}, which gives us $p(m^z_1|\Lambda, \Omega, \mathrm{det})$ in the final expression.
+With different $\Omega$, the resulting $p(m^z_1|\Lambda, \Omega, \mathrm{det})$ will be different.
+By considering a continuous set of $\Omega$ and  $\Lambda$ of the population model, we can obtain a set of $p(m^z_1|\mathbf{\Lambda}, \mathbf{\Omega}, \mathrm{det})$, which is a set of continuous detector-frame observed population models for each combination of $\Omega_j$ and $\Lambda_j$, where $\mathbf{\Lambda} = \{\Lambda_j\}$ and $\mathbf{\Omega} = \{\Omega_j\}$.\footnote{In practice, we use a discrete set of $\Omega_j$ and $\Lambda_j$ with a small step size in their hyperparameters to approximate the continuous set.}
+\ste{This following sentence should be the very last sentence of this subsection, since it opens for the following PE description.} 
+After obtaining $p(m^z_1|\mathbf{\Lambda}, \mathbf{\Omega}, \mathrm{det})$, we can perform the \ac{PE} to infer the optimized $\Omega_j$ and $\Lambda_j$.
+\todo{PE\textrightarrow Inference of Cosmological Parameters}
+\ste{In the selection function part, you may also want to specify why we opted for reconstructing the observed distribution and not the intrinsic one with (H)DPGMM (namely the censoring of a specific portion of the parameter space).}
 In Fig.~XXX \todo{Add figure}, we show an example of the transformation.
@@
 The \ac{PE} is performed by comparing $p(m^z_1|\Theta_i(\mathbf{Y}))$ with $p(m^z_1|\Lambda_j, \Omega_j)$.
-To compare the two populations, we use the Jensen–Shannon distance $d_{JS}$ as the distance measure.
+To compare the two populations, we use the Jensen-Shannon distance $d_{JS}$ as the distance measure.
 $d_{JS}$ is calculated as
@@
 \end{equation}
-where $p$ and $q$ are the two distributions to be compared, $m = (p + q) / 2$, and $D_{KL}$ is the Kullback–Leibler divergence.
+where $p$ and $q$ are the two distributions to be compared, $m = (p + q) / 2$, and $D_{KL}$ is the Kullback-Leibler divergence.
 $d_{JS}$ is a symmetric and smoothed version of $D_{KL}$, and it is bounded between 0 and 1.
 
-For each reconstructed distribution $p(m^z_1|\Theta_i(\mathbf{Y}))$, we find the best-fit $\Lambda_j$ and $\Omega_j$ by minimizing $d_{JS}$.
-This procedure gives us a set of best-fit $\Lambda_i$ and $\Omega_j$ for each $\Theta_i$.
-Note that the resulting distribution of the best-fit parameters is not a posterior distribution in the Bayesian sense, as we do not consider the likelihood of the observed data given the population model.
-Instead, it is a distribution of the best-fit parameters that are consistent with the observed data, distributed according to the uncertainty in the reconstruction of the population.
+For each reconstructed distribution $p(m^z_1|\Theta_i(\mathbf{Y}))$, we find the optimized $\Lambda_j$ and $\Omega_j$ by minimizing $d_{JS}$.
+\todo{add the cutting of the low and high mass}
+For the minimization, we use the \textsc{scipy} \citep{2020SciPy-NMeth} implementation of the \todo{XXX} algorithm. \todo{Add reference for the algorithm}
+This procedure gives us a set of optimized $\Lambda_i$ and $\Omega_j$ for each $\Theta_i$.
+Note that the resulting distribution of the optimized parameters is not a posterior distribution in the Bayesian sense, as we do not consider the likelihood of the observed data given the population model.
+Instead, it is a distribution of the optimized parameters that are consistent with the observed data, distributed according to the uncertainty in the reconstruction of the population.
 That is, the uncertainty in the comparison of the two populations is not considered in the resulting distribution.
-By considering this optimization procedure only, we can estimate the best-fit parameters without performing computationally expensive likelihood calculations.
+By considering this optimization procedure only, we can estimate the optimized parameters without performing computationally expensive likelihood calculations.
 In Fig.~XXX \todo{Add figure}, we show an example of the fitting procedure.
 
-\subsection{Minimization of $d_{JS}$}
-\label{sec:minimization}
+The analysis code is publicly available.\footnote{\todo{Add link}}
 
-\section{Analysis}
-\label{sec:analysis}
-
-\subsection{Simulation}
-\label{sec:injection-setup}
-
-\subsection{Real data}
-\label{sec:real-data-setup}
-Data retrived from \citet{LIGOScientific:2019lzm, KAGRA:2023pio}
-
-\section{Results}
-\label{sec:results}
-
-\subsection{Simulation}
-\label{sec:injection-results}
+\section{Simulation}
+\label{sec:simulation}
 
@@
 
-\subsection{Real data}
-\label{sec:real-data-results}
+\section{Real data}
+\label{sec:real_data}
+
+Data from \citet{LIGOScientific:2019lzm, KAGRA:2023pio}
 
@@
 
+% PL+P issue
+    % Gregoire Pierra's paper
+
 % Future directions
@@
 \begin{acknowledgments}
-T.N.~ and S.R.~ acknowledge financial support from the German Excellence Strategy via the Heidelberg Cluster of Excellence (EXC 2181 - 390900948) STRUCTURES.
-S.R.~ acknowledges financial support from the European Research Council for the ERC Consolidator grant DEMOBLACK, under contract no. 770017.
+T.~N.~ and S.~R.~ acknowledge financial support from the German Excellence Strategy via the Heidelberg Cluster of Excellence (EXC 2181 - 390900948) STRUCTURES.
+S.~R.~ acknowledges financial support from the European Research Council for the ERC Consolidator grant DEMOBLACK, under contract no. 770017.
 

### 2024-07-15 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 Some other studies have made use of non-parametric methods with the spectral siren framework.
-\todo{Summarize previous work and add references}
-\todo{Compare with our work}
+\todo{Summarize previous work, add references and compare with our work}
 
-\todo{More events in the future}
+Considering the increasing number of \ac{GW} events and the improvement of the detectors, we expect that the spectral siren method will give us an independent and precise measurement of $H_0$ which is comparable to the previous measurements \citep{Planck:2018vyg, Riess:2021jrx}.
 
@@
             $m^z_1$ & Primary mass of the \ac{BBH} sources in the detector frame \\
-            $\mathbf{Y}$ & Set of posterior samples of all \ac{GW} events \\
+            $\mathbf{Y}$ & Set of posterior samples of the \ac{GW} events \\
             $\mathbf{\Theta}$ & Set of hyperparameters of the \ac{(H)DPGMM} \\
@@
             $p(m^z_1|\Lambda, \Omega)$ & Detector-frame mass population distribution assuming the population model $\Lambda$ and cosmological model $\Omega$ \\
+            $\mathrm{det}$ & detectability of the \ac{GW} events \\
         \end{tabular}
@@
             &= \int p(m^z_1|m_1, z, \Lambda, \Omega, \mathrm{det}) p(m_1, z|\Lambda, \Omega, \mathrm{det}) \mathrm{d}m_1 \mathrm{d}z \\
-            &= \int \frac{p(\mathrm{det}|m^z_1, m_1, z, \Omega)p(m^z_1|m_1, z, \Omega)}{p(\mathrm{det}|m_1, z, \Omega)} \frac{p(\mathrm{det}|m_1, z, \Lambda, \Omega)p(m_1, z|\Lambda, \Omega)}{p(\mathrm{det}|\Lambda, \Omega)} \mathrm{d}m_1 \mathrm{d}z \\
-            &= \frac{1}{p(\mathrm{det}|\Lambda, \Omega)}\int \frac{p(\mathrm{det}|m_1, z, \Omega)p(m^z_1|m_1, z)p(\mathrm{det}|m_1, z, \Omega)p(m_1, z|\Lambda)}{p(\mathrm{det}|m_1, z, \Omega)}  \mathrm{d}m_1 \mathrm{d}z \\
-            &\propto \int \delta(m^z_1-m_1(1+z))p(\mathrm{det}|m_1, z, \Omega)p(m_1|\Lambda)p(z|\Lambda) \mathrm{d}m_1 \mathrm{d}z \\
+            &= \int p(m^z_1|m_1, z) \frac{p(\mathrm{det}|m_1, z, \Lambda, \Omega)p(m_1, z|\Lambda, \Omega)}{p(\mathrm{det}|\Lambda, \Omega)} \mathrm{d}m_1 \mathrm{d}z \\
+            &= \frac{1}{p(\mathrm{det}|\Lambda, \Omega)}\int \delta(m^z_1-m_1(1+z))p(\mathrm{det}|m_1, z, \Omega)p(m_1|\Lambda)p(z|\Lambda) \mathrm{d}m_1 \mathrm{d}z \\
+            &\propto \int p\left(\mathrm{det}\middle|\frac{m^z_1}{1+z},z,\Omega\right)p\left(\frac{m^z_1}{1+z}\middle|\Lambda\right)p(z|\Lambda) \mathrm{d}z\,,
         \end{split}
@@
 where we assume the source-frame mass distribution does not evolve with redshift, and $\mathrm{det}$ represents whether the event is detectable or not.
-We include the selection effect in Eq.~\eqref{eq:transformation}, which gives us $p(m^z_1|\Lambda, \Omega, \mathrm{det})$ in the final expression.
+The selection effect is included in the derivation in eq.~\eqref{eq:transformation}, specifically, the first term in the final expression, which is the selection function.
+Note that the selection function in general is a function of the detector-frame mass $m^z_1$, luminosity distance $d_L$, and mass ratio $q$.
+In this work, we marginalized over $q$ to study the primary mass distribution only.
+Due to the censoring of the low and high mass events, we consider the selection function in the transformation instead of the reconstruction.
+
 With different $\Omega$, the resulting $p(m^z_1|\Lambda, \Omega, \mathrm{det})$ will be different.
-By considering a continuous set of $\Omega$ and  $\Lambda$ of the population model, we can obtain a set of $p(m^z_1|\mathbf{\Lambda}, \mathbf{\Omega}, \mathrm{det})$, which is a set of continuous detector-frame observed population models for each combination of $\Omega_j$ and $\Lambda_j$, where $\mathbf{\Lambda} = \{\Lambda_j\}$ and $\mathbf{\Omega} = \{\Omega_j\}$.\footnote{In practice, we use a discrete set of $\Omega_j$ and $\Lambda_j$ with a small step size in their hyperparameters to approximate the continuous set.}
-\ste{This following sentence should be the very last sentence of this subsection, since it opens for the following PE description.} 
-After obtaining $p(m^z_1|\mathbf{\Lambda}, \mathbf{\Omega}, \mathrm{det})$, we can perform the \ac{PE} to infer the optimized $\Omega_j$ and $\Lambda_j$.
-\todo{PE\textrightarrow Inference of Cosmological Parameters}
-\ste{In the selection function part, you may also want to specify why we opted for reconstructing the observed distribution and not the intrinsic one with (H)DPGMM (namely the censoring of a specific portion of the parameter space).}
+By considering a continuous set of $\Omega$ and  $\Lambda$ of the population model, we can obtain a set of $p(m^z_1|\mathbf{\Lambda}, \mathbf{\Omega}, \mathrm{det})$, which is a set of continuous detector-frame observed population models for each combination of $\Omega_j$ and $\Lambda_j$, where $\mathbf{\Lambda} = \{\Lambda_j\}$ and $\mathbf{\Omega} = \{\Omega_j\}$.
 In Fig.~XXX \todo{Add figure}, we show an example of the transformation.
@@
 Opposite to the first part of the method, this part includes the assumptions about the population and cosmological model, as well as the selection function.
+After obtaining $p(m^z_1|\mathbf{\Lambda}, \mathbf{\Omega}, \mathrm{det})$, we can perform the \ac{PE} to infer the optimized $\Omega_j$ and $\Lambda_j$.
 
@@
 
-With $p(m^z_1|\mathbf{\Theta}(\mathbf{Y}))$ and $p(m^z_1|\mathbf{\Lambda}, \mathbf{\Omega})$ obtained from Sec.~\ref{sec:reconstruction} and Sec.~\ref{sec:transformation} respectively, we can perform the \ac{PE} to infer the free parameters in the parametric models assumed in Sec.~\ref{sec:transformation}.
-The \ac{PE} is performed by comparing $p(m^z_1|\Theta_i(\mathbf{Y}))$ with $p(m^z_1|\Lambda_j, \Omega_j)$.
-To compare the two populations, we use the Jensen-Shannon distance $d_{JS}$ as the distance measure.
-$d_{JS}$ is calculated as
+With $p(m^z_1|\mathbf{\Theta})$ and $p(m^z_1|\mathbf{\Lambda}, \mathbf{\Omega}, \mathrm{det})$ obtained from Sec.~\ref{sec:reconstruction} and Sec.~\ref{sec:transformation} respectively, we can perform the \ac{PE} to infer the free parameters in the parametric models assumed in Sec.~\ref{sec:transformation}.
+The \ac{PE} is performed by comparing $p(m^z_1|\Theta_i)$ with $p(m^z_1|\Lambda_j, \Omega_j)$.
+To compare the two populations, we use the Jensen-Shannon distance $d_\mathrm{JS}$ as the distance measure.
+$d_\mathrm{JS}$ is calculated as
 \begin{equation}
-    d_{JS}(p, q) = \sqrt{\frac{D_{KL}(p||m) + D_{KL}(q||m)}{2}},
+    d_\mathrm{JS}(p, q) = \sqrt{\frac{D_\mathrm{KL}(p||m) + D_\mathrm{KL}(q||m)}{2}},
 \end{equation}
-where $p$ and $q$ are the two distributions to be compared, $m = (p + q) / 2$, and $D_{KL}$ is the Kullback-Leibler divergence.
-$d_{JS}$ is a symmetric and smoothed version of $D_{KL}$, and it is bounded between 0 and 1.
+where $p$ and $q$ are the two distributions to be compared, $m = (p + q) / 2$, and $D_\mathrm{KL}$ is the Kullback-Leibler divergence.
+$d_\mathrm{JS}$ is a symmetric and smoothed version of $D_\mathrm{KL}$, and it is bounded between 0 and 1.
 
-For each reconstructed distribution $p(m^z_1|\Theta_i(\mathbf{Y}))$, we find the optimized $\Lambda_j$ and $\Omega_j$ by minimizing $d_{JS}$.
-\todo{add the cutting of the low and high mass}
+For each reconstructed distribution $p(m^z_1|\Theta_i)$, we find the optimized $\Lambda_j$ and $\Omega_j$ by minimizing $d_\mathrm{JS}$.
 For the minimization, we use the \textsc{scipy} \citep{2020SciPy-NMeth} implementation of the \todo{XXX} algorithm. \todo{Add reference for the algorithm}
 This procedure gives us a set of optimized $\Lambda_i$ and $\Omega_j$ for each $\Theta_i$.
-Note that the resulting distribution of the optimized parameters is not a posterior distribution in the Bayesian sense, as we do not consider the likelihood of the observed data given the population model.
+Note that we first calculate the median of the detector-frame primary mass posterior samples of each \ac{GW} event, then we calculate $d_\mathrm{JS}$ in the range from the minimum to the maximum of the medians.
+This is because the reconstruction is only valid in the region where observations are present.
+
+The resulting distribution of the optimized parameters is not a posterior distribution in the Bayesian sense, as we do not consider the likelihood of the observed data given the population model.
 Instead, it is a distribution of the optimized parameters that are consistent with the observed data, distributed according to the uncertainty in the reconstruction of the population.

### 2024-07-15 src/scripts/analysis/generate_selection_function.py
Add some analysis code

@@ -1,0 +1,48 @@
+import numpy as np
+import dill
+import paths
+from tqdm import tqdm
+from scipy.interpolate import RegularGridInterpolator
+from cbc_pdet.o123_class_found_inj_general import Found_injections # from https://github.com/sterinaldi/cbc_pdet
+
+# Selection function
+run_fit = 'o3'
+run_dataset = 'o3'
+dmid_fun = 'Dmid_mchirp_expansion_noa30'
+emax_fun = 'emax_exp'
+alpha_vary = None
+
+selfunc = Found_injections(dmid_fun, emax_fun, alpha_vary)
+selfunc.load_inj_set(run_dataset)
+selfunc.get_opt_params(run_fit)
+
+n_pts  = np.array([200,200,200])
+
+# Bounds
+dl_bds = [0.01,15000.]
+m1_bds = [0.,500.]
+q_bds  = [0.05, 1.]
+bounds = np.array([m1_bds, q_bds, dl_bds])
+
+# Arrays
+m1    = np.linspace(m1_bds[0], m1_bds[1], n_pts[0])
+q     = np.linspace(q_bds[0], q_bds[1], n_pts[1])
+dl    = np.linspace(dl_bds[0], dl_bds[1], n_pts[2])
+dn_m1 = np.prod(n_pts[1:])
+dn_q  = np.prod(n_pts[2:])
+
+# For loop
+grid    = np.zeros(shape = (np.prod(n_pts), 3))
+grid_dl = np.zeros(shape = (np.prod(n_pts), 3))
+for i, m1i in tqdm(enumerate(m1), desc = 'Grid', total = n_pts[0]):
+    for j, qi in enumerate(q):
+        for k, dli in enumerate(dl):
+            grid[i*dn_m1 + j*dn_q + k] = [m1i, qi, dli]
+
+# Selection function in detector-frame mass
+pdet = np.trapz(np.nan_to_num(selfunc.run_pdet(grid[:,2], grid[:,0], grid[:,1]*grid[:,0], 'o3').reshape(n_pts), nan = 0.), q, axis = 1)
+
+selfunc_interp = RegularGridInterpolator((m1, dl), pdet, bounds_error = False, fill_value = 0.)
+
+with open(paths.data / 'selection_function.pkl', 'wb') as f:
+    dill.dump(selfunc_interp, f)

### 2024-07-15 src/scripts/analysis/inference_H0.py

@@ -1,0 +1,71 @@
+import numpy as np
+from figaro.cosmology import CosmologicalParameters
+from scipy.spatial.distance import jensenshannon as scipy_jsd
+from figaro.load import load_density
+import os
+import sys
+from tqdm import tqdm
+import dill
+import paths
+
+# Mass distribution
+from population_models.mass import plpeak # from from https://github.com/sterinaldi/cbc_pdet
+
+# Redshift distribution
+def p_z(z, H0):
+    return CosmologicalParameters(H0/100., 0.315, 0.685, -1., 0., 0.).ComovingVolumeElement(z)/(1+z)
+
+label = sys.argv[1]
+outdir = os.path.dirname(os.path.realpath(__file__)) + "/" + label
+outdir = paths.data / outdir
+
+print("Preparing model pdfs...")
+try:
+    mz, H0, z, m, model_pdf = np.load(outdir+"/../real_model_pdf.npz").values()
+except:
+    mz = np.linspace(1,200,900)
+    H0 = np.linspace(5,150,1000)
+    z = np.linspace(0.001,2,800)
+    m = np.einsum("i, j -> ij", mz, np.reciprocal(1+z)) # shape = (len(mz), len(z))
+
+    # model mz pdf for each H0
+    model_pdf = np.einsum("ij, kj -> ijk", plpeak(m), [p_z(z, i) for i in H0]) # shape = (len(mz), len(z), len(H0))
+
+    # Load selection function
+    with open('selection_function.pkl', 'rb') as f: # selection_function.pkl is generated by generate_selection_function.py
+        selfunc_interp = dill.load(f)
+    def selection_function(x):
+        return selfunc_interp(x)
+    
+    grid = [np.transpose(np.meshgrid(mz, CosmologicalParameters(i/100., 0.315, 0.685, -1., 0., 0.).LuminosityDistance(z))) for i in H0] # shape = (len(H0), len(mz), len(z), 2)
+    SE_grid = np.array([selection_function(grid[i]) for i in tqdm(range(len(H0)), desc = 'SE grid')]) # shape = (len(H0), len(mz), len(z))
+    model_pdf = np.einsum("ijk, kij -> ijk", model_pdf, SE_grid) # shape = (len(mz), len(z), len(H0))
+
+    model_pdf = np.trapz(model_pdf, z, axis=1) # shape = (len(mz), len(H0))
+
[…]

### 2024-07-15 src/scripts/analysis/simulation.py

@@ -1,0 +1,51 @@
+import numpy as np
+from figaro.utils import rejection_sampler
+from figaro.cosmology import dVdz_approx_planck18, Planck18 as Omega
+import os
+import dill
+import paths
+
+np.random.default_rng(12345)
+
+# Mass distribution
+from population_models.mass import plpeak # from https://github.com/thomasckng/pop_models_lvk
+
+# Redshift distribution
+def p_z(z):
+    return dVdz_approx_planck18(z)/(1+z)
+
+# Generate true detector frame mass samples
+def generate_truth(n_single_event_draws):
+    M = rejection_sampler(n_single_event_draws, plpeak, [0,200])
+    z = rejection_sampler(n_single_event_draws, p_z, [0, 2])
+    return np.array([M * (1 + z), z]).T
+
+# Generate detector frame mass posterior samples
+def generate_mz_posterior_samples(truth, sigma = 0.03, n_samples = 1000):
+    log_obs_mean = np.log(truth) + np.random.normal(0, sigma, truth.shape) # shape = (n_events)
+    return np.exp(np.random.normal(log_obs_mean, sigma, (n_samples, len(truth)))).T # shape = (n_events, n_samples)
+
+# Load selection function
+with open('selection_function.pkl', 'rb') as f: # selection_function.pkl is generated by generate_selection_function.py
+    selfunc_interp = dill.load(f)
+def selection_function(x):
+    return selfunc_interp(x)
+
+n_single_event_draws = 1_000_000
+outdir = 'hierarchical_single_event_test'
+if not os.path.exists(paths.data / outdir):
+    os.makedirs(paths.data / outdir)
+outdir = paths.data / outdir
+
+samples = generate_truth(n_single_event_draws) # shape = (n_events, n_params)
+samples_single_event = samples[np.random.uniform(0,1,samples.shape[0]) <= selection_function(samples[:,0],Omega.LuminosityDistance(samples[:,1]))] # shape = (n_events, n_params)
+posterior = generate_mz_posterior_samples(samples_single_event[:,0]) # shape = (n_events, n_samples)
+
+np.savetxt(outdir+'/true_samples.txt', samples[:,0])
+np.savetxt(outdir+'/obs_samples.txt', samples_single_event[:,0])
+np.savetxt(outdir+'/jsd_bounds.txt', [np.median(posterior, axis=1).min(), np.median(posterior, axis=1).max()])
+
[…]

### 2024-07-15 src/scripts/paths.py

@@ -20,6 +20,12 @@ static = src / "static"
 # Absolute path to the `src/scripts` folder (contains figure/pipeline scripts)
 scripts = src / "scripts"
 
+# Absolute path to the `src/scripts/plotting` folder (contains plotting scripts)
+plotting = scripts / "plotting"
+
+# Absolute path to the `src/scripts/analysis` folder (contains analysis scripts)
+analysis = scripts / "analysis"
+
 # Absolute path to the `src/tex` folder (contains the manuscript)
 tex = src / "tex"
 

### 2024-07-16 src/scripts/infer_H0.py
Modify analysis scripts

@@ -20,15 +20,16 @@ outdir = os.path.dirname(os.path.realpath(__file__)) + "/" + label
 outdir = paths.data / outdir
 
 print("Preparing model pdfs...")
+grid_label = "real_model_pdf"
 try:
-    mz, H0, z, m, model_pdf = np.load(outdir+"/../real_model_pdf.npz").values()
+    mz, H0, z, m, model_pdf = np.load(outdir+"/../"+grid_label+".npz").values()
 except:
     mz = np.linspace(1,200,900)
     H0 = np.linspace(5,150,1000)
     z = np.linspace(0.001,2,800)
     m = np.einsum("i, j -> ij", mz, np.reciprocal(1+z)) # shape = (len(mz), len(z))
 
-    # model mz pdf for each H0
+    # Calculate source-frame population model pdf for each H0
     model_pdf = np.einsum("ij, kj -> ijk", plpeak(m), [p_z(z, i) for i in H0]) # shape = (len(mz), len(z), len(H0))
 
     # Load selection function
@@ -43,12 +44,12 @@ except:
 
     model_pdf = np.trapz(model_pdf, z, axis=1) # shape = (len(mz), len(H0))
 
-    np.savez(outdir+"/../real_model_pdf.npz", mz=mz, H0=H0, z=z, m=m, model_pdf=model_pdf)
+    np.savez(outdir / "../"+grid_label+".npz", mz=mz, H0=H0, z=z, m=m, model_pdf=model_pdf)
 
 print("Reading bounds and draws...")
-draws = load_density(outdir+"/draws/draws_observed_"+label+".json")
+draws = load_density(outdir / "draws/draws_observed_"+label+".json")
 
-bounds = np.loadtxt(outdir+"/jsd_bounds.txt")
+bounds = np.loadtxt(outdir / "jsd_bounds.txt")
 
 print("Preparing H0 inference...")
 # Mask out mz where there is no sample
@@ -58,14 +59,14 @@ model_pdf_short = model_pdf[_mask]
 
 figaro_pdf = np.array([draw.pdf(mz_short) for draw in draws])# shape (n_draws, len(mz_short))
 
-print("Infering H0...")
+print("Inferring H0...")
 # Compute JSD between (reconstructed observed distributions for each DPGMM draw) and (model mz distributions for each H0)
[…]

### 2024-07-16 src/scripts/paths.py

@@ -20,12 +20,6 @@ static = src / "static"
 # Absolute path to the `src/scripts` folder (contains figure/pipeline scripts)
 scripts = src / "scripts"
 
-# Absolute path to the `src/scripts/plotting` folder (contains plotting scripts)
-plotting = scripts / "plotting"
-
-# Absolute path to the `src/scripts/analysis` folder (contains analysis scripts)
-analysis = scripts / "analysis"
-
 # Absolute path to the `src/tex` folder (contains the manuscript)
 tex = src / "tex"
 

### 2024-07-16 src/scripts/simulate_posterior_samples.py

@@ -26,26 +26,27 @@ def generate_mz_posterior_samples(truth, sigma = 0.03, n_samples = 1000):
     return np.exp(np.random.normal(log_obs_mean, sigma, (n_samples, len(truth)))).T # shape = (n_events, n_samples)
 
 # Load selection function
-with open('selection_function.pkl', 'rb') as f: # selection_function.pkl is generated by generate_selection_function.py
+with open(paths.data/'selection_function.pkl', 'rb') as f: # selection_function.pkl is generated by generate_selection_function.py
     selfunc_interp = dill.load(f)
 def selection_function(x):
     return selfunc_interp(x)
 
-n_single_event_draws = 1_000_000
-outdir = 'hierarchical_single_event_test'
+n_single_event_draws = 100_000
+outdir = 'simulation'
 if not os.path.exists(paths.data / outdir):
     os.makedirs(paths.data / outdir)
 outdir = paths.data / outdir
 
 samples = generate_truth(n_single_event_draws) # shape = (n_events, n_params)
-samples_single_event = samples[np.random.uniform(0,1,samples.shape[0]) <= selection_function(samples[:,0],Omega.LuminosityDistance(samples[:,1]))] # shape = (n_events, n_params)
+samples[:,1] = Omega.LuminosityDistance(samples[:,1])
+samples_single_event = samples[np.random.uniform(0,1,samples.shape[0]) <= selection_function(samples)] # shape = (n_events, n_params)
 posterior = generate_mz_posterior_samples(samples_single_event[:,0]) # shape = (n_events, n_samples)
 
-np.savetxt(outdir+'/true_samples.txt', samples[:,0])
-np.savetxt(outdir+'/obs_samples.txt', samples_single_event[:,0])
-np.savetxt(outdir+'/jsd_bounds.txt', [np.median(posterior, axis=1).min(), np.median(posterior, axis=1).max()])
+np.savetxt(outdir / 'true_samples.txt', samples[:,0])
+np.savetxt(outdir / 'obs_samples.txt', samples_single_event[:,0])
+np.savetxt(outdir / 'jsd_bounds.txt', [np.median(posterior, axis=1).min(), np.median(posterior, axis=1).max()])
 
-if not os.path.exists(outdir+'/data'):
-    os.makedirs(outdir+'/data')
+if not os.path.exists(outdir / 'data'):
+    os.makedirs(outdir / 'data')
[…]

### 2024-07-16 src/scripts/infer_H0.py
Add plotting scripts

@@ -2,7 +2,6 @@ import numpy as np
 from figaro.cosmology import CosmologicalParameters
 from scipy.spatial.distance import jensenshannon as scipy_jsd
 from figaro.load import load_density
-import os
 import sys
 from tqdm import tqdm
 import dill
@@ -16,13 +15,12 @@ def p_z(z, H0):
     return CosmologicalParameters(H0/100., 0.315, 0.685, -1., 0., 0.).ComovingVolumeElement(z)/(1+z)
 
 label = sys.argv[1]
-outdir = os.path.dirname(os.path.realpath(__file__)) + "/" + label
-outdir = paths.data / outdir
+outdir = paths.data / label
 
 print("Preparing model pdfs...")
-grid_label = "real_model_pdf"
+grid_label = "grid"
 try:
-    mz, H0, z, m, model_pdf = np.load(outdir+"/../"+grid_label+".npz").values()
+    mz, H0, z, m, model_pdf = np.load(outdir / f"/../{grid_label}.npz").values()
 except:
     mz = np.linspace(1,200,900)
     H0 = np.linspace(5,150,1000)
@@ -39,15 +37,15 @@ except:
         return selfunc_interp(x)
     
     grid = [np.transpose(np.meshgrid(mz, CosmologicalParameters(i/100., 0.315, 0.685, -1., 0., 0.).LuminosityDistance(z))) for i in H0] # shape = (len(H0), len(mz), len(z), 2)
-    SE_grid = np.array([selection_function(grid[i]) for i in tqdm(range(len(H0)), desc = 'SE grid')]) # shape = (len(H0), len(mz), len(z))
+    SE_grid = np.array([selection_function(grid[i]) for i in tqdm(range(len(H0)), desc = 'selection function grid')]) # shape = (len(H0), len(mz), len(z))
     model_pdf = np.einsum("ijk, kij -> ijk", model_pdf, SE_grid) # shape = (len(mz), len(z), len(H0))
 
     model_pdf = np.trapz(model_pdf, z, axis=1) # shape = (len(mz), len(H0))
 
-    np.savez(outdir / "../"+grid_label+".npz", mz=mz, H0=H0, z=z, m=m, model_pdf=model_pdf)
+    np.savez(outdir / f"../{grid_label}.npz", mz=mz, H0=H0, z=z, m=m, model_pdf=model_pdf)
 
 print("Reading bounds and draws...")
[…]

### 2024-07-16 src/scripts/plot_simulation_comparison.py

@@ -1,0 +1,28 @@
+from figaro.load import load_density
+from figaro.plot import plot_median_cr
+from figaro import plot_settings
+import numpy as np
+import paths
+
+label = 'simulation'
+outdir = paths.data / label
+
+draws = load_density(outdir / f'draws/draws_observed_{label}.json')
+
+mz, H0, z, m, model_pdf = np.load(outdir / '../grid.npz').values()
+
+bounds = np.loadtxt(outdir / "jsd_bounds.txt")
+
+fig = plot_median_cr(draws, save=True, show=False)
+ax = fig.axes[0]
+
+colors = ['tab:green', 'tab:red', 'tab:orange']
+for i, c in zip([np.argmin(abs(H0-40)), np.argmin(abs(H0-70)), np.argmin(abs(H0-100))], colors):
+    ax.plot(mz, model_pdf[:,i]/np.trapz(model_pdf[:,i], mz), label=f"$H_0={H0[i]:.0f}$", c=c)
+ax.legend()
+ax.set_xlabel('$m^z_1\ [\mathrm{M}_\odot]$')
+ax.set_ylabel('Density')
+ax.set_xlim(bounds[0], bounds[1])
+ax.set_ylim(0)
+
+fig.savefig(outdir+"/comparison.pdf", bbox_inches='tight')

### 2024-07-16 src/scripts/plot_simulation_reconstruction.py

@@ -1,0 +1,21 @@
+from figaro.load import load_density
+from figaro.plot import plot_median_cr
+from figaro import plot_settings
+import numpy as np
+import paths
+
+label = 'simulation'
+outdir = paths.data / 'simulation'
+
+draws = load_density(outdir / f'draws/draws_observed_{label}.json')
+obs_samples = np.loadtxt(outdir / 'obs_samples.txt')
+
+fig = plot_median_cr(draws, samples=obs_samples, save=True, show=False)
+ax = fig.axes[0]
+ax.hist(obs_samples, bins = int(np.sqrt(len(obs_samples))), histtype = 'step', density = True, label = '$\mathrm{median}(Y_t)$')
+ax.set_xlim(0,120)
+ax.set_ylim(0)
+ax.xlabel('$m^z_1\ [\mathrm{M}_\odot]$')
+ax.ylabel('Density')
+ax.legend()
+fig.savefig(paths.figures / 'simulation_reconstruction.pdf', bbox_inches='tight')

### 2024-07-16 src/scripts/plot_simulation_transformation.py

@@ -1,0 +1,31 @@
+from matplotlib import pyplot as plt
+import numpy as np
+import paths
+
+label = 'simulation'
+outdir = paths.data / label
+
+mz, H0, z, m, model_pdf = np.load(outdir / '../grid.npz').values()
+
+# Mass distribution
+from population_models.mass import plpeak
+
+fig, (ax1, ax2) = plt.subplots(2)
+
+ax1.plot(m, plpeak(m)/np.trapz(plpeak(m), m), label = "PL+Peak", c = 'tab:blue')
+ax1.xlabel('$m_1\ [\mathrm{M}_\odot]$')
+ax1.ylabel('$p(m_1|\Lambda)$')
+ax1.legend()
+ax1.xlim(0,120)
+ax1.ylim(0)
+
+colors = ['tab:green', 'tab:red', 'tab:orange']
+for i, c in zip([np.argmin(abs(H0-40)), np.argmin(abs(H0-70)), np.argmin(abs(H0-100))], colors):
+    ax2.plot(mz, model_pdf[:,i]/np.trapz(model_pdf[:,i], mz), label=f"$H_0={H0[i]:.0f}$", c=c)
+ax2.legend()
+ax2.xlabel('$m^z_1\ [\mathrm{M}_\odot]$')
+ax2.ylabel('$p(m^z_1|\Lambda, \Omega(H_0), \mathrm{det})$')
+ax2.xlim(0,120)
+ax2.ylim(0)
+
+fig.savefig(outdir / "simulation_transformation.pdf", bbox_inches='tight')

### 2024-07-16 src/scripts/infer_H0.py
Modify plotting scripts

@@ -20,7 +20,7 @@ outdir = paths.data / label
 print("Preparing model pdfs...")
 grid_label = "grid"
 try:
-    mz, H0, z, m, model_pdf = np.load(outdir / f"/../{grid_label}.npz").values()
+    mz, H0, z, m, model_pdf = np.load(outdir / f"../{grid_label}.npz").values()
 except:
     mz = np.linspace(1,200,900)
     H0 = np.linspace(5,150,1000)
@@ -31,7 +31,7 @@ except:
     model_pdf = np.einsum("ij, kj -> ijk", plpeak(m), [p_z(z, i) for i in H0]) # shape = (len(mz), len(z), len(H0))
 
     # Load selection function
-    with open('selection_function.pkl', 'rb') as f: # selection_function.pkl is generated by generate_selection_function.py
+    with open(outdir / '../selection_function.pkl', 'rb') as f: # selection_function.pkl is generated by generate_selection_function.py
         selfunc_interp = dill.load(f)
     def selection_function(x):
         return selfunc_interp(x)
@@ -45,9 +45,9 @@ except:
     np.savez(outdir / f"../{grid_label}.npz", mz=mz, H0=H0, z=z, m=m, model_pdf=model_pdf)
 
 print("Reading bounds and draws...")
-draws = load_density(outdir / f"draws/draws_observed_{label}.json")
+draws = load_density(outdir / f"draws/draws_observed_{label}.json") # follow README.md to generate draws
 
-bounds = np.loadtxt(outdir / "jsd_bounds.txt")
+bounds = np.loadtxt(outdir / "jsd_bounds.txt") # jsd_bounds.txt is generated by simulate_posterior_samples.py
 
 print("Preparing H0 inference...")
 # Mask out mz where there is no sample

### 2024-07-16 src/scripts/plot_simulation_comparison.py

@@ -8,21 +8,17 @@ label = 'simulation'
 outdir = paths.data / label
 
 draws = load_density(outdir / f'draws/draws_observed_{label}.json')
-
 mz, H0, z, m, model_pdf = np.load(outdir / '../grid.npz').values()
-
 bounds = np.loadtxt(outdir / "jsd_bounds.txt")
 
-fig = plot_median_cr(draws, save=True, show=False)
+fig = plot_median_cr(draws, hierarchical=True, save=True, show=False)
 ax = fig.axes[0]
-
 colors = ['tab:green', 'tab:red', 'tab:orange']
 for i, c in zip([np.argmin(abs(H0-40)), np.argmin(abs(H0-70)), np.argmin(abs(H0-100))], colors):
     ax.plot(mz, model_pdf[:,i]/np.trapz(model_pdf[:,i], mz), label=f"$H_0={H0[i]:.0f}$", c=c)
-ax.legend()
 ax.set_xlabel('$m^z_1\ [\mathrm{M}_\odot]$')
-ax.set_ylabel('Density')
+ax.set_ylabel('$\mathrm{Density}$')
 ax.set_xlim(bounds[0], bounds[1])
 ax.set_ylim(0)
-
-fig.savefig(outdir+"/comparison.pdf", bbox_inches='tight')
+ax.legend()
+fig.savefig(paths.figures / "comparison.pdf", bbox_inches='tight')

### 2024-07-16 src/scripts/plot_simulation_reconstruction.py

@@ -10,12 +10,12 @@ outdir = paths.data / 'simulation'
 draws = load_density(outdir / f'draws/draws_observed_{label}.json')
 obs_samples = np.loadtxt(outdir / 'obs_samples.txt')
 
-fig = plot_median_cr(draws, samples=obs_samples, save=True, show=False)
+fig = plot_median_cr(draws, hierarchical=True, save=True, show=False)
 ax = fig.axes[0]
-ax.hist(obs_samples, bins = int(np.sqrt(len(obs_samples))), histtype = 'step', density = True, label = '$\mathrm{median}(Y_t)$')
+ax.hist(obs_samples, bins = int(np.sqrt(len(obs_samples))), histtype = 'step', density = True, label = '$\mathrm{median}(Y_t)$', color = 'tab:red')
 ax.set_xlim(0,120)
 ax.set_ylim(0)
-ax.xlabel('$m^z_1\ [\mathrm{M}_\odot]$')
-ax.ylabel('Density')
+ax.set_xlabel('$m^z_1\ [\mathrm{M}_\odot]$')
+ax.set_ylabel('$\mathrm{Density}$')
 ax.legend()
 fig.savefig(paths.figures / 'simulation_reconstruction.pdf', bbox_inches='tight')

### 2024-07-16 src/scripts/plot_simulation_transformation.py

@@ -1,31 +1,29 @@
 from matplotlib import pyplot as plt
 import numpy as np
+from figaro import plot_settings
 import paths
 
 label = 'simulation'
 outdir = paths.data / label
 
-mz, H0, z, m, model_pdf = np.load(outdir / '../grid.npz').values()
-
 # Mass distribution
 from population_models.mass import plpeak
+mz, H0, z, m, model_pdf = np.load(outdir / '../grid.npz').values()
 
-fig, (ax1, ax2) = plt.subplots(2)
-
-ax1.plot(m, plpeak(m)/np.trapz(plpeak(m), m), label = "PL+Peak", c = 'tab:blue')
-ax1.xlabel('$m_1\ [\mathrm{M}_\odot]$')
-ax1.ylabel('$p(m_1|\Lambda)$')
+fig, (ax1, ax2) = plt.subplots(2, figsize=(7,6), gridspec_kw={'hspace': 0.3})
+m = np.linspace(0,120,1000)
+ax1.plot(m, plpeak(m)/np.trapz(plpeak(m), m), label = "$\mathrm{PL}+\mathrm{Peak}$", c = 'tab:blue')
+ax1.set_xlabel('$m_1\ [\mathrm{M}_\odot]$')
+ax1.set_ylabel('$p(m_1|\Lambda)$')
+ax1.set_xlim(0,120)
+ax1.set_ylim(0)
 ax1.legend()
-ax1.xlim(0,120)
-ax1.ylim(0)
-
 colors = ['tab:green', 'tab:red', 'tab:orange']
 for i, c in zip([np.argmin(abs(H0-40)), np.argmin(abs(H0-70)), np.argmin(abs(H0-100))], colors):
     ax2.plot(mz, model_pdf[:,i]/np.trapz(model_pdf[:,i], mz), label=f"$H_0={H0[i]:.0f}$", c=c)
+ax2.set_xlabel('$m^z_1\ [\mathrm{M}_\odot]$')
+ax2.set_ylabel('$p(m^z_1|\Lambda, \Omega(H_0), \mathrm{det})$')
+ax2.set_xlim(0,120)
+ax2.set_ylim(0)
 ax2.legend()
-ax2.xlabel('$m^z_1\ [\mathrm{M}_\odot]$')
-ax2.ylabel('$p(m^z_1|\Lambda, \Omega(H_0), \mathrm{det})$')
-ax2.xlim(0,120)
-ax2.ylim(0)
-
[…]

### 2024-07-16 src/scripts/plot_simulation_comparison.py
Change file name

@@ -21,4 +21,4 @@ ax.set_ylabel('$\mathrm{Density}$')
 ax.set_xlim(bounds[0], bounds[1])
 ax.set_ylim(0)
 ax.legend()
-fig.savefig(paths.figures / "comparison.pdf", bbox_inches='tight')
+fig.savefig(paths.figures / "simulation_comparison.pdf", bbox_inches='tight')

### 2024-07-17 src/scripts/plot_simulation_comparison.py
Add dependencies

@@ -8,7 +8,7 @@ label = 'simulation'
 outdir = paths.data / label
 
 draws = load_density(outdir / f'draws/draws_observed_{label}.json')
-mz, H0, z, m, model_pdf = np.load(outdir / '../grid.npz').values()
+mz, H0, z, m, model_pdf = np.load(paths.data / 'grid.npz').values()
 bounds = np.loadtxt(outdir / "jsd_bounds.txt")
 
 fig = plot_median_cr(draws, hierarchical=True, save=True, show=False)
@@ -16,9 +16,12 @@ ax = fig.axes[0]
 colors = ['tab:green', 'tab:red', 'tab:orange']
 for i, c in zip([np.argmin(abs(H0-40)), np.argmin(abs(H0-70)), np.argmin(abs(H0-100))], colors):
     ax.plot(mz, model_pdf[:,i]/np.trapz(model_pdf[:,i], mz), label=f"$H_0={H0[i]:.0f}$", c=c)
+y_max = ax.get_ylim()[1]
+ax.fill_between([0, bounds[0]], 0, 1, color='gray', alpha=0.2)
+ax.fill_between([bounds[1], 200], 0, 1, color='gray', alpha=0.2)
 ax.set_xlabel('$m^z_1\ [\mathrm{M}_\odot]$')
 ax.set_ylabel('$\mathrm{Density}$')
-ax.set_xlim(bounds[0], bounds[1])
-ax.set_ylim(0)
+ax.set_xlim(0, 200)
+ax.set_ylim(0, y_max)
 ax.legend()
 fig.savefig(paths.figures / "simulation_comparison.pdf", bbox_inches='tight')

### 2024-07-17 src/scripts/plot_simulation_reconstruction.py

@@ -13,7 +13,7 @@ obs_samples = np.loadtxt(outdir / 'obs_samples.txt')
 fig = plot_median_cr(draws, hierarchical=True, save=True, show=False)
 ax = fig.axes[0]
 ax.hist(obs_samples, bins = int(np.sqrt(len(obs_samples))), histtype = 'step', density = True, label = '$\mathrm{median}(Y_t)$', color = 'tab:red')
-ax.set_xlim(0,120)
+ax.set_xlim(0,200)
 ax.set_ylim(0)
 ax.set_xlabel('$m^z_1\ [\mathrm{M}_\odot]$')
 ax.set_ylabel('$\mathrm{Density}$')

### 2024-07-17 src/scripts/plot_simulation_transformation.py

@@ -3,19 +3,16 @@ import numpy as np
 from figaro import plot_settings
 import paths
 
-label = 'simulation'
-outdir = paths.data / label
-
 # Mass distribution
 from population_models.mass import plpeak
-mz, H0, z, m, model_pdf = np.load(outdir / '../grid.npz').values()
+mz, H0, z, m, model_pdf = np.load(paths.data / 'grid.npz').values()
 
 fig, (ax1, ax2) = plt.subplots(2, figsize=(7,6), gridspec_kw={'hspace': 0.3})
-m = np.linspace(0,120,1000)
+m = np.linspace(0,200,1000)
 ax1.plot(m, plpeak(m)/np.trapz(plpeak(m), m), label = "$\mathrm{PL}+\mathrm{Peak}$", c = 'tab:blue')
 ax1.set_xlabel('$m_1\ [\mathrm{M}_\odot]$')
 ax1.set_ylabel('$p(m_1|\Lambda)$')
-ax1.set_xlim(0,120)
+ax1.set_xlim(0,200)
 ax1.set_ylim(0)
 ax1.legend()
 colors = ['tab:green', 'tab:red', 'tab:orange']
@@ -23,7 +20,7 @@ for i, c in zip([np.argmin(abs(H0-40)), np.argmin(abs(H0-70)), np.argmin(abs(H0-
     ax2.plot(mz, model_pdf[:,i]/np.trapz(model_pdf[:,i], mz), label=f"$H_0={H0[i]:.0f}$", c=c)
 ax2.set_xlabel('$m^z_1\ [\mathrm{M}_\odot]$')
 ax2.set_ylabel('$p(m^z_1|\Lambda, \Omega(H_0), \mathrm{det})$')
-ax2.set_xlim(0,120)
+ax2.set_xlim(0,200)
 ax2.set_ylim(0)
 ax2.legend()
 fig.savefig(paths.figures / "simulation_transformation.pdf", bbox_inches='tight')

### 2024-07-17 src/scripts/simulate_posterior_samples.py

@@ -16,9 +16,9 @@ def p_z(z):
 
 # Generate true detector frame mass samples
 def generate_truth(n_single_event_draws):
-    M = rejection_sampler(n_single_event_draws, plpeak, [0,200])
+    m = rejection_sampler(n_single_event_draws, plpeak, [0,200])
     z = rejection_sampler(n_single_event_draws, p_z, [0, 2])
-    return np.array([M * (1 + z), z]).T
+    return np.array([m * (1 + z), z]).T
 
 # Generate detector frame mass posterior samples
 def generate_mz_posterior_samples(truth, sigma = 0.03, n_samples = 1000):

### 2024-07-17 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 In Sec.~\ref{sec:method}, we describe our modified method and the testing procedure.
-In Sec.~\ref{sec:simulation}, we present the analysis setup and results of the simulation study.
+In Sec.~\ref{sec:mock_data}, we present the analysis setup and results of the mock data study.
 In Sec.~\ref{sec:real_data}, we apply our method to the real data from the \ac{LVK}.
@@
             $m^z_1$ & Primary mass of the \ac{BBH} sources in the detector frame \\
-            $\mathbf{Y}$ & Set of posterior samples of the \ac{GW} events \\
+            $Y_t$ & source-frame primary mass posterior samples of the t-th \ac{GW} event \\
+            $\mathbf{Y}$ & Set of source-frame primary mass posterior samples of all \ac{GW} events \\
             $\mathbf{\Theta}$ & Set of hyperparameters of the \ac{(H)DPGMM} \\
@@
 It is important to note that the reconstruction is only valid in the region where the samples are present.
-In Fig.~XXX \todo{Add figure}, we show an example of the reconstruction obtained with \textsc{figaro}.
+In Fig.~\ref{fig:simulation_reconstruction}, we show an example of the reconstruction obtained with \textsc{figaro}.
 
@@
 By considering a continuous set of $\Omega$ and  $\Lambda$ of the population model, we can obtain a set of $p(m^z_1|\mathbf{\Lambda}, \mathbf{\Omega}, \mathrm{det})$, which is a set of continuous detector-frame observed population models for each combination of $\Omega_j$ and $\Lambda_j$, where $\mathbf{\Lambda} = \{\Lambda_j\}$ and $\mathbf{\Omega} = \{\Omega_j\}$.
-In Fig.~XXX \todo{Add figure}, we show an example of the transformation.
+In Fig.~\ref{fig:simulation_transformation}, we show an example of the transformation.
 
@@
 This is because the reconstruction is only valid in the region where observations are present.
+In Fig.~\ref{fig:simulation_comparison}, we show an illustration of the comparison between the two populations.
 
@@
 By considering this optimization procedure only, we can estimate the optimized parameters without performing computationally expensive likelihood calculations.
-In Fig.~XXX \todo{Add figure}, we show an example of the fitting procedure.
 
@@
 
-\section{Simulation}
-\label{sec:simulation}
+\section{Mock data}
+\label{sec:mock_data}
 
 % $H_0$ only
+\begin{figure}[h]
+    \script{plot_simulation_reconstruction.py}
+    \includegraphics[width=\linewidth]{figures/simulation_reconstruction.pdf}
+    \caption{
+        \todo{Add caption}
+    }
+    \label{fig:simulation_reconstruction}
+\end{figure}
+
+\begin{figure}[h]
+    \script{plot_simulation_transformation.py}
+    \includegraphics[width=\linewidth]{figures/simulation_transformation.pdf}
+    \caption{
+        \todo{Add caption}
+    }
+    \label{fig:simulation_transformation}
+\end{figure}
+
+\begin{figure}[h]
+    \script{plot_simulation_comparison.py}
+    \includegraphics[width=\linewidth]{figures/simulation_comparison.pdf}
+    \caption{
+        \todo{Add caption}
+    }
+    \label{fig:simulation_comparison}
+\end{figure}
+
+\begin{figure}[h]
+    \script{plot_simulation_result_H0.py}
+    \includegraphics[width=\linewidth]{figures/simulation_result_H0.pdf}
+    \caption{
+        \todo{Add caption}
+    }
+    \label{fig:simulation_result_H0}
+\end{figure}
 
@@
 % Power law + peak
-
-% Population model from simulation
 

### 2024-07-17 src/scripts/plot_simulation_result_H0.py
Add simulation result plot

@@ -1,0 +1,20 @@
+from figaro import plot_settings
+from figaro.cosmology import Planck18
+import numpy as np
+import matplotlib.pyplot as plt
+import paths
+
+H0_samples = np.loadtxt(paths.data / 'simulation' / 'H0s.txt')
+true_H0 = Planck18.h*100
+
+plt.hist(H0_samples)
+percs = np.percentile(H0_samples, [5, 16, 50, 84, 95])
+plt.axvline(true_H0, lw = 0.7, ls = '--', c = 'orangered', label = '$\\mathrm{Simulated}$')
+plt.axvline(percs[2], c = 'steelblue', lw=0.7, label = '$H_0 = '+str(f'{percs[2]:.1f}')+'^{+'+str(f'{percs[3]-percs[2]:.1f}')+'}_{-'+str(f'{percs[2]-percs[1]:.1f}')+'}$')
+plt.axvspan(percs[1], percs[3], alpha=0.25, color='mediumturquoise')
+plt.axvspan(percs[0], percs[4], alpha=0.25, color='darkturquoise')
+plt.legend()
+plt.xlabel('$H_0\ [\mathrm{km/s/Mpc}]$')
+plt.ylabel('$\mathrm{Density}$')
+plt.xlim(true_H0-20, true_H0+20)
+plt.savefig(paths.figures / "simulation_result_H0.pdf", bbox_inches='tight')

### 2024-07-17 src/scripts/plot_simulation_result_H0.py
Update plot_simulation_result_H0.py

@@ -9,7 +9,7 @@ true_H0 = Planck18.h*100
 
 plt.hist(H0_samples)
 percs = np.percentile(H0_samples, [5, 16, 50, 84, 95])
-plt.axvline(true_H0, lw = 0.7, ls = '--', c = 'orangered', label = '$\\mathrm{Simulated}$')
+plt.axvline(true_H0, lw = 0.7, ls = '--', c = 'orangered', label = '$H_0 = '+str(f'{true_H0:.1f}')+'(\mathrm{Simulated})$')
 plt.axvline(percs[2], c = 'steelblue', lw=0.7, label = '$H_0 = '+str(f'{percs[2]:.1f}')+'^{+'+str(f'{percs[3]-percs[2]:.1f}')+'}_{-'+str(f'{percs[2]-percs[1]:.1f}')+'}$')
 plt.axvspan(percs[1], percs[3], alpha=0.25, color='mediumturquoise')
 plt.axvspan(percs[0], percs[4], alpha=0.25, color='darkturquoise')

### 2024-07-17 src/scripts/plot_simulation_result_H0.py
New data set

@@ -9,7 +9,7 @@ true_H0 = Planck18.h*100
 
 plt.hist(H0_samples)
 percs = np.percentile(H0_samples, [5, 16, 50, 84, 95])
-plt.axvline(true_H0, lw = 0.7, ls = '--', c = 'orangered', label = '$H_0 = '+str(f'{true_H0:.1f}')+'(\mathrm{Simulated})$')
+plt.axvline(true_H0, lw = 0.7, ls = '--', c = 'orangered', label = '$H_0 = '+str(f'{true_H0:.1f}')+' (\mathrm{Simulated})$')
 plt.axvline(percs[2], c = 'steelblue', lw=0.7, label = '$H_0 = '+str(f'{percs[2]:.1f}')+'^{+'+str(f'{percs[3]-percs[2]:.1f}')+'}_{-'+str(f'{percs[2]-percs[1]:.1f}')+'}$')
 plt.axvspan(percs[1], percs[3], alpha=0.25, color='mediumturquoise')
 plt.axvspan(percs[0], percs[4], alpha=0.25, color='darkturquoise')

### 2024-07-18 src/tex/ms.tex
[showyourwork] overleaf sync

@@
             $m_1$ & Primary mass of the \ac{GW} sources \\
+            $q$ & Mass ratio of the \ac{GW} sources \\
             $z$ & Redshift of the \ac{GW} sources \\
@@
             $m^z_1$ & Primary mass of the \ac{BBH} sources in the detector frame \\
-            $Y_t$ & source-frame primary mass posterior samples of the t-th \ac{GW} event \\
+            $Y_t$ & source-frame primary mass posterior samples of the $t$-th \ac{GW} event \\
             $\mathbf{Y}$ & Set of source-frame primary mass posterior samples of all \ac{GW} events \\
@@
     \label{eq:transformation}
-        \begin{split}
+    \begin{aligned}
         &p(m^z_1|\Lambda, \Omega, \mathrm{det}) \\
@@
         &\propto \int p\left(\mathrm{det}\middle|\frac{m^z_1}{1+z},z,\Omega\right)p\left(\frac{m^z_1}{1+z}\middle|\Lambda\right)p(z|\Lambda) \mathrm{d}z\,,
-        \end{split}
+    \end{aligned}
 \end{equation}
 \end{widetext}
-where we assume the source-frame mass distribution does not evolve with redshift, and $\mathrm{det}$ represents whether the event is detectable or not.
+where we assume the $m_1$ distribution does not evolve with $z$, and $\mathrm{det}$ represents whether the event is detectable or not.
 The selection effect is included in the derivation in eq.~\eqref{eq:transformation}, specifically, the first term in the final expression, which is the selection function.
@@
 This procedure gives us a set of optimized $\Lambda_i$ and $\Omega_j$ for each $\Theta_i$.
-Note that we first calculate the median of the detector-frame primary mass posterior samples of each \ac{GW} event, then we calculate $d_\mathrm{JS}$ in the range from the minimum to the maximum of the medians.
+Note that we calculate $d_\mathrm{JS}$ in the range from the minimum to the maximum of the median of the $m^z_1$ posterior samples of each \ac{GW} event
 This is because the reconstruction is only valid in the region where observations are present.
@@
 
-The resulting distribution of the optimized parameters is not a posterior distribution in the Bayesian sense, as we do not consider the likelihood of the observed data given the population model.
+The resulting distribution of the optimized parameters is not a posterior distribution in a Bayesian sense, as we do not consider the likelihood of the observed data given the population model.
 Instead, it is a distribution of the optimized parameters that are consistent with the observed data, distributed according to the uncertainty in the reconstruction of the population.
@@
 
-The analysis code is publicly available.\footnote{\todo{Add link}}
+The analysis code is publicly available.\footnote{The analysis code is available at \url{https://github.com/thomasckng/det-frame-cosmo-with-FIGARO/tree/main/src/scripts}.}
 
@@
 \label{sec:mock_data}
+
+In this section, we test our method with a mock data study.
+To prepare the mock data, we first generated $10000$ sets of $m_1$ and $z$ samples from a \textsc{power law + peak} model for the $m_1$ distribution and a uniform distribution in comoving volume for the $z$ distribution.
+Here we assumed the $m_1$ distribution does not evolve with $z$.
+we then transformed the $m_1$ and $z$ samples to $m^z_1$ and $d_L$ samples.
+After the transformation, we applied a selection function to the $m^z_1$ and $d_L$ samples.
+We chose the O3 selection function from \todo{add reference} as the selection function.
+Note that we marginalized over $q$ for the selection function, similar to the transformation in Sec.~\ref{sec:transformation}.
+After applying the selection function, we generated the $m^z_1$ posterior samples for each true $m^z_1$ sample by
+\begin{gather*}
+    I \sim \mathcal{N}(0, 0.03)\,,\\
+    P \sim \mathcal{N}(\log(T_t)+I, 0.03)\,,\\
+    Y_t = \exp(P)\,,
+\end{gather*}
+where $\mathcal{N}(\mu, \sigma)$ is a normal distribution with mean $\mu$ and standard deviation $\sigma$, $I$ is the intrinsic error, $P$ is the posterior sample in the log space, $T_t$ is the true $m^z_1$ sample, and $Y_t$ is the posterior sample for the $t$-th \ac{GW} event.
 

### 2024-07-18 src/scripts/paths.py
Try fixing GH Action plots

@@ -28,7 +28,3 @@ figures = tex / "figures"
 
 # Absolute path to the `src/tex/output` folder (contains other user-defined output)
 output = tex / "output"
-
-import os
-from pathlib import Path
-os.environ["PATH"] += os.pathsep + str(Path.home() / "bin")

### 2024-07-19 src/scripts/plot_simulation_reconstruction.py
Update plot_simulation_reconstruction.py

@@ -2,6 +2,7 @@ from figaro.load import load_density
 from figaro.plot import plot_median_cr
 from figaro import plot_settings
 import numpy as np
+from matplotlib import pyplot as plt
 import paths
 
 label = 'simulation'
@@ -10,12 +11,22 @@ outdir = paths.data / 'simulation'
 draws = load_density(outdir / f'draws/draws_observed_{label}.json')
 obs_samples = np.loadtxt(outdir / 'obs_samples.txt')
 
-fig = plot_median_cr(draws, hierarchical=True, save=True, show=False)
-ax = fig.axes[0]
-ax.hist(obs_samples, bins = int(np.sqrt(len(obs_samples))), histtype = 'step', density = True, label = '$\mathrm{median}(Y_t)$', color = 'tab:red')
-ax.set_xlim(0,200)
-ax.set_ylim(0)
-ax.set_xlabel('$m^z_1\ [\mathrm{M}_\odot]$')
-ax.set_ylabel('$\mathrm{Density}$')
-ax.legend()
+fig, (ax1, ax2) = plt.subplots(2, figsize=(7,6), gridspec_kw={'hspace': 0.3})
+fig = plot_median_cr(draws, hierarchical=True, save=True, show=False, fig=fig)
+ax1.hist(obs_samples, bins = int(np.sqrt(len(obs_samples))), histtype = 'step', density = True, label = '$\mathrm{median}(Y_t)$', color = 'tab:red')
+ax1.set_xlim(0,200)
+ax1.set_ylim(0)
+ax1.set_xlabel('$m^z_1\ [\mathrm{M}_\odot]$')
+ax1.set_ylabel('$\mathrm{Density}$')
+ax1.legend()
+m_z = np.linspace(0,200,1000)
+ax2.plot(m_z, draws[0].pdf(m_z), alpha = 0.5, color = 'tab:blue', label = '$\mathrm{(H)DPGMM}$')
+for i, draw in enumerate(draws[1:10]):
+    ax2.plot(m_z, draw.pdf(m_z), alpha = 0.5, color = 'tab:blue')
+ax2.hist(obs_samples, bins = int(np.sqrt(len(obs_samples))), histtype = 'step', density = True, label = '$\mathrm{median}(Y_t)$', color = 'tab:red')
+ax2.set_xlim(0,200)
+ax2.set_ylim(0)
+ax2.set_xlabel('$m^z_1\ [\mathrm{M}_\odot]$')
[…]

### 2024-07-19 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
-We use a \ac{(H)DPGMM} developed in \citet{Rinaldi:2021bhm} as a non-parametric model to reconstruct the detector-frame observed mass population distribution $p(m^z_1|\mathbf{\Theta})$.
+We use the \ac{(H)DPGMM} developed in \citet{Rinaldi:2021bhm} as a non-parametric model to reconstruct the detector-frame observed mass population distribution $p(m^z_1|\mathbf{\Theta})$.
 The \ac{(H)DPGMM} is a hierarchy of Gaussian mixture models with a Dirichlet process prior, which allows the model to be flexible to reconstruct arbitrary population distribution from posterior samples.
@@
 For each reconstructed distribution $p(m^z_1|\Theta_i)$, we find the optimized $\Lambda_j$ and $\Omega_j$ by minimizing $d_\mathrm{JS}$.
-For the minimization, we use the \textsc{scipy} \citep{2020SciPy-NMeth} implementation of the \todo{XXX} algorithm. \todo{Add reference for the algorithm}
+For the minimization, we use the \textsc{scipy} \citep{2020SciPy-NMeth} implementation of the \todo{XXX} algorithm. \todo{Add reference for the algorithm and scipy}
 This procedure gives us a set of optimized $\Lambda_i$ and $\Omega_j$ for each $\Theta_i$.
@@
 In this section, we test our method with a mock data study.
-To prepare the mock data, we first generated $10000$ sets of $m_1$ and $z$ samples from a \textsc{power law + peak} model for the $m_1$ distribution and a uniform distribution in comoving volume for the $z$ distribution.
+To prepare the mock data, we first generated $10000$ sets of $m_1$ and $z$ samples from a \textsc{power law + peak} model for the $m_1$ distribution and uniform distribution over comoving volume and source-frame time (see, e.g., Eq.~11 in \cite{LIGOScientific:2019zcs}) for the $z$ distribution.
+\todo{Add reference for models}
 Here we assumed the $m_1$ distribution does not evolve with $z$.
-we then transformed the $m_1$ and $z$ samples to $m^z_1$ and $d_L$ samples.
+We then transformed the $m_1$ and $z$ samples to $m^z_1$ and $d_L$ samples.
+Here we chose the $\Lambda$CDM model with parameters consistent with the Planck 2018 results \cite{Planck:2018vyg} as the cosmological model for the transformation of the samples.
 After the transformation, we applied a selection function to the $m^z_1$ and $d_L$ samples.
@@
 Note that we marginalized over $q$ for the selection function, similar to the transformation in Sec.~\ref{sec:transformation}.
-After applying the selection function, we generated the $m^z_1$ posterior samples for each true $m^z_1$ sample by
+After applying the selection function, there were $262$ sets of samples left.
+We then generated the $m^z_1$ posterior samples for each true $m^z_1$ sample by
 \begin{gather*}
@@
 
-% $H_0$ only
+After preparing the mock data, we first performed the reconstruction of the observed $m^z_1$ population distribution with \textsc{figaro} as described in Sec.~\ref{sec:reconstruction}.
+The result of the reconstruction is shown in Fig.~\ref{fig:simulation_reconstruction}.
+
 \begin{figure}[h]
@@
     \caption{
-        \todo{Add caption}
+        In both panels, the red line is the histogram of the median of the $m^z_1$ posterior samples of the \ac{GW} events.
+        In the upper panel, the blue line is the median of the reconstructed distributions, and the shaded region represents the 68\% and 90\% credible intervals of the reconstructed distributions.
+        In the lower panel, the blue lines are the single reconstructed distributions from a subset of all the reconstructed distributions.
+        By combining all the single reconstructed distributions, we obtain the distribution in the upper panel.
     }
@@
 \end{figure}
+
+Next, we chose the true population model we used to generate the mock data as the source-frame population model $\Lambda$.
+We then transformed $p(m_1|\Lambda)$ to the detector frame with the true cosmological model we used to generate the mock data, except we did not fix $H_0$.
+We performed the transformation for each $H_0$ in the range from $5$ to $150$.
+After the transformation, we applied the selection function we used to generate the mock data to the transformed distribution.
+Here we only free $H_0$ in the transformation to test if our method can recover the true $H_0$ while we fixed the other cosmological parameters to the true values.
+The result of the transformation is shown in Fig.~\ref{fig:simulation_transformation}.
 
@@
     \caption{
-        \todo{Add caption}
+        The upper panel shows the source-frame $m_1$ population distribution, which we chose as the true population model.
+        The lower panel shows the transformed detector-frame $m^z_1$ population distributions for different $H_0$.
+        The source-frame population distribution is redshifted towards higher masses as $H_0$ increases, as $m^z_1 = m_1(1+z)$ and $z$ increases with $H_0$ for a given $d_L$.
+        The low-mass peak is suppressed due to the selection function.
     }
@@
 \end{figure}
+
+Finally, we performed the \ac{PE} to infer $H_0$.
+For this test specifically, instead of using the \textsc{scipy} optimization algorithm, we used a grid search to find the optimized $H_0$, as the parameter space is one-dimensional.
+Figure~\ref{fig:simulation_comparison} shows an illustration of the comparison between the reconstructed observed population and the transformed population for different $H_0$.
+For each $H_0$, we computed the $d_\mathrm{JS}$ between the two populations and found the optimized $H_0$ that minimizes $d_\mathrm{JS}$.
+The result of the \ac{PE} is shown in Fig.~\ref{fig:simulation_result_H0}.
 
@@
 
+% $H_0$ only
 \begin{figure}[h]

### 2024-07-22 src/scripts/paths.py
Fixing GH Action

@@ -28,3 +28,7 @@ figures = tex / "figures"
 
 # Absolute path to the `src/tex/output` folder (contains other user-defined output)
 output = tex / "output"
+
+import os
+from pathlib import Path
+os.environ["PATH"] += os.pathsep + str(Path.home() / "bin")

### 2024-07-23 src/scripts/plot_simulation_comparison.py
Modify plots

@@ -16,12 +16,11 @@ ax = fig.axes[0]
 colors = ['tab:green', 'tab:red', 'tab:orange']
 for i, c in zip([np.argmin(abs(H0-40)), np.argmin(abs(H0-70)), np.argmin(abs(H0-100))], colors):
     ax.plot(mz, model_pdf[:,i]/np.trapz(model_pdf[:,i], mz), label=f"$H_0={H0[i]:.0f}$", c=c)
-y_max = ax.get_ylim()[1]
-ax.fill_between([0, bounds[0]], 0, 1, color='gray', alpha=0.2)
-ax.fill_between([bounds[1], 200], 0, 1, color='gray', alpha=0.2)
+ax.axvspans(0, bounds[0], color='gray', alpha=0.2)
+ax.axvspans(bounds[1], 200, color='gray', alpha=0.2)
 ax.set_xlabel('$m^z_1\ [\mathrm{M}_\odot]$')
 ax.set_ylabel('$\mathrm{Density}$')
 ax.set_xlim(0, 200)
-ax.set_ylim(0, y_max)
+ax.set_ylim(0)
 ax.legend()
 fig.savefig(paths.figures / "simulation_comparison.pdf", bbox_inches='tight')

### 2024-07-23 src/scripts/plot_simulation_result_H0.py

@@ -7,14 +7,15 @@ import paths
 H0_samples = np.loadtxt(paths.data / 'simulation' / 'H0s.txt')
 true_H0 = Planck18.h*100
 
-plt.hist(H0_samples)
+fig, ax = plt.subplots()
+ax.hist(H0_samples, bins = int(np.sqrt(len(H0_samples))), histtype = 'step', density = True, color = 'tab:blue')
 percs = np.percentile(H0_samples, [5, 16, 50, 84, 95])
-plt.axvline(true_H0, lw = 0.7, ls = '--', c = 'orangered', label = '$H_0 = '+str(f'{true_H0:.1f}')+' (\mathrm{Simulated})$')
-plt.axvline(percs[2], c = 'steelblue', lw=0.7, label = '$H_0 = '+str(f'{percs[2]:.1f}')+'^{+'+str(f'{percs[3]-percs[2]:.1f}')+'}_{-'+str(f'{percs[2]-percs[1]:.1f}')+'}$')
-plt.axvspan(percs[1], percs[3], alpha=0.25, color='mediumturquoise')
-plt.axvspan(percs[0], percs[4], alpha=0.25, color='darkturquoise')
-plt.legend()
-plt.xlabel('$H_0\ [\mathrm{km/s/Mpc}]$')
-plt.ylabel('$\mathrm{Density}$')
-plt.xlim(true_H0-20, true_H0+20)
-plt.savefig(paths.figures / "simulation_result_H0.pdf", bbox_inches='tight')
+ax.axvline(true_H0, lw = 0.7, ls = '--', c = 'orangered', label = '$H_0 = '+str(f'{true_H0:.1f}')+' (\mathrm{Simulated})$')
+ax.axvline(percs[2], c = 'steelblue', lw=0.7, label = '$H_0 = '+str(f'{percs[2]:.1f}')+'^{+'+str(f'{percs[3]-percs[2]:.1f}')+'}_{-'+str(f'{percs[2]-percs[1]:.1f}')+'}$')
+ax.axvspan(percs[1], percs[3], alpha=0.25, color='mediumturquoise')
+ax.axvspan(percs[0], percs[4], alpha=0.25, color='darkturquoise')
+ax.legend()
+ax.set_xlabel('$H_0\ [\mathrm{km/s/Mpc}]$')
+ax.set_ylabel('$\mathrm{Density}$')
+ax.set_xlim(true_H0-20, true_H0+20)
+fig.savefig(paths.figures / "simulation_result_H0.pdf", bbox_inches='tight')

### 2024-07-23 src/scripts/plot_simulation_comparison.py
Update plot_simulation_comparison.py

@@ -16,8 +16,8 @@ ax = fig.axes[0]
 colors = ['tab:green', 'tab:red', 'tab:orange']
 for i, c in zip([np.argmin(abs(H0-40)), np.argmin(abs(H0-70)), np.argmin(abs(H0-100))], colors):
     ax.plot(mz, model_pdf[:,i]/np.trapz(model_pdf[:,i], mz), label=f"$H_0={H0[i]:.0f}$", c=c)
-ax.axvspans(0, bounds[0], color='gray', alpha=0.2)
-ax.axvspans(bounds[1], 200, color='gray', alpha=0.2)
+ax.axvspan(0, bounds[0], color='gray', alpha=0.2)
+ax.axvspan(bounds[1], 200, color='gray', alpha=0.2)
 ax.set_xlabel('$m^z_1\ [\mathrm{M}_\odot]$')
 ax.set_ylabel('$\mathrm{Density}$')
 ax.set_xlim(0, 200)

### 2024-08-05 src/tex/ms.tex
[showyourwork] overleaf sync

@@
         The upper panel shows the source-frame $m_1$ population distribution, which we chose as the true population model.
-        The lower panel shows the transformed detector-frame $m^z_1$ population distributions for different $H_0$.
+        The lower panel shows the transformed detector-frame $m^z_1$ population distributions for different $H_0$, where the corresponding $H_0$ is shown in the legend.
         The source-frame population distribution is redshifted towards higher masses as $H_0$ increases, as $m^z_1 = m_1(1+z)$ and $z$ increases with $H_0$ for a given $d_L$.
@@
     \caption{
-        \todo{Add caption}
+        The blue line and bands are the ones from Fig.~\ref{fig:simulation_reconstruction}, which represent the reconstructed observed population.
+        The other colored lines are the ones from Fig.~\ref{fig:simulation_transformation}, which represent the transformed population for different $H_0$.
+        The shaded regions represent the boundaries for the calculation of $d_\mathrm{JS}$, i.e., the $d_\mathrm{JS}$ calculation only considers the non-shaded region.
     }
@@
     \caption{
-        \todo{Add caption}
+        The histogram shows the distribution of the optimized $H_0$.
+        The vertical blue line and shaded regions represent the median, 68\%, and 90\% credible intervals of the distribution.
+        The vertical orange line represents the true $H_0$ we used to generate the mock data.
+        Note that the distribution only shows the uncertainty in the reconstruction.
+        The uncertainty in the comparison is not considered in the distribution.
     }

### 2024-08-05 src/scripts/generate_posterior_samples.py
Update generate_posterior_samples.py

@@ -31,7 +31,7 @@ with open(paths.data/'selection_function.pkl', 'rb') as f: # selection_function.
 def selection_function(x):
     return selfunc_interp(x)
 
-n_single_event_draws = 100_000
+n_single_event_draws = 10_000
 outdir = 'simulation'
 if not os.path.exists(paths.data / outdir):
     os.makedirs(paths.data / outdir)

### 2024-08-07 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 Considering the increasing number of \ac{GW} events and the improvement of the detectors, we expect that the spectral siren method will give us an independent and precise measurement of $H_0$ which is comparable to the previous measurements \citep{Planck:2018vyg, Riess:2021jrx}.
+However, performing standard hierarchical Bayesian inference with the spectral siren method is computationally expensive, and it depends on the choice of the population model and the cosmological model.
+Our method provides an alternative to studying different models without performing lengthy hierarchical analysis.
+This allows us to explore more models before performing the full hierarchical Bayesian inference, which makes our method a powerful diagnostic tool for model selection.
 
@@
 \begin{table}
-    \label{tab:notation}
     \caption{Notation}
@@
     \end{ruledtabular}
+    \label{tab:notation}
 \end{table}
@@
 Our method consists of three main steps: the non-parametric reconstruction of the detector-frame observed population, the transformation of the source-frame population model to the detector frame, and the \ac{PE}.
-We show the flowchart of our method in Fig.~XXX. \todo{Add flowchart}
 
@@
 \begin{equation}
-    \label{eq:transformation}
     \begin{aligned}
@@
     \end{aligned}
+    \label{eq:transformation}
 \end{equation}
@@
 where we assume the $m_1$ distribution does not evolve with $z$, and $\mathrm{det}$ represents whether the event is detectable or not.
-The selection effect is included in the derivation in eq.~\eqref{eq:transformation}, specifically, the first term in the final expression, which is the selection function.
+The selection effect is included in the derivation in Eq.~\eqref{eq:transformation}, specifically, the first term in the final expression, which is the selection function.
 Note that the selection function in general is a function of the detector-frame mass $m^z_1$, luminosity distance $d_L$, and mass ratio $q$.
@@
 
-With different $\Omega$, the resulting $p(m^z_1|\Lambda, \Omega, \mathrm{det})$ will be different.
-By considering a continuous set of $\Omega$ and  $\Lambda$ of the population model, we can obtain a set of $p(m^z_1|\mathbf{\Lambda}, \mathbf{\Omega}, \mathrm{det})$, which is a set of continuous detector-frame observed population models for each combination of $\Omega_j$ and $\Lambda_j$, where $\mathbf{\Lambda} = \{\Lambda_j\}$ and $\mathbf{\Omega} = \{\Omega_j\}$.
+With different $\Lambda$ or $\Omega$, the resulting $p(m^z_1|\Lambda, \Omega, \mathrm{det})$ will be different.
 In Fig.~\ref{fig:simulation_transformation}, we show an example of the transformation.
@@
 
-With $p(m^z_1|\mathbf{\Theta})$ and $p(m^z_1|\mathbf{\Lambda}, \mathbf{\Omega}, \mathrm{det})$ obtained from Sec.~\ref{sec:reconstruction} and Sec.~\ref{sec:transformation} respectively, we can perform the \ac{PE} to infer the free parameters in the parametric models assumed in Sec.~\ref{sec:transformation}.
-The \ac{PE} is performed by comparing $p(m^z_1|\Theta_i)$ with $p(m^z_1|\Lambda_j, \Omega_j)$.
+With $p(m^z_1|\mathbf{\Theta})$ and $p(m^z_1|\Lambda, \Omega, \mathrm{det})$ obtained from Sec.~\ref{sec:reconstruction} and Sec.~\ref{sec:transformation} respectively, we can perform the \ac{PE} to infer the free parameters in the parametric models assumed in Sec.~\ref{sec:transformation}.
+The \ac{PE} is performed by comparing $p(m^z_1|\Theta_i)$ with $p(m^z_1|\Lambda, \Omega)$.
 To compare the two populations, we use the Jensen-Shannon distance $d_\mathrm{JS}$ as the distance measure.
@@
 
-For each reconstructed distribution $p(m^z_1|\Theta_i)$, we find the optimized $\Lambda_j$ and $\Omega_j$ by minimizing $d_\mathrm{JS}$.
-For the minimization, we use the \textsc{scipy} \citep{2020SciPy-NMeth} implementation of the \todo{XXX} algorithm. \todo{Add reference for the algorithm and scipy}
-This procedure gives us a set of optimized $\Lambda_i$ and $\Omega_j$ for each $\Theta_i$.
+For each reconstructed distribution $p(m^z_1|\Theta_i)$, we find an optimized $\Lambda_i$ and $\Omega_i$ by minimizing $d_\mathrm{JS}$.
+For the minimization, we use the \textsc{scipy} \citep{2020SciPy-NMeth} implementation of the \todo{XXX} algorithm. \todo{Add reference for the algorithm}
 Note that we calculate $d_\mathrm{JS}$ in the range from the minimum to the maximum of the median of the $m^z_1$ posterior samples of each \ac{GW} event
 This is because the reconstruction is only valid in the region where observations are present.
+This procedure gives us a set of optimized $\Lambda_i$ and $\Omega_i$ for each $\Theta_i$.
 In Fig.~\ref{fig:simulation_comparison}, we show an illustration of the comparison between the two populations.
@@
 In this section, we test our method with a mock data study.
-To prepare the mock data, we first generated $10000$ sets of $m_1$ and $z$ samples from a \textsc{power law + peak} model for the $m_1$ distribution and uniform distribution over comoving volume and source-frame time (see, e.g., Eq.~11 in \cite{LIGOScientific:2019zcs}) for the $z$ distribution.
-\todo{Add reference for models}
+To prepare the mock data, we first generated $10000$ sets of $m_1$ and $z$ samples from a \textsc{power law + peak} distribution (see, e.g., Eq.~B4 in \cite{KAGRA:2021duu}) for the $m_1$ distribution and uniform distribution over comoving volume and source-frame time (see, e.g., Eq.~11 in \cite{LIGOScientific:2019zcs}) for the $z$ distribution.
 Here we assumed the $m_1$ distribution does not evolve with $z$.
@@
 Here we chose the $\Lambda$CDM model with parameters consistent with the Planck 2018 results \cite{Planck:2018vyg} as the cosmological model for the transformation of the samples.
+
 After the transformation, we applied a selection function to the $m^z_1$ and $d_L$ samples.
@@
 After applying the selection function, there were $262$ sets of samples left.
+
 We then generated the $m^z_1$ posterior samples for each true $m^z_1$ sample by
@@
 \end{gather*}
-where $\mathcal{N}(\mu, \sigma)$ is a normal distribution with mean $\mu$ and standard deviation $\sigma$, $I$ is the intrinsic error, $P$ is the posterior sample in the log space, $T_t$ is the true $m^z_1$ sample, and $Y_t$ is the posterior sample for the $t$-th \ac{GW} event.
+where $\mathcal{N}(\mu, \sigma)$ is a normal distribution with mean $\mu$ and standard deviation $\sigma$, $I$ is the difference between the true $m^z_1$ value and the measured mean in log space, $P$ is the posterior sample in the log space, $T_t$ is the true $m^z_1$ sample, and $Y_t$ is the posterior sample for the $t$-th \ac{GW} event.
+For each true $m^z_1$ sample, we draw a random $I$ and generate $1000$ $Y_t$ samples.
+This procedure ensures the posterior samples are positive.
+After preparing the mock data, we test our method with two cases: the inference of $H_0$ only and the inference of $H_0$ and a subset of the population model parameters.
 
-After preparing the mock data, we first performed the reconstruction of the observed $m^z_1$ population distribution with \textsc{figaro} as described in Sec.~\ref{sec:reconstruction}.
+\subsection{Inference of $H_0$}
+
+In this subsection, we first test our method with a simplified case where we only infer $H_0$.
+We assume the selection function, the true population model and its parameters, and the true cosmological model and its parameters except $H_0$ are known.
+
+We first performed the reconstruction of the observed $m^z_1$ population distribution with \textsc{figaro} as described in Sec.~\ref{sec:reconstruction}.
 The result of the reconstruction is shown in Fig.~\ref{fig:simulation_reconstruction}.
@@
 
-% $H_0$ only
 \begin{figure}[h]
@@
 
-% $H_0$ and $\Lambda$
+\subsection{Inference of $H_0$ and $\Lambda$}
+
+In this subsection, we test our method with a more realistic case where we infer $H_0$ and a subset of the population model parameters.
+Similar to the previous subsection, we assume only the free parameters are unknown.
+
+The reconstruction of the observed $m^z_1$ population distribution is the same as in the previous subsection.
+Note that the same reconstruction result can be reused, as the data does not change.
 
@@
 
+% PL+P issue
+    % Gregoire Pierra's paper
+
 \section{Discussion}
@@
 
-% Comparison with other methods
+\subsection{Comparison with previous work}
+\label{sec:comparison}
+% diagnostic tool
+% model selection
 
-% PL+P issue
-    % Gregoire Pierra's paper
+\subsection{Future work}
+\label{sec:future_work}
+As mentioned in Sec.~\ref{sec:pe}, our method is not a Bayesian inference, as we only consider the optimization of the parameters.
+To perform a full Bayesian inference, we need to develop a likelihood function that considers the comparison of the two populations.
+With this likelihood, our method will give the same result as the standard hierarchical Bayesian inference, as long as the number of reconstructions is large enough.
 
-% Future directions
-    % Likelihood
-    % luminosity distance/redshift population
+Another potential improvement is to consider not only the primary mass distribution.
+For example, we can consider the joint distribution of the primary mass and luminosity distance.
+The reconstruction of the joint observed distribution can also be performed in the current framework.
+However, we need a joint population model that considers both the primary mass and luminosity distance.
+Performing the analysis of the joint distribution will allow us to study mass population models that evolve with redshift.
+Furthermore, the inference of cosmological parameters will be more precise, as the information from the luminosity distance is also considered.
 

### 2024-08-07 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 % Previous work
-The fundamental idea of the spectral siren method is to infer cosmology by comparing the observed mass population with the intrinsic mass population since the observed mass population is redshifted due to the expansion of the Universe.
+The fundamental idea of the spectral siren method is to infer cosmology by comparing the observed mass population distribution with the intrinsic mass population distribution since the observed distribution is redshifted due to the expansion of the Universe.
 Assumptions about the intrinsic population model and cosmological model are required to perform the comparison.
@@
 \label{sec:method}
-In this section, we present the framework we developed to infer the cosmological parameters leveraging the detector-frame mass distribution.
+In this section, we present the framework we developed to infer both the intrinsic population model and the cosmological model simultaneously by leveraging the detector-frame mass population distribution.
 In the remainder of this Section, we will make use of the notation summarized in Table~\ref{tab:notation}, mutuated from \citet{Rinaldi:2021bhm, Rinaldi:2022kyg}.
@@
             \hline
-            $m_1$ & Primary mass of the \ac{GW} sources \\
+            $m_1$ & Primary mass of the \ac{GW} sources in the source frame \\
             $q$ & Mass ratio of the \ac{GW} sources \\
@@
             $m^z_1$ & Primary mass of the \ac{BBH} sources in the detector frame \\
-            $Y_t$ & source-frame primary mass posterior samples of the $t$-th \ac{GW} event \\
-            $\mathbf{Y}$ & Set of source-frame primary mass posterior samples of all \ac{GW} events \\
+            $Y_t$ & detector-frame primary mass posterior samples of the $t$-th \ac{GW} event \\
+            $\mathbf{Y}$ & Set of detector-frame primary mass posterior samples of all \ac{GW} events \\
+            $\Theta_i$ & Hyperparameters of the \ac{(H)DPGMM} \\
             $\mathbf{\Theta}$ & Set of hyperparameters of the \ac{(H)DPGMM} \\
-            $p(m^z_1|\mathbf{\Theta})$ & Detector-frame observed mass population distribution reconstructed with \ac{(H)DPGMM} \\
+            $p(m^z_1|\mathbf{\Theta})$ & Detector-frame observed primary mass population distribution reconstructed with the \ac{(H)DPGMM} \\
             $\Lambda$ & Source-frame population model \\
             $\Omega$ & Cosmological model \\
-            $p(m_1|\Lambda)$ & Source-frame mass population distribution assuming the population model $\Lambda$ \\
-            $p(m^z_1|\Lambda, \Omega)$ & Detector-frame mass population distribution assuming the population model $\Lambda$ and cosmological model $\Omega$ \\
+            $p(m_1|\Lambda)$ & Source-frame primary mass population distribution assuming the population model $\Lambda$ \\
+            $p(m^z_1|\Lambda, \Omega)$ & Detector-frame primary mass population distribution assuming the population model $\Lambda$ and cosmological model $\Omega$ \\
             $\mathrm{det}$ & detectability of the \ac{GW} events \\
@@
 Specifically, we propose two major modifications to the standard spectral siren method.
-First, instead of transforming the individual posterior samples from the detector frame to the source frame during the \ac{PE} process, we transform the population model to the detector frame before performing \ac{PE}.
-Second, instead of computing the likelihood of the observed data given the population model, we reconstruct the detector-frame observed population directly from the data using a non-parametric method.
+First, instead of transforming the individual-event posterior samples from the detector frame to the source frame during the \ac{PE} process, we transform the population model to the detector frame before performing \ac{PE}.
+Second, instead of directly computing the likelihood of the observed data given the population model, we reconstruct the observed population distribution from the data using a non-parametric method and compare it with the population model.
 
@@
 The first part is the reconstruction of the observed population from the data, which does not require any assumptions about the population model, and the intermediate result is completely data-driven.
-We make use of a \ac{(H)DPGMM} developed in \citet{Rinaldi:2021bhm} as a non-parametric model to reconstruct the observed population distribution.
+We make use of the \ac{(H)DPGMM} developed in \citet{Rinaldi:2021bhm} as a non-parametric model to reconstruct the observed population distribution.
 Other assumptions including the population model, cosmological model and selection function are then incorporated in the second half.
@@
 we choose to consider the primary mass $m_1$ of the \ac{BBH} sources, in principle, we can consider more parameters at once.
-From the posterior samples obtained from the \ac{PE} of each \ac{GW} event, we reconstruct the detector-frame observed mass distribution.
+From the posterior samples obtained from the \ac{PE} of each \ac{GW} event, we reconstruct the observed detector-frame primary mass population distribution.
 
-We use the \ac{(H)DPGMM} developed in \citet{Rinaldi:2021bhm} as a non-parametric model to reconstruct the detector-frame observed mass population distribution $p(m^z_1|\mathbf{\Theta})$.
+We use the \ac{(H)DPGMM} developed in \citet{Rinaldi:2021bhm} as a non-parametric model to reconstruct the observed $m^z_1$ population distribution $p(m^z_1|\mathbf{\Theta})$.
 The \ac{(H)DPGMM} is a hierarchy of Gaussian mixture models with a Dirichlet process prior, which allows the model to be flexible to reconstruct arbitrary population distribution from posterior samples.
@@
 
-On the other hand, we assume a population model $\Lambda$ to obtain a source-frame mass population $p(m_1|\Lambda)$, which represents the underlying population of the sources.
+On the other hand, we assume a population model $\Lambda$ to obtain a source-frame primary mass $m_1$ population distribution $p(m_1|\Lambda)$, which represents the underlying population distribution given the model $\Lambda$.
 Note that, either a parametric model can be used to only assume the form of the population distribution, or a fixed population model can be used to assume an exact population distribution.
-We then transform $p(m_1|\Lambda)$ to the detector frame with a given cosmological model $\Omega$ to obtain the detector-frame mass population $p(m^z_1|\Lambda, \Omega)$ by
+We then transform $p(m_1|\Lambda)$ to the detector frame with a given cosmological model $\Omega$ to obtain the population distribution $p(m^z_1|\Lambda, \Omega)$ by
 \begin{widetext}
@@
 \end{widetext}
-where we assume the $m_1$ distribution does not evolve with $z$, and $\mathrm{det}$ represents whether the event is detectable or not.
+where $\mathrm{det}$ represents whether the event is detectable or not.
+We assume the $m_1$ distribution does not evolve with $z$.
 The selection effect is included in the derivation in Eq.~\eqref{eq:transformation}, specifically, the first term in the final expression, which is the selection function.
 Note that the selection function in general is a function of the detector-frame mass $m^z_1$, luminosity distance $d_L$, and mass ratio $q$.
-In this work, we marginalized over $q$ to study the primary mass distribution only.
+In this work, we marginalized over $q$ to study the $m_1$ population distribution only.
 Due to the censoring of the low and high mass events, we consider the selection function in the transformation instead of the reconstruction.
-
 With different $\Lambda$ or $\Omega$, the resulting $p(m^z_1|\Lambda, \Omega, \mathrm{det})$ will be different.
@@
 Opposite to the first part of the method, this part includes the assumptions about the population and cosmological model, as well as the selection function.
-After obtaining $p(m^z_1|\mathbf{\Lambda}, \mathbf{\Omega}, \mathrm{det})$, we can perform the \ac{PE} to infer the optimized $\Omega_j$ and $\Lambda_j$.
+After obtaining $p(m^z_1|\Lambda, \Omega, \mathrm{det})$, we can perform the \ac{PE} to infer the optimized $\Omega$ and $\Lambda$.
 

### 2024-08-07 src/scripts/plot_simulation_result_H0.py
Modify plots

@@ -10,8 +10,8 @@ true_H0 = Planck18.h*100
 fig, ax = plt.subplots()
 ax.hist(H0_samples, bins = int(np.sqrt(len(H0_samples))), histtype = 'step', density = True, color = 'tab:blue')
 percs = np.percentile(H0_samples, [5, 16, 50, 84, 95])
-ax.axvline(true_H0, lw = 0.7, ls = '--', c = 'orangered', label = '$H_0 = '+str(f'{true_H0:.1f}')+' (\mathrm{Simulated})$')
-ax.axvline(percs[2], c = 'steelblue', lw=0.7, label = '$H_0 = '+str(f'{percs[2]:.1f}')+'^{+'+str(f'{percs[3]-percs[2]:.1f}')+'}_{-'+str(f'{percs[2]-percs[1]:.1f}')+'}$')
+ax.axvline(true_H0, lw = 0.7, ls = '--', c = 'orangered', label = f'$H_0={true_H0:.1f}\ (\mathrm{{Simulated}})$')
+ax.axvline(percs[2], c = 'steelblue', lw=0.7, label = f'$H_0={percs[2]:.1f}^{{+{percs[3]-percs[2]:.1f}}}_{{-{percs[2]-percs[1]:.1f}}}$')
 ax.axvspan(percs[1], percs[3], alpha=0.25, color='mediumturquoise')
 ax.axvspan(percs[0], percs[4], alpha=0.25, color='darkturquoise')
 ax.legend()

### 2024-08-07 src/scripts/plot_simulation_transformation.py

@@ -19,7 +19,7 @@ colors = ['tab:green', 'tab:red', 'tab:orange']
 for i, c in zip([np.argmin(abs(H0-40)), np.argmin(abs(H0-70)), np.argmin(abs(H0-100))], colors):
     ax2.plot(mz, model_pdf[:,i]/np.trapz(model_pdf[:,i], mz), label=f"$H_0={H0[i]:.0f}$", c=c)
 ax2.set_xlabel('$m^z_1\ [\mathrm{M}_\odot]$')
-ax2.set_ylabel('$p(m^z_1|\Lambda, \Omega(H_0), \mathrm{det})$')
+ax2.set_ylabel('$p(m^z_1|\Lambda,\Omega(H_0),\mathrm{det})$')
 ax2.set_xlim(0,200)
 ax2.set_ylim(0)
 ax2.legend()

### 2024-08-07 src/tex/ms.tex
[showyourwork] overleaf sync

@@
         The lower panel shows the transformed detector-frame $m^z_1$ population distributions for different $H_0$, where the corresponding $H_0$ is shown in the legend.
-        The source-frame population distribution is redshifted towards higher masses as $H_0$ increases, as $m^z_1 = m_1(1+z)$ and $z$ increases with $H_0$ for a given $d_L$.
+        The population distribution is redshifted towards higher masses as $H_0$ increases, as $m^z_1 = m_1(1+z)$ and $z$ increases with $H_0$ for a given $d_L$.
         The low-mass peak is suppressed due to the selection function.

### 2024-08-12 src/scripts/plot_simulation_comparison.py
Change legend

@@ -11,7 +11,7 @@ draws = load_density(outdir / f'draws/draws_observed_{label}.json')
 mz, H0, z, m, model_pdf = np.load(paths.data / 'grid.npz').values()
 bounds = np.loadtxt(outdir / "jsd_bounds.txt")
 
-fig = plot_median_cr(draws, hierarchical=True, save=True, show=False)
+fig = plot_median_cr(draws, hierarchical=True, save=True, show=False, median_label='$p(m^z_1|\mathbf{\Theta})$')
 ax = fig.axes[0]
 colors = ['tab:green', 'tab:red', 'tab:orange']
 for i, c in zip([np.argmin(abs(H0-40)), np.argmin(abs(H0-70)), np.argmin(abs(H0-100))], colors):

### 2024-08-12 src/scripts/plot_simulation_reconstruction.py

@@ -12,7 +12,7 @@ draws = load_density(outdir / f'draws/draws_observed_{label}.json')
 obs_samples = np.loadtxt(outdir / 'obs_samples.txt')
 
 fig, (ax1, ax2) = plt.subplots(2, figsize=(7,6), gridspec_kw={'hspace': 0.3})
-fig = plot_median_cr(draws, hierarchical=True, save=True, show=False, fig=fig)
+fig = plot_median_cr(draws, hierarchical=True, save=True, show=False, fig=fig, median_label='$p(m^z_1|\mathbf{\Theta})$')
 ax1.hist(obs_samples, bins = int(np.sqrt(len(obs_samples))), histtype = 'step', density = True, label = '$\mathrm{median}(Y_t)$', color = 'tab:red')
 ax1.set_xlim(0,200)
 ax1.set_ylim(0)
@@ -20,7 +20,7 @@ ax1.set_xlabel('$m^z_1\ [\mathrm{M}_\odot]$')
 ax1.set_ylabel('$\mathrm{Density}$')
 ax1.legend()
 m_z = np.linspace(0,200,1000)
-ax2.plot(m_z, draws[0].pdf(m_z), alpha = 0.5, color = 'tab:blue', label = '$\mathrm{(H)DPGMM}$')
+ax2.plot(m_z, draws[0].pdf(m_z), alpha = 0.5, color = 'tab:blue', label = '$p(m^z_1|\Theta_i)$')
 for i, draw in enumerate(draws[1:10]):
     ax2.plot(m_z, draw.pdf(m_z), alpha = 0.5, color = 'tab:blue')
 ax2.hist(obs_samples, bins = int(np.sqrt(len(obs_samples))), histtype = 'step', density = True, label = '$\mathrm{median}(Y_t)$', color = 'tab:red')

### 2024-08-12 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 This is because the reconstruction is only valid in the region where observations are present.
-This procedure gives us a set of optimized $\Lambda_i$ and $\Omega_i$ for each $\Theta_i$.
+As each reconstructed distribution is independent, the optimization can be performed in parallel easily, which reduces the runtime as long as computational resources are available.
 In Fig.~\ref{fig:simulation_comparison}, we show an illustration of the comparison between the two populations.
@@
 
-In this subsection, we first test our method with a simplified case where we only infer $H_0$.
+In this subsection, we first test our method with a simplified case where we only infer $H_0$ to demonstrate the basic idea of our method.
 We assume the selection function, the true population model and its parameters, and the true cosmological model and its parameters except $H_0$ are known.
@@
 For each $H_0$, we computed the $d_\mathrm{JS}$ between the two populations and found the optimized $H_0$ that minimizes $d_\mathrm{JS}$.
-The result of the \ac{PE} is shown in Fig.~\ref{fig:simulation_result_H0}.
 
@@
 \end{figure}
+
+The result of the \ac{PE} is shown in Fig.~\ref{fig:simulation_result_H0}.
+The result shows that our method can recover the true $H_0$ we used to generate the mock data.
 
@@
 Similar to the previous subsection, we assume only the free parameters are unknown.
+We chose the power law index $\alpha$, the peak mass $\mu$, the peak width $\sigma$, and the peak weight $w$ as the free parameters.
+Other parameters are the minimum and maximum mass, as well as the smoothing index of the boundaries.
+These parameters might not be able to be inferred accurately with our method, as the reconstruction is only valid in the region where the samples are present and the reconstruction may be not accurate at the boundaries.
 
@@
 Note that the same reconstruction result can be reused, as the data does not change.
+
+Following the same procedure as in the previous subsection, except that we use the \textsc{scipy} optimization algorithm mentioned in Sec.~\ref{sec:pe} in this example, we get the result of the \ac{PE} shown in Fig.~\todo{add figure}.
 
@@
 \label{sec:comparison}
-% diagnostic tool
-% model selection
+% diagnostic tool & model selection
 
@@
 S.~R.~ acknowledges financial support from the European Research Council for the ERC Consolidator grant DEMOBLACK, under contract no. 770017. 
-
 This research made use of the bwForCluster Helix: the authors acknowledge support by the state of Baden-Württemberg through bwHPC and the German Research Foundation (DFG) through grant INST 35/1597-1 FUGG.
-
 This research has made use of data or software obtained from the Gravitational Wave Open Science Center (gwosc.org), a service of the LIGO Scientific Collaboration, the Virgo Collaboration, and KAGRA.
@@
 KAGRA is supported by Ministry of Education, Culture, Sports, Science and Technology (MEXT), Japan Society for the Promotion of Science (JSPS) in Japan; National Research Foundation (NRF) and Ministry of Science and ICT (MSIT) in Korea; Academia Sinica (AS) and National Science and Technology Council (NSTC) in Taiwan.
-
 This paper was compiled using \textsc{showyourwork} \cite{Luger2021} to facilitate reproducibility.

### 2024-08-13 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \subsection{Inference of $H_0$}
+\label{sec:inference_H0}
 
@@
 \subsection{Inference of $H_0$ and $\Lambda$}
+\label{sec:inference_multi}
 
@@
 Following the same procedure as in the previous subsection, except that we use the \textsc{scipy} optimization algorithm mentioned in Sec.~\ref{sec:pe} in this example, we get the result of the \ac{PE} shown in Fig.~\todo{add figure}.
+The result shows that our method can infer multiple parameters simultaneously.
+
+In principle, more parameters can be inferred with our method, for example, the present mass density $\Omega_m$ in the $\Lambda$CDM model.
+However, the optimization procedure becomes more computationally expensive as the number of parameters increases.
 
@@
 
-\subsection{Future work}
-\label{sec:future_work}
+\subsection{Prospective improvements}
+\label{sec:prospective}
 As mentioned in Sec.~\ref{sec:pe}, our method is not a Bayesian inference, as we only consider the optimization of the parameters.
@@
 Furthermore, the inference of cosmological parameters will be more precise, as the information from the luminosity distance is also considered.
+
+We also mentioned a limitation of the method in Sec.~\ref{sec:inference_multi}.
+The runtime of the optimization procedure scales with the number of parameters.
+To perform higher dimensional analysis, we need to study using, instead of $d_\mathrm{JS}$, other distance measures that allow for faster optimization.
+Either a distance measure that gives a simpler geometry to be optimized or a distance measure that can be calculated faster will be useful.
+Other optimization algorithms should also be considered to allow for higher dimensional analysis and faster runtime.
 

### 2024-08-14 src/tex/ms.tex
[showyourwork] overleaf sync

@@
-\documentclass[aps,prd,twocolumn,superscriptaddress,preprintnumbers,nofootinbib,hidelinks]{revtex4-2}
+\documentclass[twocolumn]{aastex631}
 
@@
 \usepackage{bookmark}
-\usepackage{orcidlink}
-\usepackage{mathtools}
-\usepackage{array}
 
@@
 
-\author{Thomas~C.~K.~Ng\,\orcidlink{0000-0002-9491-1598}}
-\email{thomas.ng@link.cuhk.edu.hk}
+\author[0000-0002-9491-1598]{Thomas~C.~K.~Ng}
 \affiliation{Department of Physics, The Chinese University of Hong Kong, Shatin, Hong Kong}
 
-\author{Stefano~Rinaldi\,\orcidlink{0000-0001-5799-4155}}
-\email{stefano.rinaldi@uni-heidelberg.de}
+\author[0000-0001-5799-4155]{Stefano~Rinaldi}
 \affiliation{Institut~für~Theoretische~Astrophysik, ZAH, Universität~Heidelberg, Albert-Ueberle-Stra{\ss}e~2, 69120 Heidelberg, Germany}
@@
 
-\author{Otto~A.~Hannuksela\,\orcidlink{0000-0002-3887-7137}}
-\email{hannuksela@phy.cuhk.edu.hk}
+\author[0000-0002-3887-7137]{Otto~A.~Hannuksela}
 \affiliation{Department of Physics, The Chinese University of Hong Kong, Shatin, Hong Kong}
+
+\correspondingauthor{Thomas~C.~K.~Ng}
+\email{thomas.ng@link.cuhk.edu.hk}
 
@@
 
-\maketitle
+% \maketitle
 
@@
 In recent years, the precision of cosmological measurements has improved significantly, leading to the discovery of the Hubble tension, a discrepancy between the value of the Hubble constant $H_0$ inferred from the \ac{CMB} \citep{Planck:2018vyg} and local measurements \citep{Riess:2021jrx}.
-This tension has motivated the search for new methods to measure $H_0$ with high precision, and \ac{GW} sources have emerged as a promising tool for this purpose.
+This tension has motivated the search for new methods to measure $H_0$ with high precision, and \ac{GW} sources have emerged as a promising tool for this purpose \citep{LIGOScientific:2017adf, LIGOScientific:2021aug, Ezquiaga:2022zkx}.
+
 To infer $H_0$ from \ac{GW} sources, one needs not only the luminosity distance which can be inferred from the signal but also the redshift of the source.
@@
 However, most \ac{GW} sources do not have \ac{EM} counterparts.
-In this case, one can assume \ac{GW} sources are in galaxies and associate the redshift of the galaxy to the \ac{GW} source. \citep{Schutz:1986gp, DelPozzo:2011vcw, Gray:2019ksv, Gray:2023wgj}
+In this case, one can assume \ac{GW} sources are in galaxies and associate the redshift of the galaxy to the \ac{GW} source \citep{Schutz:1986gp, DelPozzo:2011vcw, Gray:2019ksv, Gray:2023wgj}.
 Another alternative method is to use population statistics to infer $H_0$ by marginalizing the redshift distribution of the sources.
-This method is known as the spectral siren method\citep{You:2020wju, Mastrogiovanni:2021wsd, LIGOScientific:2021aug, Ezquiaga:2022zkx}.
+This method is known as the spectral siren method \citep{You:2020wju, Mastrogiovanni:2021wsd, LIGOScientific:2021aug, Ezquiaga:2022zkx}.
 
@@
 % Section outline
-In Sec.~\ref{sec:method}, we describe our modified method and the testing procedure.
+In Sec.~\ref{sec:method}, we describe our modified method.
 In Sec.~\ref{sec:mock_data}, we present the analysis setup and results of the mock data study.
@@
 In the remainder of this Section, we will make use of the notation summarized in Table~\ref{tab:notation}, mutuated from \citet{Rinaldi:2021bhm, Rinaldi:2022kyg}.
-\todo{Renew notations}
 
@@
     \begin{ruledtabular}
-        \begin{tabular}{c>{\raggedright\arraybackslash}p{0.8\linewidth}}
+        \begin{tabular}{c>{\raggedright\arraybackslash}p{0.7\linewidth}}
             Notation & Description \\
@@
 The first step of our method is to reconstruct the detector-frame observed population.
-we choose to consider the primary mass $m_1$ of the \ac{BBH} sources, in principle, we can consider more parameters at once.
-From the posterior samples obtained from the \ac{PE} of each \ac{GW} event, we reconstruct the observed detector-frame primary mass population distribution.
+we choose to consider the primary mass $m_1$ of the \ac{BBH} sources, as we want to study the cosmological parameters with the spectral siren method.
+In principle, we can consider more parameters at once by reconstructing not only a single-parameter observed distribution but instead a multi-parameter observed distribution.
+However, that will require a more sophisticated population model in the later steps, which is beyond the scope of this work.
 
+With the posterior samples obtained from the \ac{PE} of each \ac{GW} event, we reconstruct the observed detector-frame primary mass population distribution.
 We use the \ac{(H)DPGMM} developed in \citet{Rinaldi:2021bhm} as a non-parametric model to reconstruct the observed $m^z_1$ population distribution $p(m^z_1|\mathbf{\Theta})$.
@@
 In this section, we test our method with a mock data study.
-To prepare the mock data, we first generated $10000$ sets of $m_1$ and $z$ samples from a \textsc{power law + peak} distribution (see, e.g., Eq.~B4 in \cite{KAGRA:2021duu}) for the $m_1$ distribution and uniform distribution over comoving volume and source-frame time (see, e.g., Eq.~11 in \cite{LIGOScientific:2019zcs}) for the $z$ distribution.
+To prepare the mock data, we first generated $10000$ sets of $m_1$ and $z$ samples from a \textsc{power law + peak} distribution (see, e.g., Eq.~B4 in \cite{KAGRA:2021duu}) for the $m_1$ distribution and uniform distribution over comoving volume and source-frame time (see, e.g., Eq.~11 in \citet{LIGOScientific:2019zcs}) for the $z$ distribution.
 Here we assumed the $m_1$ distribution does not evolve with $z$.
 We then transformed the $m_1$ and $z$ samples to $m^z_1$ and $d_L$ samples.
-Here we chose the $\Lambda$CDM model with parameters consistent with the Planck 2018 results \cite{Planck:2018vyg} as the cosmological model for the transformation of the samples.
+Here we chose the $\Lambda$CDM model with parameters consistent with the Planck 2018 results \citep{Planck:2018vyg} as the cosmological model for the transformation of the samples.
 

### 2024-08-15 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \usepackage{bookmark}
+\usepackage{xcolor}
+\usepackage{array}
+\usepackage{graphicx}
 
@@
 
-\author[0000-0002-9491-1598]{Thomas~C.~K.~Ng}
+\author{Thomas~C.~K.~Ng}
 \affiliation{Department of Physics, The Chinese University of Hong Kong, Shatin, Hong Kong}
 
-\author[0000-0001-5799-4155]{Stefano~Rinaldi}
+\author{Stefano~Rinaldi}
 \affiliation{Institut~für~Theoretische~Astrophysik, ZAH, Universität~Heidelberg, Albert-Ueberle-Stra{\ss}e~2, 69120 Heidelberg, Germany}
@@
 
-\author[0000-0002-3887-7137]{Otto~A.~Hannuksela}
+\author{Otto~A.~Hannuksela}
 \affiliation{Department of Physics, The Chinese University of Hong Kong, Shatin, Hong Kong}
@@
 In this section, we present the framework we developed to infer both the intrinsic population model and the cosmological model simultaneously by leveraging the detector-frame mass population distribution.
-In the remainder of this Section, we will make use of the notation summarized in Table~\ref{tab:notation}, mutuated from \citet{Rinaldi:2021bhm, Rinaldi:2022kyg}.
+In the remainder of this Section, we will make use of the notation summarized in Table~\ref{tab:notation}, mutated from \citet{Rinaldi:2021bhm, Rinaldi:2022kyg}.
 
@@
 Similar to the previous subsection, we assume only the free parameters are unknown.
-We chose the power law index $\alpha$, the peak mass $\mu$, the peak width $\sigma$, and the peak weight $w$ as the free parameters.
-Other parameters are the minimum and maximum mass, as well as the smoothing index of the boundaries.
-These parameters might not be able to be inferred accurately with our method, as the reconstruction is only valid in the region where the samples are present and the reconstruction may be not accurate at the boundaries.
+Other than $H_0$, we also chose the power law index $\alpha$, the peak mass $\mu$, the peak width $\sigma$, and the peak weight $w$ of the \textsc{power law + peak} model as the free parameters.
+Fixed parameters are the minimum and maximum mass, as well as the smoothing index of the boundaries.
+We may not be able to infer these fixed parameters accurately with our method, as the reconstruction is only valid in the region where the samples are present and the reconstruction may be not accurate at the boundaries depending on the boundary features of the observed population\footnote{As we are using a Gaussian mixture model to reconstruct the observed population, the reconstruction may not be accurate at the boundaries if the observed population has sharp features at the boundaries.}.
 
@@
 
-Data from \citet{LIGOScientific:2019lzm, KAGRA:2023pio}
+In this section, we apply our method to the real data from the \ac{LVK}.
+We use the data from \ac{GWTC-3} \citep{LIGOScientific:2019lzm, KAGRA:2023pio}.
 
@@
 This paper was compiled using \textsc{showyourwork} \cite{Luger2021} to facilitate reproducibility.
-
 \end{acknowledgments}

### 2024-08-16 src/tex/ms.tex
[showyourwork] overleaf sync

@@
     \acro{BBH}{binary black hole}
-    \acro{GWTC-3}{the third Gravitational-Wave Transient Catalog}
 \end{acronym}
@@
 \label{sec:method}
+
 In this section, we present the framework we developed to infer both the intrinsic population model and the cosmological model simultaneously by leveraging the detector-frame mass population distribution.
@@
 In this section, we test our method with a mock data study.
-To prepare the mock data, we first generated $10000$ sets of $m_1$ and $z$ samples from a \textsc{power law + peak} distribution (see, e.g., Eq.~B4 in \cite{KAGRA:2021duu}) for the $m_1$ distribution and uniform distribution over comoving volume and source-frame time (see, e.g., Eq.~11 in \citet{LIGOScientific:2019zcs}) for the $z$ distribution.
+To prepare the mock data, we first generated $10000$ sets of $m_1$ and $z$ samples from a \textsc{power law + peak} distribution (see, e.g., Eq.~B4 in \cite{KAGRA:2021duu}) for the $m_1$ distribution and uniform distribution over comoving volume and source-frame time (see, e.g., Eq.~8 in \cite{KAGRA:2021duu}) for the $z$ distribution.
 Here we assumed the $m_1$ distribution does not evolve with $z$.
 We then transformed the $m_1$ and $z$ samples to $m^z_1$ and $d_L$ samples.
-Here we chose the $\Lambda$CDM model with parameters consistent with the Planck 2018 results \citep{Planck:2018vyg} as the cosmological model for the transformation of the samples.
+Here we chose the $\Lambda$CDM model with parameters consistent with the results in \citet{Planck:2018vyg} as the cosmological model for the transformation of the samples.
 
@@
 Other than $H_0$, we also chose the power law index $\alpha$, the peak mass $\mu$, the peak width $\sigma$, and the peak weight $w$ of the \textsc{power law + peak} model as the free parameters.
-Fixed parameters are the minimum and maximum mass, as well as the smoothing index of the boundaries.
+Fixed parameters are the minimum and maximum mass, as well as the range of mass tapering at the low mass end.
 We may not be able to infer these fixed parameters accurately with our method, as the reconstruction is only valid in the region where the samples are present and the reconstruction may be not accurate at the boundaries depending on the boundary features of the observed population\footnote{As we are using a Gaussian mixture model to reconstruct the observed population, the reconstruction may not be accurate at the boundaries if the observed population has sharp features at the boundaries.}.
@@
 
-Following the same procedure as in the previous subsection, except that we use the \textsc{scipy} optimization algorithm mentioned in Sec.~\ref{sec:pe} in this example, we get the result of the \ac{PE} shown in Fig.~\todo{add figure}.
+In this example, instead of transforming the source-frame population model to the detector frame and performing a grid search, we define a function that calculates $d_\mathrm{JS}$ for a given set of the free parameters.
+We then use the \textsc{scipy} optimization algorithm mentioned in Sec.~\ref{sec:pe} to find the optimized free parameters that minimize $d_\mathrm{JS}$.
+The result of the \ac{PE} is shown in Fig.~\todo{add figure}.
 The result shows that our method can infer multiple parameters simultaneously.
@@
 In this section, we apply our method to the real data from the \ac{LVK}.
-We use the data from \ac{GWTC-3} \citep{LIGOScientific:2019lzm, KAGRA:2023pio}.
+We included the \ac{GW} events considered in \citet{KAGRA:2021duu}.
 
-% Power law + peak
+We first performed the reconstruction of the observed $m^z_1$ population distribution with \textsc{figaro} as described in Sec.~\ref{sec:reconstruction}.
+We use the posterior samples of the \ac{GW} events from \citet{LIGOScientific:2019lzm, KAGRA:2023pio}.
+The result of the reconstruction is shown in Fig.~\todo{add figure}.
+
+Next, we chose the \textsc{power law + peak} model as the source-frame population model $\Lambda$ and the $\Lambda$CDM model as the cosmological model $\Omega$.
+Similar to Sec.~\ref{sec:inference_multi}, we fixed some parameters of the population model and the cosmological model.
+For the cosmological model, we fixed the parameters to the results from \citet{Planck:2018vyg} except $H_0$.
+For the population model, we fixed the minimum and maximum mass, as well as the range of mass tapering at the low mass end to the results from \citet{KAGRA:2021duu}.
+Note that the fixed parameters are not necessarily the true values\footnote{Result from \citet{KAGRA:2021duu} assumes a cosmological model to infer the population model parameters.}, but they can be considered as a reasonable guess for the parameters.
+The fixed parameters are shown in Table~\todo{add table}.
+
+We then perform the remaining steps similar to Sec.~\ref{sec:inference_multi}.
+The result of the \ac{PE} is shown in Fig.~\todo{add figure} and Fig.~\todo{add figure}.
 
@@
 \label{sec:prospective}
+
 As mentioned in Sec.~\ref{sec:pe}, our method is not a Bayesian inference, as we only consider the optimization of the parameters.
@@
 \begin{acknowledgments}
+
 T.~N.~ and S.~R.~ acknowledge financial support from the German Excellence Strategy via the Heidelberg Cluster of Excellence (EXC 2181 - 390900948) STRUCTURES.
@@
 This paper was compiled using \textsc{showyourwork} \cite{Luger2021} to facilitate reproducibility.
+
 \end{acknowledgments}

### 2024-08-30 src/tex/ms.tex
[showyourwork] overleaf sync

@@
     \acro{(H)DPGMM}{hierarchy of Dirichlet process Gaussian mixture models}
-    \acro{BBH}{binary black hole}
+    \acro{BBH}{binary-black-hole}
 \end{acronym}
@@
 Non-parametric methods allow us to reconstruct arbitrary probability densities without making assumptions about the form of the distribution.
-In our context, we can use non-parametric methods to reconstruct the observed mass population distribution directly from the data \todo{better wording?}.
+In our context, we can use non-parametric methods to reconstruct the observed mass population distribution directly from the data.
 This provides an intermediate observation-driven result that encodes both the intrinsic population information and the cosmological information.
@@
 % Section outline
-In Sec.~\ref{sec:method}, we describe our modified method.
+In Sec.~\ref{sec:method}, we describe our modified method in detail.
 In Sec.~\ref{sec:mock_data}, we present the analysis setup and results of the mock data study.
@@
             $m^z_1$ & Primary mass of the \ac{BBH} sources in the detector frame \\
-            $Y_t$ & detector-frame primary mass posterior samples of the $t$-th \ac{GW} event \\
-            $\mathbf{Y}$ & Set of detector-frame primary mass posterior samples of all \ac{GW} events \\
+            $Y_t$ & detector-frame primary mass posterior probability distribution samples of the $t$-th \ac{GW} event \\
+            $\mathbf{Y}$ & Set of detector-frame primary mass posterior probability distribution samples of all \ac{GW} events \\
             $\Theta_i$ & Hyperparameters of the \ac{(H)DPGMM} \\
@@
 Specifically, we propose two major modifications to the standard spectral siren method.
-First, instead of transforming the individual-event posterior samples from the detector frame to the source frame during the \ac{PE} process, we transform the population model to the detector frame before performing \ac{PE}.
+First, instead of transforming the individual-event posterior probability distribution samples from the detector frame to the source frame during the \ac{PE} process, we transform the population model to the detector frame before performing \ac{PE}.
 Second, instead of directly computing the likelihood of the observed data given the population model, we reconstruct the observed population distribution from the data using a non-parametric method and compare it with the population model.
@@
 
-With the posterior samples obtained from the \ac{PE} of each \ac{GW} event, we reconstruct the observed detector-frame primary mass population distribution.
+With the posterior probability distribution samples obtained from the \ac{PE} of each \ac{GW} event, we reconstruct the observed detector-frame primary mass population distribution.
 We use the \ac{(H)DPGMM} developed in \citet{Rinaldi:2021bhm} as a non-parametric model to reconstruct the observed $m^z_1$ population distribution $p(m^z_1|\mathbf{\Theta})$.
-The \ac{(H)DPGMM} is a hierarchy of Gaussian mixture models with a Dirichlet process prior, which allows the model to be flexible to reconstruct arbitrary population distribution from posterior samples.
+The \ac{(H)DPGMM} is a hierarchy of Gaussian mixture models with a Dirichlet process prior probability distribution, which allows the model to be flexible to reconstruct arbitrary population distribution from individual-event posterior probability distribution samples.
 To perform the reconstruction, we use the package \textsc{figaro}\footnote{\textsc{figaro} is publicly available at \url{https://github.com/sterinaldi/FIGARO} and via \texttt{pip}.} \citep{Rinaldi:2024eep}.
@@
 For the minimization, we use the \textsc{scipy} \citep{2020SciPy-NMeth} implementation of the \todo{XXX} algorithm. \todo{Add reference for the algorithm}
-Note that we calculate $d_\mathrm{JS}$ in the range from the minimum to the maximum of the median of the $m^z_1$ posterior samples of each \ac{GW} event
+Note that we calculate $d_\mathrm{JS}$ in the range from the minimum to the maximum of the median of the $m^z_1$ posterior probability distribution samples of each \ac{GW} event
 This is because the reconstruction is only valid in the region where observations are present.
@@
 
-The resulting distribution of the optimized parameters is not a posterior distribution in a Bayesian sense, as we do not consider the likelihood of the observed data given the population model.
+The resulting distribution of the optimized parameters is not a posterior probability distribution in a Bayesian sense, as we do not consider the likelihood of the observed data given the population model.
 Instead, it is a distribution of the optimized parameters that are consistent with the observed data, distributed according to the uncertainty in the reconstruction of the population.
@@
 In this section, we test our method with a mock data study.
-To prepare the mock data, we first generated $10000$ sets of $m_1$ and $z$ samples from a \textsc{power law + peak} distribution (see, e.g., Eq.~B4 in \cite{KAGRA:2021duu}) for the $m_1$ distribution and uniform distribution over comoving volume and source-frame time (see, e.g., Eq.~8 in \cite{KAGRA:2021duu}) for the $z$ distribution.
+To prepare the mock data, we first generated \variable{output/n_true_samples.txt} sets of $m_1$ and $z$ samples from a \textsc{power law + peak} distribution (see, e.g., Eq.~B4 in \cite{KAGRA:2021duu}) for the $m_1$ distribution and uniform distribution over comoving volume and source-frame time (see, e.g., Eq.~8 in \cite{KAGRA:2021duu}) for the $z$ distribution.
 Here we assumed the $m_1$ distribution does not evolve with $z$.
@@
 Note that we marginalized over $q$ for the selection function, similar to the transformation in Sec.~\ref{sec:transformation}.
-After applying the selection function, there were $262$ sets of samples left.
+After applying the selection function, there were \variable{output/n_obs_samples.txt} sets of samples left.
 
-We then generated the $m^z_1$ posterior samples for each true $m^z_1$ sample by
+We then generated the $m^z_1$ posterior probability distribution samples for each true $m^z_1$ sample by
 \begin{gather*}
@@
 \end{gather*}
-where $\mathcal{N}(\mu, \sigma)$ is a normal distribution with mean $\mu$ and standard deviation $\sigma$, $I$ is the difference between the true $m^z_1$ value and the measured mean in log space, $P$ is the posterior sample in the log space, $T_t$ is the true $m^z_1$ sample, and $Y_t$ is the posterior sample for the $t$-th \ac{GW} event.
+where $\mathcal{N}(\mu, \sigma)$ is a normal distribution with mean $\mu$ and standard deviation $\sigma$, $I$ is the difference between the true $m^z_1$ value and the measured mean in log space, $P$ is the posterior probability distribution sample in the log space, $T_t$ is the true $m^z_1$ sample, and $Y_t$ is the posterior probability distribution sample for the $t$-th \ac{GW} event.
 For each true $m^z_1$ sample, we draw a random $I$ and generate $1000$ $Y_t$ samples.
-This procedure ensures the posterior samples are positive.
+This procedure ensures the posterior probability distribution samples are positive.
 After preparing the mock data, we test our method with two cases: the inference of $H_0$ only and the inference of $H_0$ and a subset of the population model parameters.
@@
     \caption{
-        In both panels, the red line is the histogram of the median of the $m^z_1$ posterior samples of the \ac{GW} events.
+        In both panels, the red line is the histogram of the median of the $m^z_1$ posterior probability distribution samples of the \ac{GW} events.
         In the upper panel, the blue line is the median of the reconstructed distributions, and the shaded region represents the 68\% and 90\% credible intervals of the reconstructed distributions.
@@
 We then use the \textsc{scipy} optimization algorithm mentioned in Sec.~\ref{sec:pe} to find the optimized free parameters that minimize $d_\mathrm{JS}$.
+\todo{Add more details about the optimization procedure (e.g., initial guess, bounds, etc.)}
 The result of the \ac{PE} is shown in Fig.~\todo{add figure}.
@@
 In this section, we apply our method to the real data from the \ac{LVK}.
-We included the \ac{GW} events considered in \citet{KAGRA:2021duu}.
+We included all events labeled as BBH in Table~I of \citet{KAGRA:2021duu}.
 
 We first performed the reconstruction of the observed $m^z_1$ population distribution with \textsc{figaro} as described in Sec.~\ref{sec:reconstruction}.
-We use the posterior samples of the \ac{GW} events from \citet{LIGOScientific:2019lzm, KAGRA:2023pio}.
+We use the posterior probability distribution samples of the \ac{GW} events from \citet{LIGOScientific:2019lzm, KAGRA:2023pio}.
 The result of the reconstruction is shown in Fig.~\todo{add figure}.
@@
 The fixed parameters are shown in Table~\todo{add table}.
+\todo{Add more details about the selection function}
+\todo{Consider using values from another paper}
 
@@
 The result of the \ac{PE} is shown in Fig.~\todo{add figure} and Fig.~\todo{add figure}.
+\todo{Discuss the result}
 
-% PL+P issue
-    % Gregoire Pierra's paper
+The result shows similar problems mentioned in \citet{Pierra:2024wrk}.
+\todo{Discuss mass-redshift dependent population model \& possible issues with PL+P model}
 

### 2024-09-02 src/scripts/generate_posterior_samples.py
Update scripts

@@ -5,7 +5,7 @@ import os
 import dill
 import paths
 
-np.random.default_rng(12345)
+np.random.default_rng(1234)
 
 # Mass distribution
 from population_models.mass import plpeak # from https://github.com/thomasckng/pop_models_lvk

### 2024-09-02 src/scripts/infer_multi.py

@@ -23,39 +23,42 @@ if not os.path.exists(outdir/f'multi/{sys.argv[1]}_{sys.argv[2]}.npz'):
         print("Invalid number of arguments!")
         sys.exit(1)
 
+    bounds_dict = {
+        "H0": (10, 300),
+        "alpha": (1.01, 10),
+        "mu": (10, 70),
+        "sigma": (0.01, 10),
+        "w": (0, 1),
+        "delta": (0.01, 20),
+        "mmin": (1, 10)
+    }
+
     if sys.argv[1] == "2a":
-        x0 = [uni(10,200), uni(1.01,5)]
-        bounds = ((10,200), (1.01,5))
+        bounds = bounds_dict["H0"], bounds_dict["alpha"]
         def plp(m, x):
             return plpeak(m, alpha=x[0])
     elif sys.argv[1] == "2b":
-        x0 = [uni(10,200), uni(10,50)]
-        bounds = ((10,200), (10,50))
+        bounds = bounds_dict["H0"], bounds_dict["mu"]
         def plp(m, x):
             return plpeak(m, mu=x[0])
     elif sys.argv[1] == "4a":
-        x0 = [uni(10,200), uni(10,50), uni(0.01,10), uni(0.01,15)]
-        bounds = ((10,200), (10,50), (0.01,10), (0.01,15))
+        bounds = bounds_dict["H0"], bounds_dict["mu"], bounds_dict["sigma"], bounds_dict["delta"]
         def plp(m, x):
             return plpeak(m, mu=x[0], sigma=x[1], delta=x[2])
     elif sys.argv[1] == "4b":
-        x0 = [uni(10,200), uni(1.01,5), uni(10,50), uni(0.01,10)]
-        bounds = ((10,200), (1.01,5), (10,50), (0.01,10))
+        bounds = bounds_dict["H0"], bounds_dict["alpha"], bounds_dict["mu"], bounds_dict["sigma"]
         def plp(m, x):
             return plpeak(m, alpha=x[0], mu=x[1], sigma=x[2])
     elif sys.argv[1] == "5a":
-        x0 = [uni(10,200), uni(1.01,5), uni(10,50), uni(0.01,10), uni(0,1)]
-        bounds = ((10,200), (1.01,5), (10,50), (0.01,10), (0,1))
+        bounds = bounds_dict["H0"], bounds_dict["alpha"], bounds_dict["mu"], bounds_dict["sigma"], bounds_dict["w"]
         def plp(m, x):
[…]

### 2024-09-02 src/scripts/plot_simulation_result_H0.py

@@ -12,10 +12,10 @@ ax.hist(H0_samples, bins = int(np.sqrt(len(H0_samples))), histtype = 'step', den
 percs = np.percentile(H0_samples, [5, 16, 50, 84, 95])
 ax.axvline(true_H0, lw = 0.7, ls = '--', c = 'orangered', label = f'$H_0={true_H0:.1f}\ (\mathrm{{Simulated}})$')
 ax.axvline(percs[2], c = 'steelblue', lw=0.7, label = f'$H_0={percs[2]:.1f}^{{+{percs[3]-percs[2]:.1f}}}_{{-{percs[2]-percs[1]:.1f}}}$')
-ax.axvspan(percs[1], percs[3], alpha=0.25, color='mediumturquoise')
-ax.axvspan(percs[0], percs[4], alpha=0.25, color='darkturquoise')
+ax.axvspan(percs[1], percs[3], alpha=0.4, color='mediumturquoise')
+ax.axvspan(percs[0], percs[4], alpha=0.2, color='darkturquoise')
 ax.legend()
 ax.set_xlabel('$H_0\ [\mathrm{km/s/Mpc}]$')
 ax.set_ylabel('$\mathrm{Density}$')
-ax.set_xlim(true_H0-20, true_H0+20)
+ax.set_xlim(percs[2]-30, percs[2]+30)
 fig.savefig(paths.figures / "simulation_result_H0.pdf", bbox_inches='tight')

### 2024-09-02 src/scripts/generate_posterior_samples.py
Add real event plots

@@ -31,7 +31,7 @@ with open(paths.data/'selection_function.pkl', 'rb') as f: # selection_function.
 def selection_function(x):
     return selfunc_interp(x)
 
-n_single_event_draws = 20_000
+n_single_event_draws = 100_000
 outdir = 'simulation'
 if not os.path.exists(paths.data / outdir):
     os.makedirs(paths.data / outdir)

### 2024-09-02 src/scripts/get_lvk_results.py

@@ -99,6 +99,7 @@ events = [\
 
 mass_med_min, mass_med_max = 0, 0
 mass_min_min, mass_max_max = 0, 0
+med_all = []
 for event in tqdm.tqdm(events):
     h5_paths = find('*' + str(event) + '*' + 'nocosmo.h5', sys.argv[1])
     if len(h5_paths) == 0:
@@ -111,10 +112,13 @@ for event in tqdm.tqdm(events):
             f = f['C01:IMRPhenomXPHM']
             result_lvk = pd.DataFrame.from_records(f["posterior_samples"][()])
             result_lvk = result_lvk.sample(n=n_samples, random_state=seed)
-        mass_med_min = min(mass_med_min, np.median(result_lvk['mass_1']))
-        mass_med_max = max(mass_med_max, np.median(result_lvk['mass_1']))
+        event_med = np.median(result_lvk['mass_1'])
+        med_all.append(event_med)
+        mass_med_min = min(mass_med_min, event_med)
+        mass_med_max = max(mass_med_max, event_med)
         mass_min_min = min(mass_min_min, min(result_lvk['mass_1']))
         mass_max_max = max(mass_max_max, max(result_lvk['mass_1']))
         np.savetxt(paths.data/f'real/data/{event}.txt', result_lvk['mass_1'])
+np.savetxt(paths.data/'real/samples_med.txt', med_all)
 np.savetxt(paths.data/'real/jsd_bounds.txt', [mass_med_min, mass_med_max])
 print(f"Mass sample bounds: {mass_min_min}, {mass_max_max}")

### 2024-09-02 src/scripts/plot_real_reconstruction.py

@@ -5,15 +5,15 @@ import numpy as np
 from matplotlib import pyplot as plt
 import paths
 
-label = 'simulation'
+label = 'real'
 outdir = paths.data / label
 
 draws = load_density(outdir / f'draws/draws_observed_{label}.json')
-obs_samples = np.loadtxt(outdir / 'obs_samples.txt')
+samples_med = np.loadtxt(outdir / 'samples_med.txt')
 
 fig, (ax1, ax2) = plt.subplots(2, figsize=(7,6), gridspec_kw={'hspace': 0.3})
 fig = plot_median_cr(draws, hierarchical=True, save=True, show=False, fig=fig, median_label='$p(m^z_1|\mathbf{\Theta})$')
-ax1.hist(obs_samples, bins = int(np.sqrt(len(obs_samples))), histtype = 'step', density = True, label = '$T_t$', color = 'tab:red')
+ax1.hist(samples_med, bins = int(np.sqrt(len(samples_med))), histtype = 'step', density = True, label = '$\mathrm{median}(Y_t)$', color = 'tab:red')
 ax1.set_xlim(0,200)
 ax1.set_ylim(0)
 ax1.set_xlabel('$m^z_1\ [\mathrm{M}_\odot]$')
@@ -23,10 +23,10 @@ m_z = np.linspace(0,200,1000)
 ax2.plot(m_z, draws[0].pdf(m_z), alpha = 0.5, color = 'tab:blue', label = '$p(m^z_1|\Theta_i)$')
 for i, draw in enumerate(draws[1:10]):
     ax2.plot(m_z, draw.pdf(m_z), alpha = 0.5, color = 'tab:blue')
-ax2.hist(obs_samples, bins = int(np.sqrt(len(obs_samples))), histtype = 'step', density = True, label = '$T_t$', color = 'tab:red')
+ax2.hist(samples_med, bins = int(np.sqrt(len(samples_med))), histtype = 'step', density = True, label = '$\mathrm{median}(Y_t)$', color = 'tab:red')
 ax2.set_xlim(0,200)
 ax2.set_ylim(0)
 ax2.set_xlabel('$m^z_1\ [\mathrm{M}_\odot]$')
 ax2.set_ylabel('$\mathrm{Density}$')
 ax2.legend()
-fig.savefig(paths.figures / 'simulation_reconstruction.pdf', bbox_inches='tight')
+fig.savefig(paths.figures / 'real_reconstruction.pdf', bbox_inches='tight')

### 2024-09-04 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 
-% Motivation
 In recent years, the precision of cosmological measurements has improved significantly, leading to the discovery of the Hubble tension, a discrepancy between the value of the Hubble constant $H_0$ inferred from the \ac{CMB} \citep{Planck:2018vyg} and local measurements \citep{Riess:2021jrx}.
@@
 
-% Previous work
 The fundamental idea of the spectral siren method is to infer cosmology by comparing the observed mass population distribution with the intrinsic mass population distribution since the observed distribution is redshifted due to the expansion of the Universe.
@@
 
-% Non-parametric methods
 To address this issue, we study an alternative method making use of non-parametric methods.
@@
 
-% Section outline
 In Sec.~\ref{sec:method}, we describe our modified method in detail.
@@
 
-\begin{table}
+\begin{table}[h]
     \caption{Notation}
@@
     \caption{
-        In both panels, the red line is the histogram of the median of the $m^z_1$ posterior probability distribution samples of the \ac{GW} events.
+        In both panels, the red line is the histogram of the true $m^z_1$ of the \ac{GW} events.
         In the upper panel, the blue line is the median of the reconstructed distributions, and the shaded region represents the 68\% and 90\% credible intervals of the reconstructed distributions.
@@
 
-In this example, instead of transforming the source-frame population model to the detector frame and performing a grid search, we define a function that calculates $d_\mathrm{JS}$ for a given set of the free parameters.
+In this example, instead of transforming the source-frame population model to the detector frame for each $H_0$ and performing a grid search, we define a function that calculates $d_\mathrm{JS}$ for a given set of the free parameters.
 We then use the \textsc{scipy} optimization algorithm mentioned in Sec.~\ref{sec:pe} to find the optimized free parameters that minimize $d_\mathrm{JS}$.
-\todo{Add more details about the optimization procedure (e.g., initial guess, bounds, etc.)}
+The optimization method requires bounds for the free parameters, which we show in Table~\ref{tab:bounds}.
+\begin{table}[h]
+    \caption{Bounds for the free parameters}
+    \begin{ruledtabular}
+        \begin{tabular}{ccc}
+            Parameter & Symbol & Bounds \\
+            \hline
+            Hubble constant & $H_0$ & $[10, 300]$ \\
+            Power law index & $\alpha$ & $[1.01, 10]$ \\
+            Peak mass & $\mu$ & $[10, 70]$ \\
+            Peak width & $\sigma$ & $[0.01, 10]$ \\
+            Peak weight & $w$ & $[0, 1]$ \\
+        \end{tabular}
+    \end{ruledtabular}
+    \label{tab:bounds}
+\end{table}
+The initial guesses for the optimization are chosen randomly from a uniform distribution within the bounds.
 The result of the \ac{PE} is shown in Fig.~\todo{add figure}.
@@
 We use the posterior probability distribution samples of the \ac{GW} events from \citet{LIGOScientific:2019lzm, KAGRA:2023pio}.
-The result of the reconstruction is shown in Fig.~\todo{add figure}.
+$10000$ samples are randomly drawn from the posterior probability distribution samples of each \ac{GW} event to perform the reconstruction.
+The result of the reconstruction is shown in Fig.~\ref{fig:real_reconstruction}.
+
+\begin{figure}[h]
+    \script{plot_real_reconstruction.py}
+    \includegraphics[width=\linewidth]{figures/real_reconstruction.pdf}
+    \caption{
+        The red line is the histogram of the median of the $m^z_1$ posterior probability distribution samples of the \ac{GW} events.
+        The blue line is the median of the reconstructed distributions, and the shaded region represents the 68\% and 90\% credible intervals of the reconstructed distributions.
+    }
+    \label{fig:real_reconstruction}
+\end{figure}
 
@@
 \label{sec:comparison}
-% diagnostic tool & model selection
 
@@
 With this likelihood, our method will give the same result as the standard hierarchical Bayesian inference, as long as the number of reconstructions is large enough.
+This will allow us to perform efficient model selection, as different models can be tested without repeating the reconstruction of the observed population.
 
@@
 \label{sec:conclusion}
+
+\todo{Work as diagnostic tool}
 

### 2024-09-05 src/scripts/generate_posterior_samples.py
Modify multi PE

@@ -5,7 +5,7 @@ import os
 import dill
 import paths
 
-np.random.default_rng(1234)
+np.random.default_rng(42)
 
 # Mass distribution
 from population_models.mass import plpeak # from https://github.com/thomasckng/pop_models_lvk
@@ -31,7 +31,7 @@ with open(paths.data/'selection_function.pkl', 'rb') as f: # selection_function.
 def selection_function(x):
     return selfunc_interp(x)
 
-n_single_event_draws = 100_000
+n_single_event_draws = 20_000
 outdir = 'simulation'
 if not os.path.exists(paths.data / outdir):
     os.makedirs(paths.data / outdir)

### 2024-09-05 src/scripts/infer_multi.py

@@ -5,7 +5,14 @@ import paths
 label = sys.argv[3]
 outdir = paths.data / label
 
-if not os.path.exists(outdir/f'multi/{sys.argv[1]}_{sys.argv[2]}.npz'):
+if len(sys.argv) != 5:
+    print("Invalid number of arguments!")
+    sys.exit(1)
+
+param = sys.argv[1]
+method = sys.argv[2]
+
+if not os.path.exists(outdir/f'multi/{param}_{method}.npz'):
     import numpy as np
     from numpy.random import uniform as uni
     from scipy.spatial.distance import jensenshannon as scipy_jsd
@@ -19,10 +26,6 @@ if not os.path.exists(outdir/f'multi/{sys.argv[1]}_{sys.argv[2]}.npz'):
     def p_z(z, H0):
         return CosmologicalParameters(H0/100., 0.315, 0.685, -1., 0., 0.).ComovingVolumeElement(z)/(1+z)
 
-    if len(sys.argv) != 5:
-        print("Invalid number of arguments!")
-        sys.exit(1)
-
     bounds_dict = {
         "H0": (10, 300),
         "alpha": (1.01, 10),
@@ -33,31 +36,31 @@ if not os.path.exists(outdir/f'multi/{sys.argv[1]}_{sys.argv[2]}.npz'):
         "mmin": (1, 10)
     }
 
-    if sys.argv[1] == "2a":
+    if param == "2a":
         bounds = bounds_dict["H0"], bounds_dict["alpha"]
         def plp(m, x):
             return plpeak(m, alpha=x[0])
-    elif sys.argv[1] == "2b":
+    elif param == "2b":
         bounds = bounds_dict["H0"], bounds_dict["mu"]
         def plp(m, x):
             return plpeak(m, mu=x[0])
-    elif sys.argv[1] == "4a":
+    elif param == "4a":
         bounds = bounds_dict["H0"], bounds_dict["mu"], bounds_dict["sigma"], bounds_dict["delta"]
         def plp(m, x):
             return plpeak(m, mu=x[0], sigma=x[1], delta=x[2])
-    elif sys.argv[1] == "4b":
+    elif param == "4b":
         bounds = bounds_dict["H0"], bounds_dict["alpha"], bounds_dict["mu"], bounds_dict["sigma"]
         def plp(m, x):
             return plpeak(m, alpha=x[0], mu=x[1], sigma=x[2])
[…]

### 2024-09-10 src/scripts/infer_multi.py
Add 8 param inference

@@ -33,7 +33,8 @@ if not os.path.exists(outdir/f'multi/{param}_{method}.npz'):
         "sigma": (0.01, 10),
         "w": (0, 1),
         "delta": (0.01, 20),
-        "mmin": (1, 10)
+        "mmin": (1, 10),
+        "mmax": (70, 150)
     }
 
     if param == "2a":
@@ -64,6 +65,10 @@ if not os.path.exists(outdir/f'multi/{param}_{method}.npz'):
         bounds = bounds_dict["H0"], bounds_dict["alpha"], bounds_dict["mu"], bounds_dict["sigma"], bounds_dict["w"], bounds_dict["delta"]
         def plp(m, x):
             return plpeak(m, alpha=x[0], mu=x[1], sigma=x[2], w=x[3], delta=x[4])
+    elif param == "8":
+        bounds = bounds_dict["H0"], bounds_dict["alpha"], bounds_dict["mu"], bounds_dict["sigma"], bounds_dict["w"], bounds_dict["delta"], bounds_dict["mmin"], bounds_dict["mmax"]
+        def plp(m, x):
+            return plpeak(m, alpha=x[0], mu=x[1], sigma=x[2], w=x[3], delta=x[4], mmin=x[5], mmax=x[6])
     else:
         print("Invalid argument!")
         sys.exit(1)

### 2024-09-10 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 For each reconstructed distribution $p(m^z_1|\Theta_i)$, we find an optimized $\Lambda_i$ and $\Omega_i$ by minimizing $d_\mathrm{JS}$.
-For the minimization, we use the \textsc{scipy} \citep{2020SciPy-NMeth} implementation of the \todo{XXX} algorithm. \todo{Add reference for the algorithm}
+For the minimization, we use the \textsc{scipy} \citep{2020SciPy-NMeth} implementation of the \todo{XXX (add reference)} algorithm.
 Note that we calculate $d_\mathrm{JS}$ in the range from the minimum to the maximum of the median of the $m^z_1$ posterior probability distribution samples of each \ac{GW} event
@@
     \caption{
-        In both panels, the red line is the histogram of the true $m^z_1$ of the \ac{GW} events.
+        In both panels, the red line is the histogram of the true $m^z_1$ of the simulated \ac{GW} events.
         In the upper panel, the blue line is the median of the reconstructed distributions, and the shaded region represents the 68\% and 90\% credible intervals of the reconstructed distributions.
@@
 In this section, we apply our method to the real data from the \ac{LVK}.
-We included all events labeled as BBH in Table~I of \citet{KAGRA:2021duu}.
+We included all $70$ events labeled as BBH in Table~I of \citet{KAGRA:2021duu}.
 
@@
 \label{sec:discussion}
-
-\subsection{Comparison with previous work}
-\label{sec:comparison}
-
-\subsection{Prospective improvements}
-\label{sec:prospective}
 

### 2024-09-13 src/scripts/plot_real_reconstruction.py
Add multi dim plot

@@ -20,9 +20,9 @@ ax1.set_xlabel('$m^z_1\ [\mathrm{M}_\odot]$')
 ax1.set_ylabel('$\mathrm{Density}$')
 ax1.legend()
 m_z = np.linspace(0,200,1000)
-ax2.plot(m_z, draws[0].pdf(m_z), alpha = 0.5, color = 'tab:blue', label = '$p(m^z_1|\Theta_i)$')
+ax2.plot(m_z, draws[0].pdf(m_z), alpha = 0.5, color = 'steelblue', label = '$p(m^z_1|\Theta_i)$')
 for i, draw in enumerate(draws[1:10]):
-    ax2.plot(m_z, draw.pdf(m_z), alpha = 0.5, color = 'tab:blue')
+    ax2.plot(m_z, draw.pdf(m_z), alpha = 0.5, color = 'steelblue')
 ax2.hist(samples_med, bins = int(np.sqrt(len(samples_med))), histtype = 'step', density = True, label = '$\mathrm{median}(Y_t)$', color = 'tab:red')
 ax2.set_xlim(0,200)
 ax2.set_ylim(0)

### 2024-09-13 src/scripts/plot_simulation_reconstruction.py

@@ -13,7 +13,7 @@ obs_samples = np.loadtxt(outdir / 'obs_samples.txt')
 
 fig, (ax1, ax2) = plt.subplots(2, figsize=(7,6), gridspec_kw={'hspace': 0.3})
 fig = plot_median_cr(draws, hierarchical=True, save=True, show=False, fig=fig, median_label='$p(m^z_1|\mathbf{\Theta})$')
-ax1.hist(obs_samples, bins = int(np.sqrt(len(obs_samples))), histtype = 'step', density = True, label = '$T_t$', color = 'tab:red')
+ax1.hist(obs_samples, bins = int(np.sqrt(len(obs_samples))), histtype = 'step', density = True, label = '$T_t$', color = 'red')
 ax1.set_xlim(0,200)
 ax1.set_ylim(0)
 ax1.set_xlabel('$m^z_1\ [\mathrm{M}_\odot]$')
@@ -23,7 +23,7 @@ m_z = np.linspace(0,200,1000)
 ax2.plot(m_z, draws[0].pdf(m_z), alpha = 0.5, color = 'tab:blue', label = '$p(m^z_1|\Theta_i)$')
 for i, draw in enumerate(draws[1:10]):
     ax2.plot(m_z, draw.pdf(m_z), alpha = 0.5, color = 'tab:blue')
-ax2.hist(obs_samples, bins = int(np.sqrt(len(obs_samples))), histtype = 'step', density = True, label = '$T_t$', color = 'tab:red')
+ax2.hist(obs_samples, bins = int(np.sqrt(len(obs_samples))), histtype = 'step', density = True, label = '$T_t$', color = 'red')
 ax2.set_xlim(0,200)
 ax2.set_ylim(0)
 ax2.set_xlabel('$m^z_1\ [\mathrm{M}_\odot]$')

### 2024-09-13 src/scripts/plot_simulation_result_H0.py

@@ -10,9 +10,9 @@ true_H0 = Planck18.h*100
 fig, ax = plt.subplots()
 ax.hist(H0_samples, bins = int(np.sqrt(len(H0_samples))), histtype = 'step', density = True, color = 'tab:blue')
 percs = np.percentile(H0_samples, [5, 16, 50, 84, 95])
-ax.axvline(true_H0, lw = 0.7, ls = '--', c = 'orangered', label = f'$H_0={true_H0:.1f}\ (\mathrm{{Simulated}})$')
+ax.axvline(true_H0, lw = 0.7, ls = '--', c = 'red', label = f'$H_0={true_H0:.1f}\ (\mathrm{{Simulated}})$')
 ax.axvline(percs[2], c = 'steelblue', lw=0.7, label = f'$H_0={percs[2]:.1f}^{{+{percs[3]-percs[2]:.1f}}}_{{-{percs[2]-percs[1]:.1f}}}$')
-ax.axvspan(percs[1], percs[3], alpha=0.4, color='mediumturquoise')
+ax.axvspan(percs[1], percs[3], alpha=0.35, color='mediumturquoise')
 ax.axvspan(percs[0], percs[4], alpha=0.2, color='darkturquoise')
 ax.legend()
 ax.set_xlabel('$H_0\ [\mathrm{km/s/Mpc}]$')

### 2024-09-13 src/scripts/plot_simulation_result_multi.py

@@ -1,0 +1,14 @@
+import numpy as np
+import matplotlib.pyplot as plt
+from corner import corner
+from figaro.cosmology import Planck18
+from figaro import plot_settings
+import paths
+
+f = np.load(paths.data / 'simulation/multi' / '4a_Powell.npz')
+result = f['result']
+simulated_truth = {'$H_0$': Planck18.h*100, '$\\alpha$': 3.5, '$\\mu$': 35, '$\\sigma$': 5, '$\\delta$': 5, '$w$': 0.2, '$m_\mathrm{min}$': 5, '$m_\mathrm{max}$': 90}
+parameters = ["$H_0$", "$\\mu$", "$\\sigma$", "$\\delta$"]
+
+fig = corner(result, labels=parameters, truths=[simulated_truth[param] for param in parameters], color='steelblue', truth_color='red', levels = [0.3935, 0.9], plot_datapoints=False, plot_density=False, smooth=1, fill_contours=True, contourf_kwargs={'colors': ['white', 'darkturquoise', 'mediumturquoise'], 'alpha': [1, 0.2, 0.5]}, contour_kwargs={'linewidths': 1})
+fig.savefig(paths.figures / "simulation_result_multi.pdf", bbox_inches='tight')

### 2024-09-24 src/scripts/plot_real_result_multi.py
Bugfix

@@ -5,7 +5,7 @@ from figaro.cosmology import Planck18
 from figaro import plot_settings
 import paths
 
-f = np.load(paths.data / 'simulation/multi' / '4b_Powell.npz')
+f = np.load(paths.data / 'real/multi' / '4b_Powell.npz')
 result = f['result']
 parameters = ["$H_0$", "$\\alpha$", "$\\mu$", "$\\sigma$"]
 

### 2024-09-25 src/scripts/infer_H0.py
add validation

@@ -17,6 +17,10 @@ def p_z(z, H0):
 label = sys.argv[1]
 outdir = paths.data / label
 
+if label == "real":
+    print("Invalid label!")
+    sys.exit(1)
+
 print("Preparing model pdfs...")
 grid_label = "grid"
 try:

### 2024-09-25 src/scripts/infer_multi.py
Change fixed value

@@ -23,8 +23,8 @@ if not os.path.exists(outdir/f'multi/{param}_{method}.npz'):
     import dill
 
     # Redshift distribution
-    def p_z(z, H0):
-        return CosmologicalParameters(H0/100., 0.315, 0.685, -1., 0., 0.).ComovingVolumeElement(z)/(1+z)
+    def p_z(z, H0, kappa=0):
+        return CosmologicalParameters(H0/100., 0.315, 0.685, -1., 0., 0.).ComovingVolumeElement(z)*(1+z)**(kappa-1)
 
     bounds_dict = {
         "H0": (10, 300),
@@ -34,41 +34,61 @@ if not os.path.exists(outdir/f'multi/{param}_{method}.npz'):
         "w": (0, 1),
         "delta": (0.01, 20),
         "mmin": (1, 10),
-        "mmax": (70, 150)
+        "mmax": (70, 150),
+        "kappa": (-10, 10)
     }
-
-    if param == "2a":
-        bounds = bounds_dict["H0"], bounds_dict["alpha"]
-        def plp(m, x):
-            return plpeak(m, alpha=x[0])
-    elif param == "2b":
-        bounds = bounds_dict["H0"], bounds_dict["mu"]
-        def plp(m, x):
-            return plpeak(m, mu=x[0])
-    elif param == "4a":
-        bounds = bounds_dict["H0"], bounds_dict["mu"], bounds_dict["sigma"], bounds_dict["delta"]
-        def plp(m, x):
-            return plpeak(m, mu=x[0], sigma=x[1], delta=x[2])
-    elif param == "4b":
-        bounds = bounds_dict["H0"], bounds_dict["alpha"], bounds_dict["mu"], bounds_dict["sigma"]
-        def plp(m, x):
-            return plpeak(m, alpha=x[0], mu=x[1], sigma=x[2])
-    elif param == "5a":
-        bounds = bounds_dict["H0"], bounds_dict["alpha"], bounds_dict["mu"], bounds_dict["sigma"], bounds_dict["w"]
-        def plp(m, x):
-            return plpeak(m, alpha=x[0], mu=x[1], sigma=x[2], w=x[3])
-    elif param == "5b":
-        bounds = bounds_dict["H0"], bounds_dict["alpha"], bounds_dict["mu"], bounds_dict["mmin"], bounds_dict["w"]
-        def plp(m, x):
[…]

### 2024-09-26 src/scripts/infer_multi.py
Add comment

@@ -74,6 +74,7 @@ if not os.path.exists(outdir/f'multi/{param}_{method}.npz'):
             print("Invalid argument!")
             sys.exit(1)
     elif label == "real":
+        # Fixed parameters according to result from "Constraints on the Cosmic Expansion History from GWTC–3" (SNR > 10 & w0flatLCDM)
         if param == "3":
             bounds = bounds_dict["H0"], bounds_dict["alpha"], bounds_dict["mu"], bounds_dict["sigma"], bounds_dict["kappa"]
             def p_m_p_z(x):

### 2024-09-26 src/tex/ms.tex
[showyourwork] overleaf sync

@@
-\documentclass[twocolumn]{aastex631}
+\documentclass[sn-aps]{sn-jnl}
 
@@
 
-\author{Thomas~C.~K.~Ng}
-\affiliation{Department of Physics, The Chinese University of Hong Kong, Shatin, Hong Kong}
+\author*[1]{Thomas C. K. Ng}\email{thomas.ng@link.cuhk.edu.hk}
+\author[2,3]{Stefano Rinaldi}\email{stefano.rinaldi@uni-heidelberg.de}
+\author[1]{Otto A. Hannuksela}\email{hannuksela@phy.cuhk.edu.hk}
 
-\author{Stefano~Rinaldi}
-\affiliation{Institut~für~Theoretische~Astrophysik, ZAH, Universität~Heidelberg, Albert-Ueberle-Stra{\ss}e~2, 69120 Heidelberg, Germany}
-\affiliation{Dipartimento di Fisica e Astronomia ``G. Galilei'', Università di Padova, Via Marzolo 8, 35122 Padova, Italy}
+\affil[1]{Department of Physics, The Chinese University of Hong Kong, Shatin, Hong Kong}
+\affil[2]{Institut~für~Theoretische~Astrophysik, ZAH, Universität~Heidelberg, Albert-Ueberle-Stra{\ss}e~2, 69120 Heidelberg, Germany}
+\affil[3]{Dipartimento di Fisica e Astronomia ``G. Galilei'', Università di Padova, Via Marzolo 8, 35122 Padova, Italy}
 
-\author{Otto~A.~Hannuksela}
-\affiliation{Department of Physics, The Chinese University of Hong Kong, Shatin, Hong Kong}
+\abstract{
+    The challenge of understanding the Universe's dynamics, particularly the Hubble tension, requires precise measurements of the Hubble constant.
+    Building upon the existing spectral-siren method, which capitalizes on population information from gravitational-wave sources, this paper explores an alternative way to analyze the population data to obtain the cosmological parameters in $\Lambda$CDM.
+    We demonstrated how non-parametric methods, which are flexible models that can be used to agnostically reconstruct arbitrary probability densities, can be incorporated into this framework and leverage the detector-frame mass distribution to infer the cosmological parameters.
+    \todo{Rewrite abstract}
+}
 
-\correspondingauthor{Thomas~C.~K.~Ng}
-\email{thomas.ng@link.cuhk.edu.hk}
-
-\begin{abstract}
-    The challenge of understanding the Universe's dynamics, particularly the Hubble tension, requires precise measurements of the Hubble constant.
-    Building upon the existing spectral siren method, which capitalizes on population information from gravitational wave sources, this study explores an alternative way to analyze the population data to obtain the cosmological parameters in $\Lambda$CDM.
-    We will focus on how non-parametric methods, which are flexible models that can be used to agnostically reconstruct arbitrary probability densities, can be incorporated into this framework and leverage the detector-frame mass distribution to infer the cosmological parameters.
-    \todo{Rewrite abstract}
-\end{abstract}
-
-% \maketitle
+\maketitle
 
@@
 Another alternative method is to use population statistics to infer $H_0$ by marginalizing the redshift distribution of the sources.
-This method is known as the spectral siren method \citep{You:2020wju, Mastrogiovanni:2021wsd, LIGOScientific:2021aug, Ezquiaga:2022zkx}.
+This method is known as the spectral-siren method \citep{You:2020wju, Mastrogiovanni:2021wsd, LIGOScientific:2021aug, Ezquiaga:2022zkx}.
 
-The fundamental idea of the spectral siren method is to infer cosmology by comparing the observed mass population distribution with the intrinsic mass population distribution since the observed distribution is redshifted due to the expansion of the Universe.
-Assumptions about the intrinsic population model and cosmological model are required to perform the comparison.
-Previous studies often infer either the intrinsic population model or the cosmological model, but not both simultaneously, due to the expensive computational cost.
-\todo{Summarize previous work and add references}
-However, the intrinsic population model and the cosmological model are degenerate in general, and it is necessary to infer both simultaneously.
+The fundamental idea of the spectral-siren method is to infer cosmology by comparing the observed mass population distribution with the intrinsic mass population distribution since the observed distribution is redshifted due to the expansion of the Universe.
+The method works best when the intrinsic distribution is well understood.
+However, the intrinsic distribution is not known, and the method depends on the choice of the population model.
+Furthermore, the intrinsic distribution is degenerate with cosmology.
+For example, an intrinsic distribution with a linear redshift evolution is degenerate with $H_0$, as they both redshift the observed distribution in the same way.
+While this example is not physical, this degeneracy still affects the inference of $H_0$.
+Due to these issues, the spectral-siren method is not straightforward to apply, intrinsic population model parameters and cosmological parameters are required to be inferred simultaneously, which makes the method computationally expensive.
 
-To address this issue, we study an alternative method making use of non-parametric methods.
-Non-parametric methods allow us to reconstruct arbitrary probability densities without making assumptions about the form of the distribution.
+An alternative approach to the spectral-siren method is to use non-parametric methods to reconstruct the observed mass population distribution directly from the data.
+Non-parametric methods allow us to reconstruct arbitrary probability densities without making assumptions about the form of the distribution \citep{Rinaldi:2021bhm}.
 In our context, we can use non-parametric methods to reconstruct the observed mass population distribution directly from the data.
@@
 
-Some other studies have made use of non-parametric methods with the spectral siren framework.
-\todo{Summarize previous work, add references and compare with our work}
-
-Considering the increasing number of \ac{GW} events and the improvement of the detectors, we expect that the spectral siren method will give us an independent and precise measurement of $H_0$ which is comparable to the previous measurements \citep{Planck:2018vyg, Riess:2021jrx}.
-However, performing standard hierarchical Bayesian inference with the spectral siren method is computationally expensive, and it depends on the choice of the population model and the cosmological model.
-Our method provides an alternative to studying different models without performing lengthy hierarchical analysis.
-This allows us to explore more models before performing the full hierarchical Bayesian inference, which makes our method a powerful diagnostic tool for model selection.
+Some other studies have made use of non-parametric methods with the spectral-siren framework.
+\todo{summarize previous work with non-parametric methods and compare with our work}
 
@@
 In this section, we present the framework we developed to infer both the intrinsic population model and the cosmological model simultaneously by leveraging the detector-frame mass population distribution.
-In the remainder of this Section, we will make use of the notation summarized in Table~\ref{tab:notation}, mutated from \citet{Rinaldi:2021bhm, Rinaldi:2022kyg}.
+In the remainder of this Section, we will make use of the notation summarized in Table~\ref{tab:notation}, taken from \cite{Rinaldi:2021bhm, Rinaldi:2022kyg}.
 
-\begin{table}[h]
-    \caption{Notation}
-    \begin{ruledtabular}
-        \begin{tabular}{c>{\raggedright\arraybackslash}p{0.7\linewidth}}
+\begin{table}[htbp]
+    \caption{Notation used in this paper}
+    \begin{tabular}{cp{0.7\linewidth}}
+        \toprule
         Notation & Description \\
-            \hline
+        \midrule
         $m_1$ & Primary mass of the \ac{GW} sources in the source frame \\
@@
         $\mathrm{det}$ & detectability of the \ac{GW} events \\
+        \botrule
     \end{tabular}
-    \end{ruledtabular}
     \label{tab:notation}
@@
 
-Specifically, we propose two major modifications to the standard spectral siren method.
+Specifically, we propose two major modifications to the standard spectral-siren method.
 First, instead of transforming the individual-event posterior probability distribution samples from the detector frame to the source frame during the \ac{PE} process, we transform the population model to the detector frame before performing \ac{PE}.
@@
 
-With these modifications, we separate the spectral siren method into two distinct parts.
+With these modifications, we separate the spectral-siren method into two distinct parts.
 The first part is the reconstruction of the observed population from the data, which does not require any assumptions about the population model, and the intermediate result is completely data-driven.
-We make use of the \ac{(H)DPGMM} developed in \citet{Rinaldi:2021bhm} as a non-parametric model to reconstruct the observed population distribution.
+We make use of the \ac{(H)DPGMM} developed in \cite{Rinaldi:2021bhm} as a non-parametric model to reconstruct the observed population distribution.
 Other assumptions including the population model, cosmological model and selection function are then incorporated in the second half.
@@
 The first step of our method is to reconstruct the detector-frame observed population.
-we choose to consider the primary mass $m_1$ of the \ac{BBH} sources, as we want to study the cosmological parameters with the spectral siren method.
+We choose to consider the primary mass $m_1$ of the \ac{BBH} sources, as we want to study the cosmological parameters with the spectral-siren method.
 In principle, we can consider more parameters at once by reconstructing not only a single-parameter observed distribution but instead a multi-parameter observed distribution.
@@
 With the posterior probability distribution samples obtained from the \ac{PE} of each \ac{GW} event, we reconstruct the observed detector-frame primary mass population distribution.
-We use the \ac{(H)DPGMM} developed in \citet{Rinaldi:2021bhm} as a non-parametric model to reconstruct the observed $m^z_1$ population distribution $p(m^z_1|\mathbf{\Theta})$.
+We use the \ac{(H)DPGMM} developed in \cite{Rinaldi:2021bhm} as a non-parametric model to reconstruct the observed $m^z_1$ population distribution $p(m^z_1|\mathbf{\Theta})$.
 The \ac{(H)DPGMM} is a hierarchy of Gaussian mixture models with a Dirichlet process prior probability distribution, which allows the model to be flexible to reconstruct arbitrary population distribution from individual-event posterior probability distribution samples.
@@
 This part of the method is completely data-driven.
-Therefore the intermediate result is independent of the population, cosmological model and selection function, and we can reuse the result for testing different models.
+Therefore, the intermediate result is independent of the population, cosmological model and selection function, and we can reuse the result for testing different models.
 This reduces the computational cost of the method, as we only need to perform the reconstruction once to analyze different models.
@@
 We then transform $p(m_1|\Lambda)$ to the detector frame with a given cosmological model $\Omega$ to obtain the population distribution $p(m^z_1|\Lambda, \Omega)$ by
-\begin{widetext}
 \begin{equation}
@@
         &= \int p(m^z_1|m_1, z) \frac{p(\mathrm{det}|m_1, z, \Lambda, \Omega)p(m_1, z|\Lambda, \Omega)}{p(\mathrm{det}|\Lambda, \Omega)} \mathrm{d}m_1 \mathrm{d}z \\
-        &= \frac{1}{p(\mathrm{det}|\Lambda, \Omega)}\int \delta(m^z_1-m_1(1+z))p(\mathrm{det}|m_1, z, \Omega)p(m_1|\Lambda)p(z|\Lambda) \mathrm{d}m_1 \mathrm{d}z \\
-        &\propto \int p\left(\mathrm{det}\middle|\frac{m^z_1}{1+z},z,\Omega\right)p\left(\frac{m^z_1}{1+z}\middle|\Lambda\right)p(z|\Lambda) \mathrm{d}z\,,
+        &= \frac{1}{p(\mathrm{det}|\Lambda, \Omega)}\int \delta(m^z_1-m_1(1+z))p(\mathrm{det}|m_1, z, \Omega)p(m_1|\Lambda)p(z|\Lambda, \Omega) \mathrm{d}m_1 \mathrm{d}z \\
+        &\propto \int p\left(\mathrm{det}\middle|\frac{m^z_1}{1+z},z,\Omega\right)p\left(\frac{m^z_1}{1+z}\middle|\Lambda\right)p(z|\Lambda, \Omega) \mathrm{d}z\,,
     \end{aligned}
@@
 \end{equation}
-\end{widetext}
 where $\mathrm{det}$ represents whether the event is detectable or not.
 We assume the $m_1$ distribution does not evolve with $z$.
+
 The selection effect is included in the derivation in Eq.~\eqref{eq:transformation}, specifically, the first term in the final expression, which is the selection function.
@@
 
-Opposite to the first part of the method, this part includes the assumptions about the population and cosmological model, as well as the selection function.
+In contrast to the first part of the method, this part includes the assumptions about the population and cosmological model, as well as the selection function.
 After obtaining $p(m^z_1|\Lambda, \Omega, \mathrm{det})$, we can perform the \ac{PE} to infer the optimized $\Omega$ and $\Lambda$.
@@
 
-With $p(m^z_1|\mathbf{\Theta})$ and $p(m^z_1|\Lambda, \Omega, \mathrm{det})$ obtained from Sec.~\ref{sec:reconstruction} and Sec.~\ref{sec:transformation} respectively, we can perform the \ac{PE} to infer the free parameters in the parametric models assumed in Sec.~\ref{sec:transformation}.
-The \ac{PE} is performed by comparing $p(m^z_1|\Theta_i)$ with $p(m^z_1|\Lambda, \Omega)$.
-To compare the two populations, we use the Jensen-Shannon distance $d_\mathrm{JS}$ as the distance measure.
+With $p(m^z_1|\mathbf{\Theta})$ and $p(m^z_1|\Lambda, \Omega, \mathrm{det})$ obtained from Sec.~\ref{sec:reconstruction} and Sec.~\ref{sec:transformation} respectively, we need a way of translating each realization of the reconstructed distribution to a set of population and cosmological parameters.
+To do so, we define a notion of "closeness" using the Jensen-Shannon distance $d_\mathrm{JS}$ between the two distributions.
 $d_\mathrm{JS}$ is calculated as
@@
 For each reconstructed distribution $p(m^z_1|\Theta_i)$, we find an optimized $\Lambda_i$ and $\Omega_i$ by minimizing $d_\mathrm{JS}$.
-For the minimization, we use the \textsc{scipy} \citep{2020SciPy-NMeth} implementation of the \todo{XXX (add reference)} algorithm.
-Note that we calculate $d_\mathrm{JS}$ in the range from the minimum to the maximum of the median of the $m^z_1$ posterior probability distribution samples of each \ac{GW} event
+For the minimization, we use the \textsc{scipy} \citep{2020SciPy-NMeth} implementation of the modified Powell algorithm.
+Note that we calculate $d_\mathrm{JS}$ in the range from the minimum to the maximum of the median of the $m^z_1$ posterior probability distribution samples of each \ac{GW} event.
 This is because the reconstruction is only valid in the region where observations are present.
@@
 
-The analysis code is publicly available.\footnote{The analysis code is available at \url{https://github.com/thomasckng/det-frame-cosmo-with-FIGARO/tree/main/src/scripts}.}
+The analysis code we developed to implement the method presented in this section is publicly available on GitHub\footnote{ \url{https://github.com/thomasckng/det-frame-cosmo-with-FIGARO/tree/main/src/scripts}}.
 
@@
 We then transformed the $m_1$ and $z$ samples to $m^z_1$ and $d_L$ samples.
-Here we chose the $\Lambda$CDM model with parameters consistent with the results in \citet{Planck:2018vyg} as the cosmological model for the transformation of the samples.
+Here we chose the $\Lambda$CDM model with parameters consistent with the results in \cite{Planck:2018vyg} as the cosmological model for the transformation of the samples.
 
@@
 In this subsection, we first test our method with a simplified case where we only infer $H_0$ to demonstrate the basic idea of our method.
-We assume the selection function, the true population model and its parameters, and the true cosmological model and its parameters except $H_0$ are known.
+We assume the selection function, the true mass and redshift population model and its parameters, and the true cosmological model and its parameters except $H_0$ are known.
 
-We first performed the reconstruction of the observed $m^z_1$ population distribution with \textsc{figaro} as described in Sec.~\ref{sec:reconstruction}.
+We first reconstructed the observed $m^z_1$ population distribution with \textsc{figaro} as described in Sec.~\ref{sec:reconstruction}.
 The result of the reconstruction is shown in Fig.~\ref{fig:simulation_reconstruction}.
-
-\begin{figure}[h]
+\begin{figure}[htbp]
     \script{plot_simulation_reconstruction.py}
@@
         In the upper panel, the blue line is the median of the reconstructed distributions, and the shaded region represents the 68\% and 90\% credible intervals of the reconstructed distributions.
-        In the lower panel, the blue lines are the single reconstructed distributions from a subset of all the reconstructed distributions.
+        In the lower panel, the blue lines are the single-reconstructed distributions from a subset of all the reconstructed distributions.
         By combining all the single reconstructed distributions, we obtain the distribution in the upper panel.
@@
 The result of the transformation is shown in Fig.~\ref{fig:simulation_transformation}.
-
-\begin{figure}[h]
+\begin{figure}[htbp]
     \script{plot_simulation_transformation.py}
@@
 For each $H_0$, we computed the $d_\mathrm{JS}$ between the two populations and found the optimized $H_0$ that minimizes $d_\mathrm{JS}$.
-
-\begin{figure}[h]
+\begin{figure}[htbp]
     \script{plot_simulation_comparison.py}
@@
 The result of the \ac{PE} is shown in Fig.~\ref{fig:simulation_result_H0}.
-The result shows that our method can recover the true $H_0$ we used to generate the mock data.
-
-\begin{figure}[h]
+\begin{figure}[htbp]
     \script{plot_simulation_result_H0.py}
@@
         The vertical orange line represents the true $H_0$ we used to generate the mock data.
-        Note that the distribution only shows the uncertainty in the reconstruction.
-        The uncertainty in the comparison is not considered in the distribution.
     }
@@
 \end{figure}
+We find that the $H_0$ distribution, with which the uncertainty associated comes from the uncertainty in the reconstruction, is consistent with the simulated $H_0$ value.
 
@@
 Similar to the previous subsection, we assume only the free parameters are unknown.
-Other than $H_0$, we also chose the power law index $\alpha$, the peak mass $\mu$, the peak width $\sigma$, and the peak weight $w$ of the \textsc{power law + peak} model as the free parameters.
-Fixed parameters are the minimum and maximum mass, as well as the range of mass tapering at the low mass end.
-We may not be able to infer these fixed parameters accurately with our method, as the reconstruction is only valid in the region where the samples are present and the reconstruction may be not accurate at the boundaries depending on the boundary features of the observed population\footnote{As we are using a Gaussian mixture model to reconstruct the observed population, the reconstruction may not be accurate at the boundaries if the observed population has sharp features at the boundaries.}.
+Other than $H_0$, we also chose the power law index $\alpha$, the peak mass $\mu$, and the peak width $\sigma$ of the \textsc{power law + peak} model as the free parameters.
+Fixed parameters are the minimum and maximum mass, the range of mass tapering at the low mass end $\delta$, the peak weight $w$, and the cosmological parameters except $H_0$.
+We may not be able to infer some of these fixed parameters (e.g., the minimum and maximum mass, and $\delta$) accurately with our method, as the reconstruction is only valid in the region where the samples are present, and the reconstruction may be not accurate at the boundaries depending on the boundary features of the observed population\footnote{As we are using a Gaussian mixture model to reconstruct the observed population, the reconstruction may not be accurate at the boundaries if the observed population has sharp features at the boundaries.}.
 
@@
 The optimization method requires bounds for the free parameters, which we show in Table~\ref{tab:bounds}.
-\begin{table}[h]
+\begin{table}[htbp]
     \caption{Bounds for the free parameters}
-    \begin{ruledtabular}
     \begin{tabular}{ccc}
+        \toprule
         Parameter & Symbol & Bounds \\
-            \hline
+        \midrule
         Hubble constant & $H_0$ & $[10, 300]$ \\
@@
         Peak width & $\sigma$ & $[0.01, 10]$ \\
-            Peak weight & $w$ & $[0, 1]$ \\
+        \botrule
     \end{tabular}
-    \end{ruledtabular}
     \label{tab:bounds}
@@
 The initial guesses for the optimization are chosen randomly from a uniform distribution within the bounds.
-The result of the \ac{PE} is shown in Fig.~\todo{add figure}.
-The result shows that our method can infer multiple parameters simultaneously.
+The result of the \ac{PE} is shown in Fig.~\ref{fig:simulation_result_multi}.
+\begin{figure}[htbp]
+    \script{plot_simulation_result_multi.py}
+    \includegraphics[width=\linewidth]{figures/simulation_result_multi.pdf}
+    \caption{
+        The corner plot shows the distribution of the optimized parameters.
+        The diagonal histograms show the marginalized distributions.
+        The off-diagonal contour plots show the joint distributions, where the contours represent the 39.35\%, 90\% credible intervals.
+        The red lines and points represent the true values we used to generate the mock data.
+    }
+    \label{fig:simulation_result_multi}
+\end{figure}
+We find that the distribution, with the uncertainty associated comes from the uncertainty in the reconstruction, is consistent with the simulated values.
+Some features of the result are discussed together with the result of the real data in Sec.~\ref{sec:real_data}.
 
@@
 In this section, we apply our method to the real data from the \ac{LVK}.
-We included all $70$ events labeled as BBH in Table~I of \citet{KAGRA:2021duu}.
+We included all $70$ events labeled as BBH in Table~I of \cite{KAGRA:2021duu}.
 
-We first performed the reconstruction of the observed $m^z_1$ population distribution with \textsc{figaro} as described in Sec.~\ref{sec:reconstruction}.
-We use the posterior probability distribution samples of the \ac{GW} events from \citet{LIGOScientific:2019lzm, KAGRA:2023pio}.
+Similar to Sec.~\ref{sec:mock_data}, we first reconstructed the observed $m^z_1$ population distribution.
+We use the posterior probability distribution samples of the \ac{GW} events from \cite{LIGOScientific:2019lzm, KAGRA:2023pio}.
 $10000$ samples are randomly drawn from the posterior probability distribution samples of each \ac{GW} event to perform the reconstruction.
 The result of the reconstruction is shown in Fig.~\ref{fig:real_reconstruction}.
-
-\begin{figure}[h]
+\begin{figure}[htbp]
     \script{plot_real_reconstruction.py}
@@
 Similar to Sec.~\ref{sec:inference_multi}, we fixed some parameters of the population model and the cosmological model.
-For the cosmological model, we fixed the parameters to the results from \citet{Planck:2018vyg} except $H_0$.
-For the population model, we fixed the minimum and maximum mass, as well as the range of mass tapering at the low mass end to the results from \citet{KAGRA:2021duu}.
-Note that the fixed parameters are not necessarily the true values\footnote{Result from \citet{KAGRA:2021duu} assumes a cosmological model to infer the population model parameters.}, but they can be considered as a reasonable guess for the parameters.
+For the cosmological model, we fixed the parameters to the results from \cite{Planck:2018vyg} except $H_0$.
+For the population model, we fixed the minimum and maximum mass, as well as the range of mass tapering at the low mass end to the results from \cite{KAGRA:2021duu}.
+\todo{Update the fixed parameters}
+Note that the fixed parameters are not necessarily the true values\footnote{Result from \cite{KAGRA:2021duu} assumes a cosmological model to infer the population model parameters.}, but they can be considered as a reasonable guess for the parameters.
 The fixed parameters are shown in Table~\todo{add table}.
@@
 We then perform the remaining steps similar to Sec.~\ref{sec:inference_multi}.
-The result of the \ac{PE} is shown in Fig.~\todo{add figure} and Fig.~\todo{add figure}.
-\todo{Discuss the result}
-
-The result shows similar problems mentioned in \citet{Pierra:2024wrk}.
+The result of the \ac{PE} is shown in Fig.~\ref{fig:real_result_multi}.
+\begin{figure}[htbp]
+    \script{plot_real_result_multi.py}
+    \includegraphics[width=\linewidth]{figures/real_result_multi.pdf}
+    \caption{
+        The corner plot shows the distribution of the optimized parameters.
+        The diagonal histograms show the marginalized distributions.
+        The off-diagonal contour plots show the joint distributions, where the contours represent the 39.35\%, 90\% credible intervals.
+    }
+    \label{fig:real_result_multi}
+\end{figure}
+\todo{Discuss the result (choice of redshift distribution)}
+The result shows similar problems mentioned in \cite{Pierra:2023deu}.
 \todo{Discuss mass-redshift dependent population model \& possible issues with PL+P model}
@@
 
-\todo{Work as diagnostic tool}
+With the numbers of observed \ac{GW} events that are expected to grow larger and larger, potentially hitting the thousand mark by the end of the fifth observing run, performing a complete analysis from scratch with different astrophysical models will soon become too computationally expensive.
+The method we propose in this paper can help to mitigate this issue by providing a fast and efficient way to explore different models before performing the full hierarchical Bayesian inference.
+This makes our method a powerful diagnostic tool for model selection.
 
-\begin{acknowledgments}
+\backmatter
 
-T.~N.~ and S.~R.~ acknowledge financial support from the German Excellence Strategy via the Heidelberg Cluster of Excellence (EXC 2181 - 390900948) STRUCTURES.
-S.~R.~ acknowledges financial support from the European Research Council for the ERC Consolidator grant DEMOBLACK, under contract no. 770017. 
+\bmhead{Acknowledgements}
+
+T.~N. and S.~R. acknowledge financial support from the German Excellence Strategy via the Heidelberg Cluster of Excellence (EXC 2181 - 390900948) STRUCTURES.
+S.~R. acknowledges financial support from the European Research Council for the ERC Consolidator grant DEMOBLACK, under contract no. 770017. 
 This research made use of the bwForCluster Helix: the authors acknowledge support by the state of Baden-Württemberg through bwHPC and the German Research Foundation (DFG) through grant INST 35/1597-1 FUGG.
@@
 
-\end{acknowledgments}
-
 \bibliography{bib}

### 2024-09-26 src/tex/ms.tex
[showyourwork] overleaf sync

@@
-\documentclass[sn-aps]{sn-jnl}
+\documentclass[sn-aps, pdflatex]{sn-jnl}
 
@@
 
-We also mentioned a limitation of the method in Sec.~\ref{sec:inference_multi}.
-The runtime of the optimization procedure scales with the number of parameters.
-To perform higher dimensional analysis, we need to study using, instead of $d_\mathrm{JS}$, other distance measures that allow for faster optimization.
-Either a distance measure that gives a simpler geometry to be optimized or a distance measure that can be calculated faster will be useful.
-Other optimization algorithms should also be considered to allow for higher dimensional analysis and faster runtime.
-
 \section{Conclusion}

### 2024-09-27 src/scripts/infer_multi.py
Modify result plot

@@ -76,7 +76,7 @@ if not os.path.exists(outdir/f'multi/{param}_{method}.npz'):
     elif label == "real":
         # Fixed parameters according to result from "Constraints on the Cosmic Expansion History from GWTC–3" (SNR > 10 & w0flatLCDM)
         if param == "3":
-            bounds = bounds_dict["H0"], bounds_dict["alpha"], bounds_dict["mu"], bounds_dict["sigma"], bounds_dict["kappa"]
+            bounds = bounds_dict["H0"], bounds_dict["alpha"], bounds_dict["kappa"]
             def p_m_p_z(x):
                 return np.einsum("ij, j -> ij", plpeak(m, alpha=x[1], mu=31.825703461630482, sigma=3.7904971258458042, w=0.024059947239759398, delta=4.8961636235644015, mmin=5.0808821331157095, mmax=109.03299036617125), p_z(z, x[0], kappa=x[2]))
         elif param == "5c":

### 2024-09-27 src/scripts/plot_simulation_result_multi.py

@@ -7,6 +7,8 @@ import paths
 
 f = np.load(paths.data / 'simulation/multi' / '4b_Powell.npz')
 result = f['result']
+result = result[result[:, 0] < 200]
+result = result[result[:, 3] < 9]
 simulated_truth = {'$H_0$': Planck18.h*100, '$\\alpha$': 3.5, '$\\mu$': 35, '$\\sigma$': 5, '$\\delta$': 5, '$w$': 0.2, '$m_\mathrm{min}$': 5, '$m_\mathrm{max}$': 90}
 parameters = ["$H_0$", "$\\alpha$", "$\\mu$", "$\\sigma$"]
 

### 2024-09-27 src/tex/ms.tex

@@
 In our context, we can use non-parametric methods to reconstruct the observed mass population distribution directly from the data.
-This provides an intermediate observation-driven result that encodes both the intrinsic population information and the cosmological information.
-By separating the reconstruction of the observed population from the inference of the intrinsic population and cosmological model, we can explore different models easily.
+We can then compare the reconstructed distribution with the predicted distribution from some intrinsic population model and cosmological model, which allows us to infer the model parameters by minimizing the distance between the two distributions.
+
+This approach allows us to extract the information contained in the observed data without making assumptions about the population model, cosmological model, or selection function.
+The reconstructed distribution can act as an intermediate observation-driven result that encodes both the intrinsic population information and the cosmological information.
+Making use of the reconstructed distribution, we can easily explore the effect of using different models, as the reconstructed distribution can be reused for different models.
 Furthermore, the intermediate result represents the information contained in the observed data without any assumptions about the population model, cosmological model, or selection function, which allows us to study the features of the observed population directly.
@@
 We make use of the \ac{(H)DPGMM} developed in \cite{Rinaldi:2021bhm} as a non-parametric model to reconstruct the observed population distribution.
-Other assumptions including the population model, cosmological model and selection function are then incorporated in the second half.
+Other assumptions including the population model, cosmological model and selection function are then incorporated in the second half, where we transform the source-frame population model to the detector frame with a given cosmological model.
 This separation allows us to explore the impact of different assumptions easily.
@@
 In this work, we marginalized over $q$ to study the $m_1$ population distribution only.
-Due to the censoring of the low and high mass events, we consider the selection function in the transformation instead of the reconstruction.
+
+The selection effect is chosen to be considered in the transformation instead of the reconstruction, as the uncertainty in the reconstruction will diverge in the region where the selection function censoring the data.
+For example, if we consider the selection function in the reconstruction, the uncertainty in the reconstruction will be large at the low mass end, as the selection function censors the low mass events effectively.
+\cite{Essick:2023upv} also pointed out that including the selection function in the reconstruction may lead to insignificant biases in the analysis.
+Therefore, we consider the selection function in the transformation to avoid these issues.
+
+Qualitatively, the transformation described in Eq.~\eqref{eq:transformation} redshifts the source-frame intrinsic population distribution to the detector frame, where the redshift is determined by the cosmological model.
+The selection function is then applied to the redshifted distribution to obtain the detector-frame observed population distribution.
 With different $\Lambda$ or $\Omega$, the resulting $p(m^z_1|\Lambda, \Omega, \mathrm{det})$ will be different.
@@
 Another potential improvement is to consider not only the primary mass distribution.
-For example, we can consider the joint distribution of the primary mass and luminosity distance.
+For example, we can consider the joint distribution of the primary mass, mass ratio, and redshift.
 The reconstruction of the joint observed distribution can also be performed in the current framework.
-However, we need a joint population model that considers both the primary mass and luminosity distance.
+However, we need a joint population model that considers all the parameters, including the possible correlations between the parameters.
 Performing the analysis of the joint distribution will allow us to study mass population models that evolve with redshift.

### 2024-09-27 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \usepackage{graphicx}
-
-\newcommand{\todo}[1]{\textcolor{red}{[TODO: #1]}}
-\newcommand{\ste}[1]{\textcolor{blue}{[Ste: #1]}}
 
@@
     We demonstrated how non-parametric methods, which are flexible models that can be used to agnostically reconstruct arbitrary probability densities, can be incorporated into this framework and leverage the detector-frame mass distribution to infer the cosmological parameters.
-    \todo{Rewrite abstract}
 }
@@
 Some other studies have made use of non-parametric methods with the spectral-siren framework.
-\todo{summarize previous work with non-parametric methods and compare with our work}
 
@@
 Finally, we conclude in Sec.~\ref{sec:conclusion} with a summary.
-\todo{Rewrite outline}
 
@@
 The uncertainty of the reconstruction can then be quantified by the spread of the distributions.
-It is important to note that the reconstruction is only valid in the region where the samples are present.
+It is important to note that the reconstruction is only valid in the region where the samples are present and cannot be extrapolated to regions where no samples are present.
 In Fig.~\ref{fig:simulation_reconstruction}, we show an example of the reconstruction obtained with \textsc{figaro}.
@@
 Note that we calculate $d_\mathrm{JS}$ in the range from the minimum to the maximum of the median of the $m^z_1$ posterior probability distribution samples of each \ac{GW} event.
-This is because the reconstruction is only valid in the region where observations are present.
-As each reconstructed distribution is independent, the optimization can be performed in parallel easily, which reduces the runtime as long as computational resources are available.
+This is because the reconstruction is only valid in the region where observations are present as mentioned in Sec.~\ref{sec:reconstruction}.
+The optimization procedure can be performed in parallel for each reconstructed distribution, as each reconstructed distribution is independent.
+This reduces the runtime as long as computational resources are available.
 In Fig.~\ref{fig:simulation_comparison}, we show an illustration of the comparison between the two populations.
@@
 After the transformation, we applied a selection function to the $m^z_1$ and $d_L$ samples.
-We chose the O3 selection function from \todo{add reference} as the selection function.
+We used the physically modelled selection function implemented in \cite{Lorenzo-Medina:2024opt}.
 Note that we marginalized over $q$ for the selection function, similar to the transformation in Sec.~\ref{sec:transformation}.
@@
 
-We first reconstructed the observed $m^z_1$ population distribution with \textsc{figaro} as described in Sec.~\ref{sec:reconstruction}.
+We first reconstructed the observed detector-frame mass population distribution with \textsc{figaro} as described in Sec.~\ref{sec:reconstruction}.
 The result of the reconstruction is shown in Fig.~\ref{fig:simulation_reconstruction}.
@@
 
-Next, we chose the true population model we used to generate the mock data as the source-frame population model $\Lambda$.
-We then transformed $p(m_1|\Lambda)$ to the detector frame with the true cosmological model we used to generate the mock data, except we did not fix $H_0$.
-We performed the transformation for each $H_0$ in the range from $5$ to $150$.
-After the transformation, we applied the selection function we used to generate the mock data to the transformed distribution.
-Here we only free $H_0$ in the transformation to test if our method can recover the true $H_0$ while we fixed the other cosmological parameters to the true values.
+Next, under the assumption that the population model, cosmological model and selection function are known, except for the value of $H_0$, we can perform the transformation mentioned in Sec.~\ref{sec:transformation}.
+Following Eq.~\eqref{eq:transformation}, we can write the observed detector-frame mass population distribution as
+\begin{equation}
+    \begin{aligned}
+        &p(m^z_1|\Lambda, \Omega(H_0), \mathrm{det}) \\
+        &\propto \int p\left(\mathrm{det}\middle|\frac{m^z_1}{1+z},z,\Omega(H_0)\right)p\left(\frac{m^z_1}{1+z}\middle|\Lambda\right)p(z|\Lambda, \Omega(H_0)) \mathrm{d}z\,,
+    \end{aligned}
+    \label{eq:transformation_H0}
+\end{equation}
+where $\Omega(H_0)$ is the cosmological model with $H_0$ as a free parameter.
 The result of the transformation is shown in Fig.~\ref{fig:simulation_transformation}.
@@
 
-Finally, we performed the \ac{PE} to infer $H_0$.
-For this test specifically, instead of using the \textsc{scipy} optimization algorithm, we used a grid search to find the optimized $H_0$, as the parameter space is one-dimensional.
+We then applied our optimization scheme to each reconstructed distribution to infer the optimized $H_0$.
+In this simplified case, since the parameter space is one-dimensional, instead of using the optimization algorithm, we relied on a grid-based search to find the optimized $H_0$ between $5$ and $150$ $\mathrm{km}\,\mathrm{s}^{-1}\,\mathrm{Mpc}^{-1}$.
+For each $H_0$, we computed the $d_\mathrm{JS}$ between the two populations and found the optimized $H_0$ that minimizes $d_\mathrm{JS}$.
 Figure~\ref{fig:simulation_comparison} shows an illustration of the comparison between the reconstructed observed population and the transformed population for different $H_0$.
-For each $H_0$, we computed the $d_\mathrm{JS}$ between the two populations and found the optimized $H_0$ that minimizes $d_\mathrm{JS}$.
 \begin{figure}[htbp]
@@
 \end{figure}
-We find that the $H_0$ distribution, with which the uncertainty associated comes from the uncertainty in the reconstruction, is consistent with the simulated $H_0$ value.
+We find that the $H_0$ distribution, whose associated uncertainty comes from the reconstruction uncertainty, is consistent with the simulated $H_0$ value.
 
@@
 In this subsection, we test our method with a more realistic case where we infer $H_0$ and a subset of the population model parameters.
-Similar to the previous subsection, we assume only the free parameters are unknown.
+Similar to the previous subsection, we assume perfect knowledge of both models, the selection function and all the parameters that are not objects of the inference.
 Other than $H_0$, we also chose the power law index $\alpha$, the peak mass $\mu$, and the peak width $\sigma$ of the \textsc{power law + peak} model as the free parameters.
-Fixed parameters are the minimum and maximum mass, the range of mass tapering at the low mass end $\delta$, the peak weight $w$, and the cosmological parameters except $H_0$.
-We may not be able to infer some of these fixed parameters (e.g., the minimum and maximum mass, and $\delta$) accurately with our method, as the reconstruction is only valid in the region where the samples are present, and the reconstruction may be not accurate at the boundaries depending on the boundary features of the observed population\footnote{As we are using a Gaussian mixture model to reconstruct the observed population, the reconstruction may not be accurate at the boundaries if the observed population has sharp features at the boundaries.}.
+The minimum and maximum mass, and the range of mass tapering at the low mass end $\delta$ are fixed, as we may not be able to infer these parameters accurately with our method.
+This is because the reconstructions are only valid in the region where the samples are present, and they may be not accurate at the boundaries depending on the boundary features of the observed population distribution.
+In principle, more parameters can be inferred with our method, for example, the peak weight $w$ in the mass population model and the present mass density $\Omega_m$ in the $\Lambda$CDM model.
+However, the optimization procedure becomes more computationally expensive as the number of parameters increases, therefore, we only chose a subset of the parameters to demonstrate the idea of our method.
 
 The reconstruction of the observed $m^z_1$ population distribution is the same as in the previous subsection.
-Note that the same reconstruction result can be reused, as the data does not change.
+This also shows that the reconstruction is independent of the population model, cosmological model, and selection function, and we can reuse the result for different analysis setups.
+As long as the events included in the analysis do not change, the reconstruction can be reused.
+With more and more events in the near future, the standard spectral-siren method may not be feasible to apply, and this feature of our method can be advantageous.
 
 In this example, instead of transforming the source-frame population model to the detector frame for each $H_0$ and performing a grid search, we define a function that calculates $d_\mathrm{JS}$ for a given set of the free parameters.
-We then use the \textsc{scipy} optimization algorithm mentioned in Sec.~\ref{sec:pe} to find the optimized free parameters that minimize $d_\mathrm{JS}$.
+We then use the optimization algorithm mentioned in Sec.~\ref{sec:pe} to find the optimized free parameters that minimize $d_\mathrm{JS}$.
 The optimization method requires bounds for the free parameters, which we show in Table~\ref{tab:bounds}.
@@
     \caption{Bounds for the free parameters}
-    \begin{tabular}{ccc}
+    \begin{tabular}{cc}
         \toprule
-        Parameter & Symbol & Bounds \\
+        Parameter & Bounds \\
         \midrule
-        Hubble constant & $H_0$ & $[10, 300]$ \\
-        Power law index & $\alpha$ & $[1.01, 10]$ \\
-        Peak mass & $\mu$ & $[10, 70]$ \\
-        Peak width & $\sigma$ & $[0.01, 10]$ \\
+        $H_0$ & $[10, 300]$ \\
+        $\alpha$ & $[1.01, 10]$ \\
+        $\mu$ & $[10, 70]$ \\
+        $\sigma$ & $[0.01, 10]$ \\
         \botrule
@@
 \end{figure}
-We find that the distribution, with the uncertainty associated comes from the uncertainty in the reconstruction, is consistent with the simulated values.
+We find that the distribution, whose associated uncertainty comes from the reconstruction uncertainty, is consistent with the simulated values.
 Some features of the result are discussed together with the result of the real data in Sec.~\ref{sec:real_data}.
-
-In principle, more parameters can be inferred with our method, for example, the present mass density $\Omega_m$ in the $\Lambda$CDM model.
-However, the optimization procedure becomes more computationally expensive as the number of parameters increases.
 
@@
 Next, we chose the \textsc{power law + peak} model as the source-frame population model $\Lambda$ and the $\Lambda$CDM model as the cosmological model $\Omega$.
+For the redshift population model, instead of the uniform distribution over comoving volume and source-frame time, we used the \textsc{power law} redshift evolution model described in \cite{KAGRA:2021duu}.
 Similar to Sec.~\ref{sec:inference_multi}, we fixed some parameters of the population model and the cosmological model.
 For the cosmological model, we fixed the parameters to the results from \cite{Planck:2018vyg} except $H_0$.
-For the population model, we fixed the minimum and maximum mass, as well as the range of mass tapering at the low mass end to the results from \cite{KAGRA:2021duu}.
-\todo{Update the fixed parameters}
-Note that the fixed parameters are not necessarily the true values\footnote{Result from \cite{KAGRA:2021duu} assumes a cosmological model to infer the population model parameters.}, but they can be considered as a reasonable guess for the parameters.
-The fixed parameters are shown in Table~\todo{add table}.
-\todo{Add more details about the selection function}
-\todo{Consider using values from another paper}
+For the population model, we fixed the minimum mass $m_\mathrm{min}$ and maximum mass $m_\mathrm{max}$, as well as the range of mass tapering at the low mass end to the results from \cite{LIGOScientific:2021aug}.
+Note that the fixed parameters are not necessarily the true values, but they can be considered as a reasonable guess for the parameters.
+The fixed parameters of the population model are shown in Table~\ref{tab:fixed_parameters}.
+\begin{table}[htbp]
+    \caption{Fixed parameters}
+    \begin{tabular}{cc}
+        \toprule
+        Parameter & Value \\
+        \midrule
+        $m_\mathrm{min}$ & $5.08$ \\
+        $m_\mathrm{max}$ & $109.03$ \\
+        $\delta$ & $4.89$ \\
+        $w$ & $0.02$ \\
+        \botrule
+    \end{tabular}
+    \label{tab:fixed_parameters}
+\end{table}
 
@@
 \end{figure}
-\todo{Discuss the result (choice of redshift distribution)}
-The result shows similar problems mentioned in \cite{Pierra:2023deu}.
-\todo{Discuss mass-redshift dependent population model \& possible issues with PL+P model}
 
@@
 
-As mentioned in Sec.~\ref{sec:pe}, our method is not a Bayesian inference, as we only consider the optimization of the parameters.
+The result shows similar problems mentioned in \cite{Pierra:2023deu}.
+
+\section{Conclusion}
+\label{sec:conclusion}
+
+In this paper, we proposed a method to infer both the intrinsic population model and the cosmological model simultaneously by leveraging the detector-frame mass population distribution and making use of non-parametric methods.
+We demonstrated the method with a mock data study and applied it to the real data from the \ac{LVK}.
+With the numbers of observed \ac{GW} events that are expected to grow larger and larger, potentially hitting the thousand mark by the end of the fifth observing run, performing a complete analysis from scratch with different astrophysical models will soon become too computationally expensive.
+The method we propose in this paper can help mitigate this issue by providing a fast and efficient way to explore different models before performing the full hierarchical Bayesian inference.
+This makes our method a powerful diagnostic tool for model selection.
+
+To further improve the method, there are several directions we can explore.
+First, as mentioned in Sec.~\ref{sec:pe}, our method is not a Bayesian inference, as we only consider the optimization of the parameters.
 To perform a full Bayesian inference, we need to develop a likelihood function that considers the comparison of the two populations.
 With this likelihood, our method will give the same result as the standard hierarchical Bayesian inference, as long as the number of reconstructions is large enough.
-This will allow us to perform efficient model selection, as different models can be tested without repeating the reconstruction of the observed population.
+This will allow us to perform efficient Bayesian model selection, as different models can be tested without repeating the reconstruction of the observed population distribution.
+This is currently under development and will be presented in a future work.
 
@@
 Furthermore, the inference of cosmological parameters will be more precise, as the information from the luminosity distance is also considered.
-
-\section{Conclusion}
-\label{sec:conclusion}
-
-With the numbers of observed \ac{GW} events that are expected to grow larger and larger, potentially hitting the thousand mark by the end of the fifth observing run, performing a complete analysis from scratch with different astrophysical models will soon become too computationally expensive.
-The method we propose in this paper can help to mitigate this issue by providing a fast and efficient way to explore different models before performing the full hierarchical Bayesian inference.
-This makes our method a powerful diagnostic tool for model selection.
 

### 2024-10-02 src/scripts/plot_real_reconstruction.py
Remove plot

@@ -11,22 +11,12 @@ outdir = paths.data / label
 draws = load_density(outdir / f'draws/draws_observed_{label}.json')
 samples_med = np.loadtxt(outdir / 'samples_med.txt')
 
-fig, (ax1, ax2) = plt.subplots(2, figsize=(7,6), gridspec_kw={'hspace': 0.3})
+fig, ax  = plt.subplots()
 fig = plot_median_cr(draws, hierarchical=True, save=True, show=False, fig=fig, median_label='$p(m^z_1|\mathbf{\Theta})$')
-ax1.hist(samples_med, bins = int(np.sqrt(len(samples_med))), histtype = 'step', density = True, label = '$\mathrm{median}(Y_t)$', color = 'tab:red')
-ax1.set_xlim(0,200)
-ax1.set_ylim(0)
-ax1.set_xlabel('$m^z_1\ [\mathrm{M}_\odot]$')
-ax1.set_ylabel('$\mathrm{Density}$')
-ax1.legend()
-m_z = np.linspace(0,200,1000)
-ax2.plot(m_z, draws[0].pdf(m_z), alpha = 0.5, color = 'steelblue', label = '$p(m^z_1|\Theta_i)$')
-for i, draw in enumerate(draws[1:10]):
-    ax2.plot(m_z, draw.pdf(m_z), alpha = 0.5, color = 'steelblue')
-ax2.hist(samples_med, bins = int(np.sqrt(len(samples_med))), histtype = 'step', density = True, label = '$\mathrm{median}(Y_t)$', color = 'tab:red')
-ax2.set_xlim(0,200)
-ax2.set_ylim(0)
-ax2.set_xlabel('$m^z_1\ [\mathrm{M}_\odot]$')
-ax2.set_ylabel('$\mathrm{Density}$')
-ax2.legend()
+ax.hist(samples_med, bins = int(np.sqrt(len(samples_med))), histtype = 'step', density = True, label = '$\mathrm{median}(Y_t)$', color = 'tab:red')
+ax.set_xlim(0,200)
+ax.set_ylim(0)
+ax.set_xlabel('$m^z_1\ [\mathrm{M}_\odot]$')
+ax.set_ylabel('$\mathrm{Density}$')
+ax.legend()
 fig.savefig(paths.figures / 'real_reconstruction.pdf', bbox_inches='tight')

### 2024-10-02 src/scripts/get_lvk_results.py
Bugfix

@@ -97,8 +97,8 @@ events = [\
 'GW200216_220804',
 ]
 
-mass_med_min, mass_med_max = 0, 0
-mass_min_min, mass_max_max = 0, 0
+mass_med_min, mass_med_max = 1000, 0
+mass_min_min, mass_max_max = 1000, 0
 med_all = []
 for event in tqdm.tqdm(events):
     h5_paths = find('*' + str(event) + '*' + 'nocosmo.h5', sys.argv[1])

### 2024-10-02 src/scripts/infer_multi.py
Modify comments

@@ -24,6 +24,7 @@ if not os.path.exists(outdir/f'multi/{param}_{method}.npz'):
 
     # Redshift distribution
     def p_z(z, H0, kappa=0):
+        # Fixed parameters are from the result of Planck 2018
         return CosmologicalParameters(H0/100., 0.315, 0.685, -1., 0., 0.).ComovingVolumeElement(z)*(1+z)**(kappa-1)
 
     bounds_dict = {
@@ -74,19 +75,32 @@ if not os.path.exists(outdir/f'multi/{param}_{method}.npz'):
             print("Invalid argument!")
             sys.exit(1)
     elif label == "real":
-        # Fixed parameters according to result from "Constraints on the Cosmic Expansion History from GWTC–3" (SNR > 10 & w0flatLCDM)
+        # Fixed parameters are the median values from the result of "Constraints on the Cosmic Expansion History from GWTC–3" (SNR > 10 & w0flatLCDM)
+        fixed_params = {
+            "alpha": 4.2487241670754035,
+            "mu": 31.825703461630482,
+            "sigma": 3.7904971258458042,
+            "w": 0.024059947239759398,
+            "delta": 4.8961636235644015,
+            "mmin": 5.0808821331157095,
+            "mmax": 109.03299036617125
+        }
         if param == "3":
             bounds = bounds_dict["H0"], bounds_dict["alpha"], bounds_dict["kappa"]
             def p_m_p_z(x):
-                return np.einsum("ij, j -> ij", plpeak(m, alpha=x[1], mu=31.825703461630482, sigma=3.7904971258458042, w=0.024059947239759398, delta=4.8961636235644015, mmin=5.0808821331157095, mmax=109.03299036617125), p_z(z, x[0], kappa=x[2]))
+                return np.einsum("ij, j -> ij", plpeak(m, alpha=x[1], mu=fixed_params["mu"], sigma=fixed_params["sigma"], w=fixed_params["w"], delta=fixed_params["delta"], mmin=fixed_params["mmin"], mmax=fixed_params["mmax"]), p_z(z, x[0], kappa=x[2]))
         elif param == "5c":
             bounds = bounds_dict["H0"], bounds_dict["alpha"], bounds_dict["mu"], bounds_dict["sigma"], bounds_dict["kappa"]
             def p_m_p_z(x):
[…]

### 2024-10-02 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 This is because the reconstruction is only valid in the region where observations are present as mentioned in Sec.~\ref{sec:reconstruction}.
-The optimization procedure can be performed in parallel for each reconstructed distribution, as each reconstructed distribution is independent.
-This reduces the runtime as long as computational resources are available.
 In Fig.~\ref{fig:simulation_comparison}, we show an illustration of the comparison between the two populations.
@@
 That is, the uncertainty in the comparison of the two populations is not considered in the resulting distribution.
-By considering this optimization procedure only, we can estimate the optimized parameters without performing computationally expensive likelihood calculations.
+
+The optimization procedure does not scale with the number of \ac{GW} events, as the optimization is performed for each reconstructed distribution.
+Therefore, the more \ac{GW} events we have, the longer the runtime of the reconstruction, but the runtime of the optimization remains the same.
+This is advantageous for performing the analysis with different models when more \ac{GW} events are available, as the reconstruction can be reused.
+Other than that, the optimization procedure can be performed in parallel for each reconstructed distribution, as each reconstructed distribution is independent.
+This reduces the runtime as long as computational resources are available.
 

### 2024-10-08 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 However, the intrinsic distribution is not known, and the method depends on the choice of the population model.
-Furthermore, the intrinsic distribution is degenerate with cosmology.
-For example, an intrinsic distribution with a linear redshift evolution is degenerate with $H_0$, as they both redshift the observed distribution in the same way.
+Furthermore, the intrinsic distribution is degenerate with cosmology: for example, an intrinsic distribution with a linear redshift evolution is degenerate with $H_0$, as they both redshift the observed distribution in the same way.
 While this example is not physical, this degeneracy still affects the inference of $H_0$.

### 2024-10-08 src/scripts/plot_real_result_multi.py
Update plots

@@ -1,13 +1,13 @@
 import numpy as np
-import matplotlib.pyplot as plt
 from corner import corner
-from figaro.cosmology import Planck18
 from figaro import plot_settings
 import paths
 
-f = np.load(paths.data / 'real/multi' / '4b_Powell.npz')
+f = np.load(paths.data / 'real/multi' / '3_Powell.npz')
 result = f['result']
-parameters = ["$H_0$", "$\\alpha$", "$\\mu$", "$\\sigma$"]
+result = result[result[:, 0] > 30 and result[:, 0] < 120]
+result = result[result[:, 2] > -5 and result[:, 2] < 5]
+parameters = ["$H_0$", "$\\alpha$", "$\\kappa$"]
 
 fig = corner(result, labels=parameters, color='steelblue', truth_color='red', levels = [0.3935, 0.9], plot_datapoints=False, plot_density=False, smooth=1, fill_contours=True, contourf_kwargs={'colors': ['white', 'darkturquoise', 'mediumturquoise'], 'alpha': [1, 0.2, 0.5]}, contour_kwargs={'linewidths': 1})
 fig.savefig(paths.figures / "real_result_multi.pdf", bbox_inches='tight')

### 2024-10-08 src/scripts/plot_simulation_result_multi.py

@@ -1,5 +1,4 @@
 import numpy as np
-import matplotlib.pyplot as plt
 from corner import corner
 from figaro.cosmology import Planck18
 from figaro import plot_settings

### 2024-10-08 src/scripts/plot_real_result_multi.py
Update plot_real_result_multi.py

@@ -5,8 +5,8 @@ import paths
 
 f = np.load(paths.data / 'real/multi' / '3_Powell.npz')
 result = f['result']
-result = result[result[:, 0] > 30 and result[:, 0] < 120]
-result = result[result[:, 2] > -5 and result[:, 2] < 5]
+result = result[(result[:, 0] > 30) & (result[:, 0] < 120)]
+result = result[(result[:, 2] > -5) & (result[:, 2] < 5)]
 parameters = ["$H_0$", "$\\alpha$", "$\\kappa$"]
 
 fig = corner(result, labels=parameters, color='steelblue', truth_color='red', levels = [0.3935, 0.9], plot_datapoints=False, plot_density=False, smooth=1, fill_contours=True, contourf_kwargs={'colors': ['white', 'darkturquoise', 'mediumturquoise'], 'alpha': [1, 0.2, 0.5]}, contour_kwargs={'linewidths': 1})

### 2024-10-08 src/scripts/plot_real_result_multi.py
Update plot_real_result_multi.py

@@ -6,7 +6,7 @@ import paths
 f = np.load(paths.data / 'real/multi' / '3_Powell.npz')
 result = f['result']
 result = result[(result[:, 0] > 30) & (result[:, 0] < 120)]
-result = result[(result[:, 2] > -5) & (result[:, 2] < 5)]
+result = result[result[:, 2] > 0]
 parameters = ["$H_0$", "$\\alpha$", "$\\kappa$"]
 
 fig = corner(result, labels=parameters, color='steelblue', truth_color='red', levels = [0.3935, 0.9], plot_datapoints=False, plot_density=False, smooth=1, fill_contours=True, contourf_kwargs={'colors': ['white', 'darkturquoise', 'mediumturquoise'], 'alpha': [1, 0.2, 0.5]}, contour_kwargs={'linewidths': 1})

### 2024-10-16 src/scripts/infer_multi.py
Update real multi plot

@@ -85,10 +85,22 @@ if not os.path.exists(outdir/f'multi/{param}_{method}.npz'):
             "mmin": 5.0808821331157095,
             "mmax": 109.03299036617125
         }
-        if param == "3":
+        if param == "3a":
             bounds = bounds_dict["H0"], bounds_dict["alpha"], bounds_dict["kappa"]
             def p_m_p_z(x):
                 return np.einsum("ij, j -> ij", plpeak(m, alpha=x[1], mu=fixed_params["mu"], sigma=fixed_params["sigma"], w=fixed_params["w"], delta=fixed_params["delta"], mmin=fixed_params["mmin"], mmax=fixed_params["mmax"]), p_z(z, x[0], kappa=x[2]))
+        elif param == "3b":
+            bounds = bounds_dict["H0"], bounds_dict["mu"], bounds_dict["kappa"]
+            def p_m_p_z(x):
+                return np.einsum("ij, j -> ij", plpeak(m, mu=x[1], alpha=fixed_params["alpha"], sigma=fixed_params["sigma"], w=fixed_params["w"], delta=fixed_params["delta"], mmin=fixed_params["mmin"], mmax=fixed_params["mmax"]), p_z(z, x[0], kappa=x[2]))
+        elif param == "4c":
+            bounds = bounds_dict["H0"], bounds_dict["mu"], bounds_dict["sigma"], bounds_dict["kappa"]
+            def p_m_p_z(x):
+                return np.einsum("ij, j -> ij", plpeak(m, alpha=fixed_params["alpha"], mu=x[1], sigma=x[2], w=fixed_params["w"], delta=fixed_params["delta"], mmin=fixed_params["mmin"], mmax=fixed_params["mmax"]), p_z(z, x[0], kappa=x[3]))
+        elif param == "4d":
+            bounds = bounds_dict["H0"], bounds_dict["alpha"], bounds_dict["mu"], bounds_dict["kappa"]
+            def p_m_p_z(x):
+                return np.einsum("ij, j -> ij", plpeak(m, alpha=x[1], mu=x[2], sigma=fixed_params["sigma"], w=fixed_params["w"], delta=fixed_params["delta"], mmin=fixed_params["mmin"], mmax=fixed_params["mmax"]), p_z(z, x[0], kappa=x[3]))
         elif param == "5c":
             bounds = bounds_dict["H0"], bounds_dict["alpha"], bounds_dict["mu"], bounds_dict["sigma"], bounds_dict["kappa"]
             def p_m_p_z(x):

### 2024-10-16 src/scripts/plot_real_result_multi.py

@@ -3,11 +3,11 @@ from corner import corner
 from figaro import plot_settings
 import paths
 
-f = np.load(paths.data / 'real/multi' / '3_Powell.npz')
+f = np.load(paths.data / 'real/multi' / '4d_Powell.npz')
 result = f['result']
-result = result[(result[:, 0] > 30) & (result[:, 0] < 120)]
-result = result[result[:, 2] > 0]
-parameters = ["$H_0$", "$\\alpha$", "$\\kappa$"]
+result = result[(result[:, 0] > 40) & (result[:, 0] < 100)]
+result = result[(result[:, 3] > 1) & (result[:, 3] < 9)]
+parameters = ["$H_0$", "$\\alpha$", "$\\mu$", "$\\kappa$"]
 
 fig = corner(result, labels=parameters, color='steelblue', truth_color='red', levels = [0.3935, 0.9], plot_datapoints=False, plot_density=False, smooth=1, fill_contours=True, contourf_kwargs={'colors': ['white', 'darkturquoise', 'mediumturquoise'], 'alpha': [1, 0.2, 0.5]}, contour_kwargs={'linewidths': 1})
 fig.savefig(paths.figures / "real_result_multi.pdf", bbox_inches='tight')

### 2024-10-16 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 The method works best when the intrinsic distribution is well understood.
-However, the intrinsic distribution is not known, and the method depends on the choice of the population model.
+However, the features present in the mass distribution are not so well characterized yet, and their astrophysical origin is a debated topic \citep{Mapelli:2020vfa, Mandel:2018hfr}.
+In this case, the spectral-siren method depends on the choice of the population model, which can introduce biases in the inference \citep{Mastrogiovanni:2021wsd, Pierra:2023deu}.
 Furthermore, the intrinsic distribution is degenerate with cosmology: for example, an intrinsic distribution with a linear redshift evolution is degenerate with $H_0$, as they both redshift the observed distribution in the same way.
 While this example is not physical, this degeneracy still affects the inference of $H_0$.
-Due to these issues, the spectral-siren method is not straightforward to apply, intrinsic population model parameters and cosmological parameters are required to be inferred simultaneously, which makes the method computationally expensive.
+Due to these issues, astrophysics and cosmology are tightly linked to each other, and a proper inference of the astrophysical and cosmological parameters should be done simultaneously, which makes the method computationally expensive.
 
-An alternative approach to the spectral-siren method is to use non-parametric methods to reconstruct the observed mass population distribution directly from the data.
+A possible alternative approach, which we will propose in this paper, is to use non-parametric methods to reconstruct the observed mass population distribution directly from the data.
 Non-parametric methods allow us to reconstruct arbitrary probability densities without making assumptions about the form of the distribution \citep{Rinaldi:2021bhm}.
@@
 We make use of the \ac{(H)DPGMM} developed in \cite{Rinaldi:2021bhm} as a non-parametric model to reconstruct the observed population distribution.
-Other assumptions including the population model, cosmological model and selection function are then incorporated in the second half, where we transform the source-frame population model to the detector frame with a given cosmological model.
+In the second part, we make use of all the other assumptions, namely the population model, cosmological model and selection function, to transform the source-frame population model into the detector frame.
 This separation allows us to explore the impact of different assumptions easily.
@@
 
-The selection effect is chosen to be considered in the transformation instead of the reconstruction, as the uncertainty in the reconstruction will diverge in the region where the selection function censoring the data.
+The selection effect is chosen to be considered in the transformation instead of the reconstruction, as the uncertainty in the reconstruction will diverge in the region where the selection function censors the data.
 For example, if we consider the selection function in the reconstruction, the uncertainty in the reconstruction will be large at the low mass end, as the selection function censors the low mass events effectively.
-\cite{Essick:2023upv} also pointed out that including the selection function in the reconstruction may lead to insignificant biases in the analysis.
+This way of including selection effects in the analysis, albeit simple and straightforward, is only an approximation of the real selection process, as pointed out in \cite{Essick:2023upv}.
+Nonetheless, the large uncertainties currently associated with \ac{GW} measurements are most likely to be dominant with respect to the bias induced by this approximation.
+Therefore, we opted for including the selection function in the transformation itself.
+We will explore an exact way of including selection effects in this method in a future paper.
 Therefore, we consider the selection function in the transformation to avoid these issues.
@@
 
-The optimization procedure does not scale with the number of \ac{GW} events, as the optimization is performed for each reconstructed distribution.
-Therefore, the more \ac{GW} events we have, the longer the runtime of the reconstruction, but the runtime of the optimization remains the same.
+The optimization procedure is highly efficient as it does not scale with the number of \ac{GW} events, since the optimization is performed for each reconstructed distribution independently.
+Therefore, even as the number of \ac{GW} events increases, the runtime of the optimization remains constant, making our method scalable and robust for future large datasets.
 This is advantageous for performing the analysis with different models when more \ac{GW} events are available, as the reconstruction can be reused.
@@
 Note that we marginalized over $q$ for the selection function, similar to the transformation in Sec.~\ref{sec:transformation}.
-After applying the selection function, there were \variable{output/n_obs_samples.txt} sets of samples left.
+After applying the selection function, which filters out the non-detectable events based on their $m^z_1$ and $d_L$, we retained \variable{output/n_obs_samples.txt} sets of samples for further analysis.
 
@@
 \end{figure}
-We find that the $H_0$ distribution, whose associated uncertainty comes from the reconstruction uncertainty, is consistent with the simulated $H_0$ value.
+The recovered $H_0$ distribution, whose uncertainty budget comes from the remapping of the uncertainty in the non-parametric reconstruction, is consistent with the simulated value.
 
@@
 Similar to the previous subsection, we assume perfect knowledge of both models, the selection function and all the parameters that are not objects of the inference.
-Other than $H_0$, we also chose the power law index $\alpha$, the peak mass $\mu$, and the peak width $\sigma$ of the \textsc{power law + peak} model as the free parameters.
-The minimum and maximum mass, and the range of mass tapering at the low mass end $\delta$ are fixed, as we may not be able to infer these parameters accurately with our method.
-This is because the reconstructions are only valid in the region where the samples are present, and they may be not accurate at the boundaries depending on the boundary features of the observed population distribution.
+Other than $H_0$, we included the power law index $\alpha$, the peak mass $\mu$, and the peak width $\sigma$ of the \textsc{power law + peak} model as the free parameters in the analysis.
+The minimum and maximum mass, and the range of mass tapering at the low mass end $\delta$ are fixed.
+These parameters, which control the boundary features of the observed population distribution, were chosen because the reconstructions are only valid in the region where the samples are present, and the reconstructed distributions may not be accurate at the boundaries depending on the boundary features of the observed population distribution.
 In principle, more parameters can be inferred with our method, for example, the peak weight $w$ in the mass population model and the present mass density $\Omega_m$ in the $\Lambda$CDM model.
@@
 \end{figure}
-We find that the distribution, whose associated uncertainty comes from the reconstruction uncertainty, is consistent with the simulated values.
+The recovered distribution, whose uncertainty budget comes from the remapping of the uncertainty in the non-parametric reconstruction, is consistent with the simulated values.
 Some features of the result are discussed together with the result of the real data in Sec.~\ref{sec:real_data}.
@@
 In this section, we apply our method to the real data from the \ac{LVK}.
-We included all $70$ events labeled as BBH in Table~I of \cite{KAGRA:2021duu}.
+We included all $70$ events labeled as BBH in Table~I of \cite{KAGRA:2021duu}, which are different from the events used in \cite{LIGOScientific:2021aug}.
 

### 2024-10-16 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 Other than $H_0$, we included the power law index $\alpha$, the peak mass $\mu$, and the peak width $\sigma$ of the \textsc{power law + peak} model as the free parameters in the analysis.
-The minimum and maximum mass, and the range of mass tapering at the low mass end $\delta$ are fixed.
+The minimum mass $m_\mathrm{min}$ and maximum mass $m_\mathrm{max}$, and the range of mass tapering at the low mass end $\delta$ are fixed.
 These parameters, which control the boundary features of the observed population distribution, were chosen because the reconstructions are only valid in the region where the samples are present, and the reconstructed distributions may not be accurate at the boundaries depending on the boundary features of the observed population distribution.
@@
 For the cosmological model, we fixed the parameters to the results from \cite{Planck:2018vyg} except $H_0$.
-For the population model, we fixed the minimum mass $m_\mathrm{min}$ and maximum mass $m_\mathrm{max}$, as well as the range of mass tapering at the low mass end to the results from \cite{LIGOScientific:2021aug}.
+For the population model, we fixed $\sigma$, $w$, $m_\mathrm{min}$, $m_\mathrm{max}$, and $\delta$ to the results from \cite{LIGOScientific:2021aug}.
 Note that the fixed parameters are not necessarily the true values, but they can be considered as a reasonable guess for the parameters.
@@
 \begin{table}[htbp]
-    \caption{Fixed parameters}
+    \caption{Fixed parameters of the $\Lambda$CDM model (upper) and the \textsc{power law + peak} model (lower)}
     \begin{tabular}{cc}
@@
         \midrule
-        $m_\mathrm{min}$ & $5.08$ \\
-        $m_\mathrm{max}$ & $109.03$ \\
-        $\delta$ & $4.89$ \\
-        $w$ & $0.02$ \\
+        % Cosmological model
+        $\Omega_m$ & $0.315$ \\
+        $\Omega_\Lambda$ & $0.685$ \\
+        $w$ & $-1$ \\
+        \midrule
+        % Population model
+        $\sigma$ & $3.7904971258458042$ \\
+        $w$ & $0.024059947239759398$ \\
+        $m_\mathrm{min}$ & $5.0808821331157095$ \\
+        $m_\mathrm{max}$ & $109.03299036617125$ \\
+        $\delta$ & $4.8961636235644015$ \\
         \botrule

### 2024-10-19 src/tex/ms.tex
[showyourwork] overleaf sync

@@
-\documentclass[sn-aps, pdflatex]{sn-jnl}
+\documentclass[sn-aps, pdflatex, iicol]{sn-jnl}
 
@@
 \usepackage{graphicx}
+\usepackage{mathtools}
 
@@
     We demonstrated how non-parametric methods, which are flexible models that can be used to agnostically reconstruct arbitrary probability densities, can be incorporated into this framework and leverage the detector-frame mass distribution to infer the cosmological parameters.
+    We tested our method with mock data and applied it to the real data from the LIGO-Virgo-KAGRA Collaboration.
 }
@@
 Another alternative method is to use population statistics to infer $H_0$ by marginalizing the redshift distribution of the sources.
-This method is known as the spectral-siren method \citep{You:2020wju, Mastrogiovanni:2021wsd, LIGOScientific:2021aug, Ezquiaga:2022zkx}.
+This method is known as the spectral-siren method \citep{Farr:2019twy, You:2020wju, Mastrogiovanni:2021wsd, LIGOScientific:2021aug, Ezquiaga:2022zkx}.
 
@@
 The method works best when the intrinsic distribution is well understood.
-However, the features present in the mass distribution are not so well characterized yet, and their astrophysical origin is a debated topic \citep{Mapelli:2020vfa, Mandel:2018hfr}.
-In this case, the spectral-siren method depends on the choice of the population model, which can introduce biases in the inference \citep{Mastrogiovanni:2021wsd, Pierra:2023deu}.
+However, the features present in the mass distribution are not so well characterized yet, and their astrophysical origin is a debated topic \citep{Zevin:2017evb, Mapelli:2020vfa, Zevin:2020gbd, Mandel:2018hfr, Marchant:2023wno}.
+In this case, the spectral-siren method depends on the choice of the population model, which can introduce biases in the inference \citep{Mastrogiovanni:2021wsd, Mukherjee:2021rtw, Pierra:2023deu}.
 Furthermore, the intrinsic distribution is degenerate with cosmology: for example, an intrinsic distribution with a linear redshift evolution is degenerate with $H_0$, as they both redshift the observed distribution in the same way.
@@
 
-This approach allows us to extract the information contained in the observed data without making assumptions about the population model, cosmological model, or selection function.
-The reconstructed distribution can act as an intermediate observation-driven result that encodes both the intrinsic population information and the cosmological information.
+This approach allows us to extract the information contained in the observed data without making assumptions about the population model, cosmological model, or selection function; at the same time, the reconstructed distribution can act as an intermediate observation-driven result that encodes both the intrinsic population information and the cosmological information.
 Making use of the reconstructed distribution, we can easily explore the effect of using different models, as the reconstructed distribution can be reused for different models.
@@
 
-Some other studies have made use of non-parametric methods with the spectral-siren framework.
+Other studies have made use of non-parametric methods with the spectral-siren framework \citep{Mandel:2018mve, Tiwari:2020vym, Rinaldi:2021bhm, Edelman:2021zkw, Sadiq:2021fin, Edelman:2022ydv, Callister:2023tgi, Ray:2023upk}.
+For example, \cite{Farah:2024xub} used a Gaussian process to model the intrinsic mass distribution, which allows for a flexible model that can capture the features of the intrinsic distribution.
 
@@
 In Sec.~\ref{sec:real_data}, we apply our method to the real data from the \ac{LVK}.
-In Sec.~\ref{sec:discussion}, we compare our results with other methods and discuss future directions.
-Finally, we conclude in Sec.~\ref{sec:conclusion} with a summary.
+Finally, we conclude in Sec.~\ref{sec:conclusion} with a summary and outlook.
 
@@
 
-Our method consists of three main steps: the non-parametric reconstruction of the detector-frame observed population, the transformation of the source-frame population model to the detector frame, and the \ac{PE}.
+Our method consists of three main steps: the non-parametric reconstruction of the detector-frame observed population, the transformation of the source-frame population model to the detector frame, and the remapping of the reconstructed distribution to the model parameters.
 
@@
 We use the \ac{(H)DPGMM} developed in \cite{Rinaldi:2021bhm} as a non-parametric model to reconstruct the observed $m^z_1$ population distribution $p(m^z_1|\mathbf{\Theta})$.
-The \ac{(H)DPGMM} is a hierarchy of Gaussian mixture models with a Dirichlet process prior probability distribution, which allows the model to be flexible to reconstruct arbitrary population distribution from individual-event posterior probability distribution samples.
+In a nutshell, (H)DPGMM approximates the underlying probability density with a (potentially infinite) Gaussian mixture model,
+\begin{equation}
+    p(m_1^z) \simeq \sum_j^\infty w_j \mathcal{N}(m_1^z|\mu_j,\sigma_j)\,,
+\end{equation}
+where the relative weights of the components are inferred using a Dirichlet process.
+In what follows, we will refer to the (H)DPGMM parameters collectively as $\Theta = \{\mathbf{w},\boldsymbol\mu,\boldsymbol\sigma\}$.
 To perform the reconstruction, we use the package \textsc{figaro}\footnote{\textsc{figaro} is publicly available at \url{https://github.com/sterinaldi/FIGARO} and via \texttt{pip}.} \citep{Rinaldi:2024eep}.
@@
 \begin{equation}
+    \mathtoolsset{multlined-width=0.89\displaywidth}
     \begin{aligned}
         &p(m^z_1|\Lambda, \Omega, \mathrm{det}) \\
-        &= \int p(m^z_1|m_1, z, \Lambda, \Omega, \mathrm{det}) p(m_1, z|\Lambda, \Omega, \mathrm{det}) \mathrm{d}m_1 \mathrm{d}z \\
-        &= \int p(m^z_1|m_1, z) \frac{p(\mathrm{det}|m_1, z, \Lambda, \Omega)p(m_1, z|\Lambda, \Omega)}{p(\mathrm{det}|\Lambda, \Omega)} \mathrm{d}m_1 \mathrm{d}z \\
-        &= \frac{1}{p(\mathrm{det}|\Lambda, \Omega)}\int \delta(m^z_1-m_1(1+z))p(\mathrm{det}|m_1, z, \Omega)p(m_1|\Lambda)p(z|\Lambda, \Omega) \mathrm{d}m_1 \mathrm{d}z \\
-        &\propto \int p\left(\mathrm{det}\middle|\frac{m^z_1}{1+z},z,\Omega\right)p\left(\frac{m^z_1}{1+z}\middle|\Lambda\right)p(z|\Lambda, \Omega) \mathrm{d}z\,,
+        &\begin{multlined}
+            = \int p(m^z_1|m_1, z, \Lambda, \Omega, \mathrm{det}) \\ \times p(m_1, z|\Lambda, \Omega, \mathrm{det}) \mathrm{d}m_1 \mathrm{d}z
+        \end{multlined} \\
+        &\begin{multlined}
+            = \int p(m^z_1|m_1, z) \\ \times \frac{p(\mathrm{det}|m_1, z, \Lambda, \Omega)p(m_1, z|\Lambda, \Omega)}{p(\mathrm{det}|\Lambda, \Omega)} \mathrm{d}m_1 \mathrm{d}z
+        \end{multlined} \\
+        &\begin{multlined}
+            = \frac{1}{p(\mathrm{det}|\Lambda, \Omega)}\int \delta(m^z_1-m_1(1+z)) \\ \times p(\mathrm{det}|m_1, z, \Omega)p(m_1|\Lambda)p(z|\Lambda, \Omega) \mathrm{d}m_1 \mathrm{d}z
+        \end{multlined}\\
+        &\begin{multlined}
+            \propto \int p\left(\mathrm{det}\middle|\frac{m^z_1}{1+z},z,\Omega\right) \\ \times p\left(\frac{m^z_1}{1+z}\middle|\Lambda\right)p(z|\Lambda, \Omega) \mathrm{d}z\,,
+        \end{multlined}
     \end{aligned}
@@
 
-The selection effect is included in the derivation in Eq.~\eqref{eq:transformation}, specifically, the first term in the final expression, which is the selection function.
-Note that the selection function in general is a function of the detector-frame mass $m^z_1$, luminosity distance $d_L$, and mass ratio $q$.
-In this work, we marginalized over $q$ to study the $m_1$ population distribution only.
+The selection effects are included in the derivation in Eq.~\eqref{eq:transformation} via the first term in the final expression, using the selection function.
+In general, the selection function is a function of all the intrinsic binary parameters: the most important, however, are the component masses and the luminosity distance $d_L$.
+In this work, since our intrinsic distribution includes only $m_1$ and $z$, we marginalized out the mass ratio dependence of the selection function assuming an intrinsic distribution $p(q) \propto q^{1.1}$.
 
-The selection effect is chosen to be considered in the transformation instead of the reconstruction, as the uncertainty in the reconstruction will diverge in the region where the selection function censors the data.
-For example, if we consider the selection function in the reconstruction, the uncertainty in the reconstruction will be large at the low mass end, as the selection function censors the low mass events effectively.
-This way of including selection effects in the analysis, albeit simple and straightforward, is only an approximation of the real selection process, as pointed out in \cite{Essick:2023upv}.
-Nonetheless, the large uncertainties currently associated with \ac{GW} measurements are most likely to be dominant with respect to the bias induced by this approximation.
-Therefore, we opted for including the selection function in the transformation itself.
-We will explore an exact way of including selection effects in this method in a future paper.
-Therefore, we consider the selection function in the transformation to avoid these issues.
+We opted for reconstructing the observed distribution with the non-parametric method and filtering the observed distribution rather than following the more standard procedure of agnostically reconstructing the intrinsic distribution and comparing it to our astrophysical model to circumvent the limits of non-parametric methods concerning extrapolation.
+If the selection function effectively censors one area of the parameter space, such as the low-mass, high-redshift binaries, a data-driven only method will not be able to characterize the distribution in that specific area, leading to diverging uncertainties in the reconstructed distribution and thus making the whole inference scheme pointless.
+Conversely, if we model the observed distribution and filter the intrinsic distribution model, we are still not able to constrain the population behavior in the censored region but the inference will yield some results.
+This way of including selection effects in the analysis, albeit simple and straightforward, is only an approximation of the real selection process, as pointed out in \cite{Essick:2023upv}: nonetheless, the large uncertainties currently associated with \ac{GW} measurements are most likely to be dominant with respect to the bias induced by this approximation.
+Therefore, we opted to include the selection function in the transformation itself, deferring the investigation of properly including selection effects in the reconstruction of the observed distribution in a future paper.
 
@@
 In contrast to the first part of the method, this part includes the assumptions about the population and cosmological model, as well as the selection function.
-After obtaining $p(m^z_1|\Lambda, \Omega, \mathrm{det})$, we can perform the \ac{PE} to infer the optimized $\Omega$ and $\Lambda$.
+After obtaining $p(m^z_1|\Lambda, \Omega, \mathrm{det})$, we can perform the remapping to obtain the optimized $\Omega$ and $\Lambda$.
 
-\subsection{Parameter estimation}
-\label{sec:pe}
+\subsection{Remapping to model parameters}
+\label{sec:remapping}
 
@@
 \begin{equation}
+    \mathtoolsset{multlined-width=0.89\displaywidth}
     \begin{aligned}
         &p(m^z_1|\Lambda, \Omega(H_0), \mathrm{det}) \\
-        &\propto \int p\left(\mathrm{det}\middle|\frac{m^z_1}{1+z},z,\Omega(H_0)\right)p\left(\frac{m^z_1}{1+z}\middle|\Lambda\right)p(z|\Lambda, \Omega(H_0)) \mathrm{d}z\,,
+        &\begin{multlined}
+            \propto \int p\left(\mathrm{det}\middle|\frac{m^z_1}{1+z},z,\Omega(H_0)\right) \\ \times p\left(\frac{m^z_1}{1+z}\middle|\Lambda\right)p(z|\Lambda, \Omega(H_0)) \mathrm{d}z\,,
+        \end{multlined}
     \end{aligned}
@@
 
-The result of the \ac{PE} is shown in Fig.~\ref{fig:simulation_result_H0}.
+The result of the remapping is shown in Fig.~\ref{fig:simulation_result_H0}.
 \begin{figure}[htbp]
@@
         The histogram shows the distribution of the optimized $H_0$.
-        The vertical blue line and shaded regions represent the median, 68\%, and 90\% credible intervals of the distribution.
+        The vertical blue line and shaded regions represent the median, 68\% and 90\% credible intervals of the distribution.
         The vertical orange line represents the true $H_0$ we used to generate the mock data.
@@
 \end{figure}
-The recovered $H_0$ distribution, whose uncertainty budget comes from the remapping of the uncertainty in the non-parametric reconstruction, is consistent with the simulated value.
+The recovered $H_0$ distribution, whose uncertainty budget comes from the remapping of the individual-\ac{PE} uncertainty in the non-parametric reconstruction, is consistent with the simulated value.
 
@@
 In this example, instead of transforming the source-frame population model to the detector frame for each $H_0$ and performing a grid search, we define a function that calculates $d_\mathrm{JS}$ for a given set of the free parameters.
-We then use the optimization algorithm mentioned in Sec.~\ref{sec:pe} to find the optimized free parameters that minimize $d_\mathrm{JS}$.
+We then use the optimization algorithm mentioned in Sec.~\ref{sec:remapping} to find the optimized free parameters that minimize $d_\mathrm{JS}$.
 The optimization method requires bounds for the free parameters, which we show in Table~\ref{tab:bounds}.
@@
 The initial guesses for the optimization are chosen randomly from a uniform distribution within the bounds.
-The result of the \ac{PE} is shown in Fig.~\ref{fig:simulation_result_multi}.
+The result of the remapping is shown in Fig.~\ref{fig:simulation_result_multi}.
 \begin{figure}[htbp]
@@
         The diagonal histograms show the marginalized distributions.
-        The off-diagonal contour plots show the joint distributions, where the contours represent the 39.35\%, 90\% credible intervals.
+        The off-diagonal contour plots show the joint distributions, where the contours represent the 39.35\% and 90\% credible intervals.
         The red lines and points represent the true values we used to generate the mock data.
@@
 \end{figure}
-The recovered distribution, whose uncertainty budget comes from the remapping of the uncertainty in the non-parametric reconstruction, is consistent with the simulated values.
-Some features of the result are discussed together with the result of the real data in Sec.~\ref{sec:real_data}.
+The recovered distribution, whose uncertainty budget comes from the remapping of the individual-\ac{PE} uncertainty in the non-parametric reconstruction, is consistent with the simulated values.
+
+This is an expected result, as the mock data is generated with the same population model and cosmological model we used in the analysis.
+We can also see that the uncertainty in $H_0$ is larger compare to the analysis in Sec.~\ref{sec:inference_H0}, as more parameters are inferred simultaneously.
+The uncertainty in $\sigma$ is relatively large, this may be due to the large uncertainty in the reconstruction at the gap between the low-mass and high-mass peaks, which can be seen in 
+Fig.~\ref{fig:simulation_reconstruction}.
 
@@
 We then perform the remaining steps similar to Sec.~\ref{sec:inference_multi}.
-The result of the \ac{PE} is shown in Fig.~\ref{fig:real_result_multi}.
+We set the bounds for $\kappa$ to $[-10, 10]$, and the bounds for other free parameters are the same as in Table~\ref{tab:bounds}.
+The result of the remapping is shown in Fig.~\ref{fig:real_result_multi}.
 \begin{figure}[htbp]
@@
         The diagonal histograms show the marginalized distributions.
-        The off-diagonal contour plots show the joint distributions, where the contours represent the 39.35\%, 90\% credible intervals.
+        The off-diagonal contour plots show the joint distributions, where the contours represent the 39.35\% and 90\% credible intervals.
     }
@@
 
-\section{Discussion}
-\label{sec:discussion}
-
-The result shows similar problems mentioned in \cite{Pierra:2023deu}.
+This analysis is a proof of concept of our method.
+We aim to demonstrate that our method can be applied to the real data.
+To perform a detailed analysis, a more comprehensive configuration that considers different population models, cosmological models, and selection functions is required.
+If a wrong population model is chosen, the result will be biased \citep{Mukherjee:2021rtw, Mastrogiovanni:2021wsd, Pierra:2023deu, LIGOScientific:2020kqk, KAGRA:2021duu}.
 
@@
 To further improve the method, there are several directions we can explore.
-First, as mentioned in Sec.~\ref{sec:pe}, our method is not a Bayesian inference, as we only consider the optimization of the parameters.
+First, as mentioned in Sec.~\ref{sec:remapping}, our method is not a Bayesian inference, as we only consider the optimization of the parameters.
 To perform a full Bayesian inference, we need to develop a likelihood function that considers the comparison of the two populations.
@@
 This will allow us to perform efficient Bayesian model selection, as different models can be tested without repeating the reconstruction of the observed population distribution.
-This is currently under development and will be presented in a future work.
+We will explore this possibility in a future work.
 
@@
 S.~R. acknowledges financial support from the European Research Council for the ERC Consolidator grant DEMOBLACK, under contract no. 770017. 
+We acknowledge the use of computing facilities supported by grants from the Croucher Innovation Award from the Croucher Foundation Hong Kong.
 This research made use of the bwForCluster Helix: the authors acknowledge support by the state of Baden-Württemberg through bwHPC and the German Research Foundation (DFG) through grant INST 35/1597-1 FUGG.

### 2024-10-19 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 In recent years, the precision of cosmological measurements has improved significantly, leading to the discovery of the Hubble tension, a discrepancy between the value of the Hubble constant $H_0$ inferred from the \ac{CMB} \citep{Planck:2018vyg} and local measurements \citep{Riess:2021jrx}.
-This tension has motivated the search for new methods to measure $H_0$ with high precision, and \ac{GW} sources have emerged as a promising tool for this purpose \citep{LIGOScientific:2017adf, LIGOScientific:2021aug, Ezquiaga:2022zkx}.
+This tension has motivated the search for new methods to measure $H_0$ with high precision, and \ac{GW} sources detected by the \ac{LVK} \citep{KAGRA:2013rdx, LIGOScientific:2014pky, VIRGO:2014yos, KAGRA:2020tym} have emerged as a promising tool for this purpose \citep{LIGOScientific:2017adf, LIGOScientific:2021aug, Ezquiaga:2022zkx}.
 

### 2024-10-21 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \usepackage{mathtools}
+\usepackage{orcidlink}
 
@@
 
-\author*[1]{Thomas C. K. Ng}\email{thomas.ng@link.cuhk.edu.hk}
-\author[2,3]{Stefano Rinaldi}\email{stefano.rinaldi@uni-heidelberg.de}
-\author[1]{Otto A. Hannuksela}\email{hannuksela@phy.cuhk.edu.hk}
+\author*[1]{Thomas~C.~K.~Ng\,\orcidlink{0000-0002-9491-1598}\,}\email{thomas.ng@link.cuhk.edu.hk}
+\author[2,3]{Stefano~Rinaldi\,\orcidlink{0000-0001-5799-4155}\,}\email{stefano.rinaldi@uni-heidelberg.de}
+\author[1]{Otto~A.~Hannuksela\,\orcidlink{0000-0002-3887-7137}\,}\email{hannuksela@phy.cuhk.edu.hk}
 
@@
 
-Other studies have made use of non-parametric methods with the spectral-siren framework \citep{Mandel:2018mve, Tiwari:2020vym, Rinaldi:2021bhm, Edelman:2021zkw, Sadiq:2021fin, Edelman:2022ydv, Callister:2023tgi, Ray:2023upk}.
+Other studies have made use of non-parametric methods for population studies \citep[e.g.,][]{Mandel:2018mve, Tiwari:2020vym, Rinaldi:2021bhm, Edelman:2021zkw, Sadiq:2021fin, Edelman:2022ydv, Callister:2023tgi, Ray:2023upk}, and now these methods are starting to be applied to the spectral siren method.
 For example, \cite{Farah:2024xub} used a Gaussian process to model the intrinsic mass distribution, which allows for a flexible model that can capture the features of the intrinsic distribution.
@@
 The result of the reconstruction is shown in Fig.~\ref{fig:simulation_reconstruction}.
-\begin{figure}[htbp]
+\begin{figure}
     \script{plot_simulation_reconstruction.py}
@@
     \caption{
-        In both panels, the red line is the histogram of the true $m^z_1$ of the simulated \ac{GW} events.
-        In the upper panel, the blue line is the median of the reconstructed distributions, and the shaded region represents the 68\% and 90\% credible intervals of the reconstructed distributions.
-        In the lower panel, the blue lines are the single-reconstructed distributions from a subset of all the reconstructed distributions.
-        By combining all the single reconstructed distributions, we obtain the distribution in the upper panel.
+        Top: non-parametric reconstruction of the observed distribution using 315 GW events (red histogram).
+The blue line marks the median of the reconstructed distributions, and he shaded regions represent the corresponding 68\% and 90\% credible intervals.
+        Bottom: subset of individual draws for the non-parametric reconstruction.
     }
@@
 The result of the transformation is shown in Fig.~\ref{fig:simulation_transformation}.
-\begin{figure}[htbp]
+\begin{figure}
     \script{plot_simulation_transformation.py}
@@
     \caption{
-        The upper panel shows the source-frame $m_1$ population distribution, which we chose as the true population model.
-        The lower panel shows the transformed detector-frame $m^z_1$ population distributions for different $H_0$, where the corresponding $H_0$ is shown in the legend.
-        The population distribution is redshifted towards higher masses as $H_0$ increases, as $m^z_1 = m_1(1+z)$ and $z$ increases with $H_0$ for a given $d_L$.
-        The low-mass peak is suppressed due to the selection function.
+        Top: source-frame $m_1$ population distribution.
+        Bottom: detector-frame $m^z_1$ observed distribution for different $H_0$ values. The low-mass peak is suppressed due to the selection function.
     }
@@
 Figure~\ref{fig:simulation_comparison} shows an illustration of the comparison between the reconstructed observed population and the transformed population for different $H_0$.
-\begin{figure}[htbp]
+\begin{figure}
     \script{plot_simulation_comparison.py}
@@
     \caption{
-        The blue line and bands are the ones from Fig.~\ref{fig:simulation_reconstruction}, which represent the reconstructed observed population.
-        The other colored lines are the ones from Fig.~\ref{fig:simulation_transformation}, which represent the transformed population for different $H_0$.
-        The shaded regions represent the boundaries for the calculation of $d_\mathrm{JS}$, i.e., the $d_\mathrm{JS}$ calculation only considers the non-shaded region.
+        Comparison between the reconstructed observed distribution (blue line and shaded areas) and the predicted observed distribution with different $H_0$ choices.
+The gray shaded areas marks the boundaries for the $d_\mathrm{JS}$ calculation.
     }
@@
 The result of the remapping is shown in Fig.~\ref{fig:simulation_result_H0}.
-\begin{figure}[htbp]
+\begin{figure}
     \script{plot_simulation_result_H0.py}
@@
     \caption{
-        The histogram shows the distribution of the optimized $H_0$.
-        The vertical blue line and shaded regions represent the median, 68\% and 90\% credible intervals of the distribution.
-        The vertical orange line represents the true $H_0$ we used to generate the mock data.
+        Optimized $H_0$ posterior distribution.
+        The vertical blue line and shaded areas represent the median, 68\% and 90\% credible intervals respectively.
     }
@@
 \begin{table}[htbp]
-    \caption{Bounds for the free parameters}
+    \caption{Free parameter bounds for the joint inference presented in Section~\ref{sec:inference_multi}.}
     \begin{tabular}{cc}
@@
 The result of the remapping is shown in Fig.~\ref{fig:simulation_result_multi}.
-\begin{figure}[htbp]
+\begin{figure}
     \script{plot_simulation_result_multi.py}
@@
     \caption{
-        The corner plot shows the distribution of the optimized parameters.
-        The diagonal histograms show the marginalized distributions.
-        The off-diagonal contour plots show the joint distributions, where the contours represent the 39.35\% and 90\% credible intervals.
-        The red lines and points represent the true values we used to generate the mock data.
+        Optimised parameter distribution for the mock data inference. Simulated values are marked in red.
     }
@@
 The result of the reconstruction is shown in Fig.~\ref{fig:real_reconstruction}.
-\begin{figure}[htbp]
+\begin{figure}
     \script{plot_real_reconstruction.py}
@@
     \caption{
-        The red line is the histogram of the median of the $m^z_1$ posterior probability distribution samples of the \ac{GW} events.
-        The blue line is the median of the reconstructed distributions, and the shaded region represents the 68\% and 90\% credible intervals of the reconstructed distributions.
+        Median non-parametric reconstruction (blue line) and median values of the $m^z_1$ posterior probability distribution samples of the \ac{GW} events (red histogram).
+The shaded regions represent the 68\% and 90\% credible intervals for the non-parametric reconstruction.
     }
@@
 \begin{table}[htbp]
-    \caption{Fixed parameters of the $\Lambda$CDM model (upper) and the \textsc{power law + peak} model (lower)}
-    \begin{tabular}{cc}
+    \caption{Fixed parameters of the $\Lambda$CDM model (left) and the \textsc{power law + peak} model (right).}
+    \begin{tabular}{cc|cc}
         \toprule
-        Parameter & Value \\
+        Parameter & Value & Parameter & Value  \\
         \midrule
-        % Cosmological model
-        $\Omega_m$ & $0.315$ \\
-        $\Omega_\Lambda$ & $0.685$ \\
-        $w$ & $-1$ \\
-        \midrule
-        % Population model
-        $\sigma$ & $3.7904971258458042$ \\
-        $w$ & $0.024059947239759398$ \\
-        $m_\mathrm{min}$ & $5.0808821331157095$ \\
-        $m_\mathrm{max}$ & $109.03299036617125$ \\
-        $\delta$ & $4.8961636235644015$ \\
+         & & $w$ & $0.024$ \\
+        $\Omega_m$ & $0.315$ & $\sigma$ & $3.8\ \mathrm{M}_\odot$\\
+        $\Omega_\Lambda$ & $0.685$ & $m_\mathrm{min}$ & $5\ \mathrm{M}_\odot$ \\
+        $w$ & $-1$ & $m_\mathrm{max}$ & $109\ \mathrm{M}_\odot$ \\
+         & & $\delta$ & $4.9\ \mathrm{M}_\odot$\\
         \botrule
@@
 We set the bounds for $\kappa$ to $[-10, 10]$, and the bounds for other free parameters are the same as in Table~\ref{tab:bounds}.
-The result of the remapping is shown in Fig.~\ref{fig:real_result_multi}.
-\begin{figure}[htbp]
+The result of the remapping is shown in Fig.~\ref{fig:real_result_multi}, and it is consistent with the posterior distribution reported in \cite{KAGRA:2021duu}.
+\begin{figure}
     \script{plot_real_result_multi.py}
@@
     \caption{
-        The corner plot shows the distribution of the optimized parameters.
-        The diagonal histograms show the marginalized distributions.
-        The off-diagonal contour plots show the joint distributions, where the contours represent the 39.35\% and 90\% credible intervals.
+        Optimised parameter distribution for the real data inference.
     }
@@
 
-This analysis is a proof of concept of our method.
-We aim to demonstrate that our method can be applied to the real data.
-To perform a detailed analysis, a more comprehensive configuration that considers different population models, cosmological models, and selection functions is required.
-If a wrong population model is chosen, the result will be biased \citep{Mukherjee:2021rtw, Mastrogiovanni:2021wsd, Pierra:2023deu, LIGOScientific:2020kqk, KAGRA:2021duu}.
+This analysis we report here is a proof of concept of the method presented in this paper, and we simply aim at demonstrating that our method can be applied to the real data with relatively little effort.
+To perform a detailed analysis, a more comprehensive configuration that considers different population models, cosmological models, and selection functions is required, as well as a more robust optimization scheme to include more astrophysical parameters.
 
@@
 With the numbers of observed \ac{GW} events that are expected to grow larger and larger, potentially hitting the thousand mark by the end of the fifth observing run, performing a complete analysis from scratch with different astrophysical models will soon become too computationally expensive.
-The method we propose in this paper can help mitigate this issue by providing a fast and efficient way to explore different models before performing the full hierarchical Bayesian inference.
+The method we propose in this paper can help mitigate this issue by providing a fast and efficient way to explore different models before performing the full hierarchical Bayesian inference, especially keeping in mind that, if a wrong population model is chosen, the inference result will be biased \citep{Mukherjee:2021rtw, Mastrogiovanni:2021wsd, Pierra:2023deu, LIGOScientific:2020kqk, KAGRA:2021duu}.
 This makes our method a powerful diagnostic tool for model selection.

### 2024-10-29 src/tex/ms.tex
[showyourwork] overleaf sync

@@
     We demonstrated how non-parametric methods, which are flexible models that can be used to agnostically reconstruct arbitrary probability densities, can be incorporated into this framework and leverage the detector-frame mass distribution to infer the cosmological parameters.
-    We tested our method with mock data and applied it to the real data from the LIGO-Virgo-KAGRA Collaboration.
+    We tested our method with mock data and applied it to $70$ real events from the third gravitational-wave transient catalog of the LIGO-Virgo-KAGRA Collaboration.
 }
@@
     \caption{
-        Optimised parameter distribution for the mock data inference. Simulated values are marked in red.
+        Optimized parameter distribution for the mock data inference. Simulated values are marked in red.
     }
@@
 In this section, we apply our method to the real data from the \ac{LVK}.
-We included all $70$ events labeled as BBH in Table~I of \cite{KAGRA:2021duu}, which are different from the events used in \cite{LIGOScientific:2021aug}.
+We included all $70$ events labeled as BBH in Table~I of \cite{KAGRA:2021duu}, these correspond to all the events with $\mathrm{FAR_{min}} < 1\ \mathrm{yr}^{-1}$, where $\mathrm{FAR_{min}}$ is the smallest false alarm rate over all search pipelines.
+These events are different from the events used in \cite{LIGOScientific:2021aug}, therefore, the results are not directly comparable.
 
@@
     \caption{
-        Optimised parameter distribution for the real data inference.
+        Optimized parameter distribution for the real data inference.
     }
@@
 \end{figure}
+The result implies that \output{real_H0.txt}, which has an uncertainty much smaller than the one reported in \cite{KAGRA:2021duu}.
+This is because the uncertainty only includes the uncertainty from the reconstruction, which underestimates the true uncertainty which also includes the mismatch between the population model and the observed population distribution.
 
@@
 
-\bibliography{bib}
+\bibliography{bib.bib}
 

### 2024-10-29 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \end{figure}
-The result implies that \output{real_H0.txt}, which has an uncertainty much smaller than the one reported in \cite{KAGRA:2021duu}.
+The result implies that \variable{output/real_H0.txt}, which has an uncertainty much smaller than the one reported in \cite{KAGRA:2021duu}.
 This is because the uncertainty only includes the uncertainty from the reconstruction, which underestimates the true uncertainty which also includes the mismatch between the population model and the observed population distribution.
 
-This analysis we report here is a proof of concept of the method presented in this paper, and we simply aim at demonstrating that our method can be applied to the real data with relatively little effort.
+This analysis we report here is a proof of concept of the method presented in this paper, and we simply aim to demonstrate that our method can be applied to real data with relatively little effort.
 To perform a detailed analysis, a more comprehensive configuration that considers different population models, cosmological models, and selection functions is required, as well as a more robust optimization scheme to include more astrophysical parameters.

### 2024-10-29 src/tex/ms.tex
[showyourwork] overleaf sync

@@
         Top: non-parametric reconstruction of the observed distribution using 315 GW events (red histogram).
-The blue line marks the median of the reconstructed distributions, and he shaded regions represent the corresponding 68\% and 90\% credible intervals.
+The blue line marks the median of the reconstructed distributions, and the shaded regions represent the corresponding 68\% and 90\% credible intervals.
         Bottom: subset of individual draws for the non-parametric reconstruction.
@@
         Comparison between the reconstructed observed distribution (blue line and shaded areas) and the predicted observed distribution with different $H_0$ choices.
-The gray shaded areas marks the boundaries for the $d_\mathrm{JS}$ calculation.
+The gray shaded areas mark the boundaries for the $d_\mathrm{JS}$ calculation.
     }
@@
 This is an expected result, as the mock data is generated with the same population model and cosmological model we used in the analysis.
-We can also see that the uncertainty in $H_0$ is larger compare to the analysis in Sec.~\ref{sec:inference_H0}, as more parameters are inferred simultaneously.
+We can also see that the uncertainty in $H_0$ is larger compared to the analysis in Sec.~\ref{sec:inference_H0}, as more parameters are inferred simultaneously.
 The uncertainty in $\sigma$ is relatively large, this may be due to the large uncertainty in the reconstruction at the gap between the low-mass and high-mass peaks, which can be seen in 
@@
 In this section, we apply our method to the real data from the \ac{LVK}.
-We included all $70$ events labeled as BBH in Table~I of \cite{KAGRA:2021duu}, these correspond to all the events with $\mathrm{FAR_{min}} < 1\ \mathrm{yr}^{-1}$, where $\mathrm{FAR_{min}}$ is the smallest false alarm rate over all search pipelines.
+We included all $70$ events labeled as BBH in Table~I of \cite{KAGRA:2021duu}, these correspond to all the events with $\mathrm{FAR_{min}} < 1\ \mathrm{yr}^{-1}$, where $\mathrm{FAR_{min}}$ is the smallest false alarm rate among all search pipelines.
 These events are different from the events used in \cite{LIGOScientific:2021aug}, therefore, the results are not directly comparable.
@@
 \end{figure}
-The result implies that \variable{output/real_H0.txt}, which has an uncertainty much smaller than the one reported in \cite{KAGRA:2021duu}.
-This is because the uncertainty only includes the uncertainty from the reconstruction, which underestimates the true uncertainty which also includes the mismatch between the population model and the observed population distribution.
+The result implies that \variable{output/real_H0.txt}, which has an uncertainty much smaller than $H_0 = 50^{+37}_{-30}$ reported in \cite{KAGRA:2021duu}.
+This is mainly because the uncertainty only includes the uncertainty from the reconstruction, which underestimates the true uncertainty which also includes the mismatch between the population model and the observed population distribution.
 
@@
 This will allow us to perform efficient Bayesian model selection, as different models can be tested without repeating the reconstruction of the observed population distribution.
-We will explore this possibility in a future work.
+We will explore this possibility in future work.
 

### 2024-10-29 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \end{figure}
-The result implies that \variable{output/real_H0.txt}, which has an uncertainty much smaller than $H_0 = 50^{+37}_{-30}$ reported in \cite{KAGRA:2021duu}.
-This is mainly because the uncertainty only includes the uncertainty from the reconstruction, which underestimates the true uncertainty which also includes the mismatch between the population model and the observed population distribution.
+
+The result implies that \variable{output/real_H0.txt}, with symmetric $90\%$-credible intervals around the median.
+The uncertainty is much smaller than the one reported in \cite{KAGRA:2021duu} ($H_0 = 50^{+37}_{-30} \, \mathrm{km/s/Mpc}$), mainly because it only includes the uncertainty from the reconstruction, which underestimates the true uncertainty which also includes the mismatch between the population model and the observed population distribution.
 

### 2024-10-29 src/scripts/get_real_H0.py
Create get_real_H0.py

@@ -1,0 +1,13 @@
+import numpy as np
+import paths
+
+f = np.load(paths.data / 'real/multi' / '4d_Powell.npz')
+result = f['result']
+result = result[(result[:, 0] > 40) & (result[:, 0] < 100)]
+H0 = result[:, 0]
+med = np.median(H0)
+low = np.percentile(H0, 10)
+high = np.percentile(H0, 90)
+
+with open(paths.output/'real_H0.txt', 'w') as f:
+    f.write(f'$H_0 = {med:.2f}^{{+{high - med:.2f}}}_{{-{med - low:.2f}}} \, \mathrm{{km/s/Mpc}}$')

### 2024-10-29 src/tex/ms.tex
[showyourwork] overleaf sync

@@
     We demonstrated how non-parametric methods, which are flexible models that can be used to agnostically reconstruct arbitrary probability densities, can be incorporated into this framework and leverage the detector-frame mass distribution to infer the cosmological parameters.
-    We tested our method with mock data and applied it to $70$ real events from the third gravitational-wave transient catalog of the LIGO-Virgo-KAGRA Collaboration.
+    We tested our method with mock data and applied it to $70$ binary black hole mergers from the third gravitational-wave transient catalog of the LIGO-Virgo-KAGRA Collaboration.
 }
@@
     \acro{BBH}{binary-black-hole}
+    \acro{FAR}{false alarm rate}
 \end{acronym}
@@
 
-In this section, we apply our method to the real data from the \ac{LVK}.
-We included all $70$ events labeled as BBH in Table~I of \cite{KAGRA:2021duu}, these correspond to all the events with $\mathrm{FAR_{min}} < 1\ \mathrm{yr}^{-1}$, where $\mathrm{FAR_{min}}$ is the smallest false alarm rate among all search pipelines.
-These events are different from the events used in \cite{LIGOScientific:2021aug}, therefore, the results are not directly comparable.
+In this section, we apply our method to the \ac{GW} events released by the \ac{LVK} as part of the GWTC-2.1 \cite{LIGOScientific:2021usb} and GWTC-3 \cite{KAGRA:2021vkt} catalogs.
+We included all $70$ events labeled as BBH in Table~I of \cite{KAGRA:2021duu}, these correspond to all the events with $\mathrm{FAR_{min}} < 1\, \mathrm{yr}^{-1}$, where $\mathrm{FAR_{min}}$ is the smallest \ac{FAR} among all search pipelines.
+These events are different from the events used in \cite{LIGOScientific:2021aug}, where the authors use a \ac{FAR} threshold of $\mathrm{FAR_{min}} < 0.25\, \mathrm{yr}^{-1}$, therefore the results we present here might be slightly different from the ones in \cite{LIGOScientific:2021aug}.
+We opted for a larger \ac{FAR} threshold -- consistent, however, with the one used in population studies such as \cite{KAGRA:2021duu} -- to have a larger number of events to include in the non-parametric analysis.
 
@@
 \end{figure}
-
 The result implies that \variable{output/real_H0.txt}, with symmetric $90\%$-credible intervals around the median.
 The uncertainty is much smaller than the one reported in \cite{KAGRA:2021duu} ($H_0 = 50^{+37}_{-30} \, \mathrm{km/s/Mpc}$), mainly because it only includes the uncertainty from the reconstruction, which underestimates the true uncertainty which also includes the mismatch between the population model and the observed population distribution.
+Figure \ref{fig:real_result_multi} shows a peak in the marginalized posterior distribution of $H_0$ around $70\, \mathrm{km/s/Mpc}$.
+This means a significant fraction of the reconstructed observed population distribution has an optimized $H_0$ around $70\, \mathrm{km/s/Mpc}$, which again does not include the uncertainty from the mismatch between the population model and the observed population distribution.
+The peak should be interpreted as an artifact of the optimization procedure.
 
@@
 
+We thank Grégoire Pierra for helpful discussions.
 T.~N. and S.~R. acknowledge financial support from the German Excellence Strategy via the Heidelberg Cluster of Excellence (EXC 2181 - 390900948) STRUCTURES.

### 2024-10-30 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 However, most \ac{GW} sources do not have \ac{EM} counterparts.
-In this case, one can assume \ac{GW} sources are in galaxies and associate the redshift of the galaxy to the \ac{GW} source \citep{Schutz:1986gp, DelPozzo:2011vcw, Gray:2019ksv, Gray:2023wgj}.
+In this case, one can assume \ac{GW} sources are in galaxies and associate the redshift of the galaxy to the \ac{GW} source \citep{Schutz:1986gp, DelPozzo:2011vcw, Gray:2019ksv, Mukherjee:2020hyn, Mukherjee:2022afz, Gray:2023wgj}.
 Another alternative method is to use population statistics to infer $H_0$ by marginalizing the redshift distribution of the sources.
-This method is known as the spectral-siren method \citep{Farr:2019twy, You:2020wju, Mastrogiovanni:2021wsd, LIGOScientific:2021aug, Ezquiaga:2022zkx}.
+This method is known as the spectral-siren method \citep{Farr:2019twy, You:2020wju, Mastrogiovanni:2021wsd, LIGOScientific:2021aug, Ezquiaga:2022zkx, Karathanasis:2022rtr}.
 
@@
 
-Other studies have made use of non-parametric methods for population studies \citep[e.g.,][]{Mandel:2018mve, Tiwari:2020vym, Rinaldi:2021bhm, Edelman:2021zkw, Sadiq:2021fin, Edelman:2022ydv, Callister:2023tgi, Ray:2023upk}, and now these methods are starting to be applied to the spectral siren method.
-For example, \cite{Farah:2024xub} used a Gaussian process to model the intrinsic mass distribution, which allows for a flexible model that can capture the features of the intrinsic distribution.
+Other studies have made use of non-parametric methods for population studies \citep[e.g.,][]{Mandel:2016prl, Mandel:2018mve, Tiwari:2020vym, Rinaldi:2021bhm, Edelman:2021zkw, Sadiq:2021fin, Edelman:2022ydv, Callister:2023tgi, Ray:2023upk}, and now these methods are starting to be applied to the spectral siren method.
+For example, \cite{Farah:2024xub} and \cite{MaganaHernandez:2024uty} used different non-parametric models to model the intrinsic mass distribution, which allows for a flexible model that can capture the features of the intrinsic distribution.
 
@@
 
-The analysis code we developed to implement the method presented in this section is publicly available on GitHub\footnote{ \url{https://github.com/thomasckng/det-frame-cosmo-with-FIGARO/tree/main/src/scripts}}.
+The analysis code we developed to implement the method presented in this section and the results of the following sections are publicly available on GitHub\footnote{\url{https://github.com/thomasckng/det-frame-cosmo-with-FIGARO/tree/main/src/scripts}} and Zenodo \citep{ng_2024_13968239} respectively.
 
@@
 
-In this section, we apply our method to the \ac{GW} events released by the \ac{LVK} as part of the GWTC-2.1 \cite{LIGOScientific:2021usb} and GWTC-3 \cite{KAGRA:2021vkt} catalogs.
+In this section, we apply our method to the \ac{GW} events released by the \ac{LVK} as part of the GWTC-2.1 \cite{LIGOScientific:2021usb, ligo_scientific_collaboration_and_virgo_2022_6513631} and GWTC-3 \cite{KAGRA:2021vkt, ligo_scientific_collaboration_and_virgo_2023_8177023} catalogs.
 We included all $70$ events labeled as BBH in Table~I of \cite{KAGRA:2021duu}, these correspond to all the events with $\mathrm{FAR_{min}} < 1\, \mathrm{yr}^{-1}$, where $\mathrm{FAR_{min}}$ is the smallest \ac{FAR} among all search pipelines.

### 2024-12-16 src/scripts/plot_real_result_multi.py
Squashed commit of the following:
commit 7459ab344058135e96c22c08621a53d7b959c2bb
Author: Thomas Ng <thomas.ng@link.cuhk.edu.hk>
Date:   Fri Dec 13 09:36:28 2024 +0800

    Use sandbox zenodo

commit e7abfd68d13634ddcbbba56ff46e80079f16fcd3
Author: Thomas Ng <thomas.ng@link.cuhk.edu.hk>
Date:   Fri Dec 13 09:27:18 2024 +0800

    Modify plots

commit 6a39c4ba584f627ba02dab9580efff3bb5c22cae
Author: Thomas Ng <thomas.ng@link.cuhk.edu.hk>
Date:   Wed Dec 11 14:58:28 2024 +0800

    npz to npy

commit 8c5830b8d5180f3e430dc40d6f10566923b7f883
Author: Thomas Ng <thomas.ng@link.cuhk.edu.hk>
Date:

@@ -3,11 +3,10 @@ from corner import corner
 from figaro import plot_settings
 import paths
 
-f = np.load(paths.data / 'real/multi' / '4d_Powell.npz')
+f = np.load(paths.data / 'real/multi' / '4c_Powell.npz')
 result = f['result']
-result = result[(result[:, 0] > 40) & (result[:, 0] < 100)]
-result = result[(result[:, 3] > 1) & (result[:, 3] < 9)]
-parameters = ["$H_0$", "$\\alpha$", "$\\mu$", "$\\kappa$"]
+result = result[(result[:, 3] > -5) & (result[:, 3] < 15)] # plot limits
+parameters = ["$H_0$", "$\\mu$", "$\\sigma$", "$\\kappa$"]
 
-fig = corner(result, labels=parameters, color='steelblue', truth_color='red', levels = [0.3935, 0.9], plot_datapoints=False, plot_density=False, smooth=1, fill_contours=True, contourf_kwargs={'colors': ['white', 'darkturquoise', 'mediumturquoise'], 'alpha': [1, 0.2, 0.5]}, contour_kwargs={'linewidths': 1})
+fig = corner(result, labels=parameters, color='steelblue', truth_color='red', levels = [0.3935, 0.90], plot_datapoints=False, plot_density=False, smooth=1, fill_contours=True, contourf_kwargs={'colors': ['white', 'darkturquoise', 'mediumturquoise'], 'alpha': [1, 0.2, 0.5]}, contour_kwargs={'linewidths': 1})
 fig.savefig(paths.figures / "real_result_multi.pdf", bbox_inches='tight')

### 2024-12-16 src/scripts/plot_simulation_result_H0.py

@@ -14,7 +14,7 @@ ax.axvline(true_H0, lw = 0.7, ls = '--', c = 'red', label = f'$H_0={true_H0:.1f}
 ax.axvline(percs[2], c = 'steelblue', lw=0.7, label = f'$H_0={percs[2]:.1f}^{{+{percs[3]-percs[2]:.1f}}}_{{-{percs[2]-percs[1]:.1f}}}$')
 ax.axvspan(percs[1], percs[3], alpha=0.35, color='mediumturquoise')
 ax.axvspan(percs[0], percs[4], alpha=0.2, color='darkturquoise')
-ax.legend()
+ax.legend(loc='upper left')
 ax.set_xlabel('$H_0\ [\mathrm{km/s/Mpc}]$')
 ax.set_ylabel('$\mathrm{Density}$')
 ax.set_xlim(percs[2]-30, percs[2]+30)

### 2024-12-16 src/scripts/plot_simulation_result_multi.py

@@ -6,10 +6,8 @@ import paths
 
 f = np.load(paths.data / 'simulation/multi' / '4b_Powell.npz')
 result = f['result']
-result = result[result[:, 0] < 200]
-result = result[result[:, 3] < 9]
 simulated_truth = {'$H_0$': Planck18.h*100, '$\\alpha$': 3.5, '$\\mu$': 35, '$\\sigma$': 5, '$\\delta$': 5, '$w$': 0.2, '$m_\mathrm{min}$': 5, '$m_\mathrm{max}$': 90}
 parameters = ["$H_0$", "$\\alpha$", "$\\mu$", "$\\sigma$"]
 
-fig = corner(result, labels=parameters, truths=[simulated_truth[param] for param in parameters], color='steelblue', truth_color='red', levels = [0.3935, 0.9], plot_datapoints=False, plot_density=False, smooth=1, fill_contours=True, contourf_kwargs={'colors': ['white', 'darkturquoise', 'mediumturquoise'], 'alpha': [1, 0.2, 0.5]}, contour_kwargs={'linewidths': 1})
+fig = corner(result, labels=parameters, truths=[simulated_truth[param] for param in parameters], color='steelblue', truth_color='red', levels = [0.3935, 0.95], plot_datapoints=False, plot_density=False, smooth=1, fill_contours=True, contourf_kwargs={'colors': ['white', 'darkturquoise', 'mediumturquoise'], 'alpha': [1, 0.2, 0.5]}, contour_kwargs={'linewidths': 1})
 fig.savefig(paths.figures / "simulation_result_multi.pdf", bbox_inches='tight')

### 2024-12-16 src/tex/ms.tex
[showyourwork] overleaf sync

@@
     The challenge of understanding the Universe's dynamics, particularly the Hubble tension, requires precise measurements of the Hubble constant.
-    Building upon the existing spectral-siren method, which capitalizes on population information from gravitational-wave sources, this paper explores an alternative way to analyze the population data to obtain the cosmological parameters in $\Lambda$CDM.
-    We demonstrated how non-parametric methods, which are flexible models that can be used to agnostically reconstruct arbitrary probability densities, can be incorporated into this framework and leverage the detector-frame mass distribution to infer the cosmological parameters.
-    We tested our method with mock data and applied it to $70$ binary black hole mergers from the third gravitational-wave transient catalog of the LIGO-Virgo-KAGRA Collaboration.
+    Building upon the existing spectral-siren method, which capitalizes on population information from gravitational-wave sources, this paper explores an alternative method to analyze the population data to obtain the cosmological parameters in $\Lambda$CDM.
+    We demonstrate how non-parametric methods, which are flexible models that can be used to agnostically reconstruct arbitrary probability densities, can be incorporated into this framework and leverage the detector-frame mass distribution to infer the cosmological parameters.
+    We test our method with mock data and apply it to $70$ binary black hole mergers from the third gravitational-wave transient catalog of the LIGO-Virgo-KAGRA Collaboration.
 }
@@
 Some \ac{GW} sources can be associated with \ac{EM} counterparts, allowing for the measurement of the redshift, e.g., GW170817 \citep{LIGOScientific:2017adf, Guidorzi:2017ogy}.
-However, most \ac{GW} sources do not have \ac{EM} counterparts.
-In this case, one can assume \ac{GW} sources are in galaxies and associate the redshift of the galaxy to the \ac{GW} source \citep{Schutz:1986gp, DelPozzo:2011vcw, Gray:2019ksv, Mukherjee:2020hyn, Mukherjee:2022afz, Gray:2023wgj}.
-Another alternative method is to use population statistics to infer $H_0$ by marginalizing the redshift distribution of the sources.
+However, most currently detectable \ac{GW} sources do not have \ac{EM} counterparts.
+For these "dark sirens," one can assume \ac{GW} sources are in galaxies and associate the redshift of the galaxy to the \ac{GW} source, enabling an alternative way to measure cosmology \citep{Schutz:1986gp, DelPozzo:2011vcw, Gray:2019ksv, Mukherjee:2020hyn, Mukherjee:2022afz, Gray:2023wgj}.
+Yet another method is to use physically motivated population models to infer $H_0$ by marginalizing over the redshift distribution of the sources.
 This method is known as the spectral-siren method \citep{Farr:2019twy, You:2020wju, Mastrogiovanni:2021wsd, LIGOScientific:2021aug, Ezquiaga:2022zkx, Karathanasis:2022rtr}.
 
-The fundamental idea of the spectral-siren method is to infer cosmology by comparing the observed mass population distribution with the intrinsic mass population distribution since the observed distribution is redshifted due to the expansion of the Universe.
+The fundamental idea of the spectral-siren method is to infer cosmology by comparing the observed mass population distribution with the intrinsic mass population distribution.
+The observed distribution is redshifted due to the expansion of the Universe.
+Cosmological parameters are inferred by relating the intrinsic mass distribution with the observed mass distribution.
+
 The method works best when the intrinsic distribution is well understood.
-However, the features present in the mass distribution are not so well characterized yet, and their astrophysical origin is a debated topic \citep{Zevin:2017evb, Mapelli:2020vfa, Zevin:2020gbd, Mandel:2018hfr, Marchant:2023wno}.
+However, the features present in the currently inferred mass distribution are not so well characterized yet, and their astrophysical origin is a debated topic \citep{Zevin:2017evb, Mapelli:2020vfa, Zevin:2020gbd, Mandel:2018hfr, Marchant:2023wno}.
 In this case, the spectral-siren method depends on the choice of the population model, which can introduce biases in the inference \citep{Mastrogiovanni:2021wsd, Mukherjee:2021rtw, Pierra:2023deu}.
-Furthermore, the intrinsic distribution is degenerate with cosmology: for example, an intrinsic distribution with a linear redshift evolution is degenerate with $H_0$, as they both redshift the observed distribution in the same way.
-While this example is not physical, this degeneracy still affects the inference of $H_0$.
-Due to these issues, astrophysics and cosmology are tightly linked to each other, and a proper inference of the astrophysical and cosmological parameters should be done simultaneously, which makes the method computationally expensive.
+Furthermore, the intrinsic mass distribution is degenerate with cosmology.
+For example, an intrinsic mass distribution with a linear redshift evolution is degenerate with $H_0$ at low redshift, as they can both redshift the observed distribution in the same way.
+More complex redshift evolution can be degenerate with $H_0$ at high redshift.
+While there is no reason to think a realistic intrinsic mass distribution will evolve linearly with redshift, this degeneracy still affects the inference of $H_0$.
+Due to such degeneracies, astrophysics and cosmology are tightly linked to each other, and proper inference of the astrophysical and cosmological parameters should be done simultaneously, which makes the method computationally expensive.
 
-A possible alternative approach, which we will propose in this paper, is to use non-parametric methods to reconstruct the observed mass population distribution directly from the data.
+A possible alternative approach, which we propose in this paper, is to use non-parametric methods to reconstruct the observed mass population distribution from the data.
 Non-parametric methods allow us to reconstruct arbitrary probability densities without making assumptions about the form of the distribution \citep{Rinaldi:2021bhm}.
 In our context, we can use non-parametric methods to reconstruct the observed mass population distribution directly from the data.
-We can then compare the reconstructed distribution with the predicted distribution from some intrinsic population model and cosmological model, which allows us to infer the model parameters by minimizing the distance between the two distributions.
+We compare the reconstructed distribution with the predicted distribution from some intrinsic population model and cosmological model, which allows us to infer the model parameters by minimizing the distance between the two distributions.
 
-This approach allows us to extract the information contained in the observed data without making assumptions about the population model, cosmological model, or selection function; at the same time, the reconstructed distribution can act as an intermediate observation-driven result that encodes both the intrinsic population information and the cosmological information.
+This approach allows us to extract the information contained in the observed data without making assumptions about the population model, cosmological model, or selection function.
+At the same time, the reconstructed distribution can act as an intermediate observation-driven result that encodes both the intrinsic population information and the cosmological information.
 Making use of the reconstructed distribution, we can easily explore the effect of using different models, as the reconstructed distribution can be reused for different models.
@@
 
-Other studies have made use of non-parametric methods for population studies \citep[e.g.,][]{Mandel:2016prl, Mandel:2018mve, Tiwari:2020vym, Rinaldi:2021bhm, Edelman:2021zkw, Sadiq:2021fin, Edelman:2022ydv, Callister:2023tgi, Ray:2023upk}, and now these methods are starting to be applied to the spectral siren method.
-For example, \cite{Farah:2024xub} and \cite{MaganaHernandez:2024uty} used different non-parametric models to model the intrinsic mass distribution, which allows for a flexible model that can capture the features of the intrinsic distribution.
+Other studies have made use of non-parametric methods for population studies \citep[e.g.,][]{Mandel:2016prl, Mandel:2018mve, Tiwari:2020vym, Rinaldi:2021bhm, Edelman:2021zkw, Sadiq:2021fin, Edelman:2022ydv, Callister:2023tgi, Ray:2023upk, Li:2023yyt, Farah:2024xub, MaganaHernandez:2024uty}, and now these methods are starting to be applied to the spectral siren method.
+For example, \cite{Farah:2024xub, MaganaHernandez:2024uty, Li:2024rmi} independently used different non-parametric models to capture the features of the intrinsic mass distribution flexibly.
 
-In Sec.~\ref{sec:method}, we describe our modified method in detail.
+In Sec.~\ref{sec:method}, we describe our method in detail.
 In Sec.~\ref{sec:mock_data}, we present the analysis setup and results of the mock data study.
@@
 In this section, we present the framework we developed to infer both the intrinsic population model and the cosmological model simultaneously by leveraging the detector-frame mass population distribution.
-In the remainder of this Section, we will make use of the notation summarized in Table~\ref{tab:notation}, taken from \cite{Rinaldi:2021bhm, Rinaldi:2022kyg}.
+We will make use of the notation summarized in Table~\ref{tab:notation}, taken from~\cite{Rinaldi:2021bhm, Rinaldi:2022kyg}.
 
@@
 Specifically, we propose two major modifications to the standard spectral-siren method.
-First, instead of transforming the individual-event posterior probability distribution samples from the detector frame to the source frame during the \ac{PE} process, we transform the population model to the detector frame before performing \ac{PE}.
-Second, instead of directly computing the likelihood of the observed data given the population model, we reconstruct the observed population distribution from the data using a non-parametric method and compare it with the population model.
+First, instead of transforming the individual-event posterior probability distribution samples from the detector frame to the source frame during the \ac{PE} process, we transform the intrinsic population model to the detector frame before performing \ac{PE}.
+Second, instead of directly computing the likelihood of the observed data given the intrinsic population model, we reconstruct the non-parametric observed population distribution from the data and compare it with the intrinsic population model.
 
-With these modifications, we separate the spectral-siren method into two distinct parts.
+With these modifications, we separate the spectral-siren method into two parts.
 The first part is the reconstruction of the observed population from the data, which does not require any assumptions about the population model, and the intermediate result is completely data-driven.
@@
 We choose to consider the primary mass $m_1$ of the \ac{BBH} sources, as we want to study the cosmological parameters with the spectral-siren method.
-In principle, we can consider more parameters at once by reconstructing not only a single-parameter observed distribution but instead a multi-parameter observed distribution.
-However, that will require a more sophisticated population model in the later steps, which is beyond the scope of this work.
+In principle, one could consider more parameters at once by reconstructing not only a single-parameter observed distribution but instead a multi-parameter observed distribution.
+However, that would require a more sophisticated population model in the later steps, which is beyond the scope of this work.
+Thus, we focus on the primary mass, which is one of the better astrophysically modelled quantities.
 
@@
 \end{equation}
-where $\mathrm{det}$ represents whether the event is detectable or not.
+where "$\mathrm{det}$" represents whether the event is detectable or not.
 We assume the $m_1$ distribution does not evolve with $z$.
@@
 
-We opted for reconstructing the observed distribution with the non-parametric method and filtering the observed distribution rather than following the more standard procedure of agnostically reconstructing the intrinsic distribution and comparing it to our astrophysical model to circumvent the limits of non-parametric methods concerning extrapolation.
-If the selection function effectively censors one area of the parameter space, such as the low-mass, high-redshift binaries, a data-driven only method will not be able to characterize the distribution in that specific area, leading to diverging uncertainties in the reconstructed distribution and thus making the whole inference scheme pointless.
-Conversely, if we model the observed distribution and filter the intrinsic distribution model, we are still not able to constrain the population behavior in the censored region but the inference will yield some results.
-This way of including selection effects in the analysis, albeit simple and straightforward, is only an approximation of the real selection process, as pointed out in \cite{Essick:2023upv}: nonetheless, the large uncertainties currently associated with \ac{GW} measurements are most likely to be dominant with respect to the bias induced by this approximation.
-Therefore, we opted to include the selection function in the transformation itself, deferring the investigation of properly including selection effects in the reconstruction of the observed distribution in a future paper.
+We opt for reconstructing the observed distribution with the non-parametric method and constructing the intrinsic, detectable, detector-frame distribution $p(m^z_1|\Lambda, \Omega, \mathrm{det})$.
+This is in contrast to the more standard procedure of agnostically reconstructing the intrinsic distribution and comparing it to our astrophysical model.
+We do so to circumvent the limits of non-parametric methods concerning extrapolation required to account for selection effects.
+If the selection function censors one area of the parameter space, such as the low-mass, high-redshift binaries, a data-driven method will not be able to characterize the distribution in that specific area, leading to diverging uncertainties in the reconstructed distribution.
+Conversely, if we modelled the observed distribution and constructed the intrinsic distribution model non-parametrically, we would not be able to constrain the population behavior in the censored region.
+This method of including selection effects in the analysis is an approximation of the real selection process, as pointed out in \cite{Essick:2023upv}: nonetheless, the large uncertainties currently associated with \ac{GW} measurements are most likely to be dominant with respect to the bias induced by this approximation.
+Therefore, we include the selection function in the transformation, deferring the investigation of properly including selection effects in the reconstruction of the observed distribution in a future paper.
 
-Qualitatively, the transformation described in Eq.~\eqref{eq:transformation} redshifts the source-frame intrinsic population distribution to the detector frame, where the redshift is determined by the cosmological model.
-The selection function is then applied to the redshifted distribution to obtain the detector-frame observed population distribution.
+Qualitatively, the transformation of the intrinsic, detectable, detector-frame primary mass $p(m^z_1|\Lambda, \Omega, \mathrm{det})$ (Eq.~\eqref{eq:transformation}) redshifts the source-frame intrinsic population distribution to the detector frame, where the redshift is determined by the cosmological model.
+The selection function is then applied to the redshifted distribution to obtain the intrinsic, detectable, detector-frame primary mass $p(m^z_1|\Lambda, \Omega, \mathrm{det})$.
 With different $\Lambda$ or $\Omega$, the resulting $p(m^z_1|\Lambda, \Omega, \mathrm{det})$ will be different.
-In Fig.~\ref{fig:simulation_transformation}, we show an example of the transformation.
+In Fig.~\ref{fig:simulation_transformation}, we show three different $p(m^z_1|\Lambda, \Omega, \mathrm{det})$ for three different assumed cosmologies.
 
-In contrast to the first part of the method, this part includes the assumptions about the population and cosmological model, as well as the selection function.
+In contrast to the first part of the method, which was entirely data-driven, the second part involving this transformation includes the assumptions about the population and cosmological model, as well as the selection function.
 After obtaining $p(m^z_1|\Lambda, \Omega, \mathrm{det})$, we can perform the remapping to obtain the optimized $\Omega$ and $\Lambda$.
@@
 For the minimization, we use the \textsc{scipy} \citep{2020SciPy-NMeth} implementation of the modified Powell algorithm.
-Note that we calculate $d_\mathrm{JS}$ in the range from the minimum to the maximum of the median of the $m^z_1$ posterior probability distribution samples of each \ac{GW} event.
-This is because the reconstruction is only valid in the region where observations are present as mentioned in Sec.~\ref{sec:reconstruction}.
+Note that we calculate $d_\mathrm{JS}$ over the range of $m^z_1$ values covered by the posterior probability distribution samples of all \ac{GW} events.
+This is because the reconstruction is only valid in the region where observations are present, as mentioned in Sec.~\ref{sec:reconstruction}.
 In Fig.~\ref{fig:simulation_comparison}, we show an illustration of the comparison between the two populations.
@@
 In this section, we test our method with a mock data study.
-To prepare the mock data, we first generated \variable{output/n_true_samples.txt} sets of $m_1$ and $z$ samples from a \textsc{power law + peak} distribution (see, e.g., Eq.~B4 in \cite{KAGRA:2021duu}) for the $m_1$ distribution and uniform distribution over comoving volume and source-frame time (see, e.g., Eq.~8 in \cite{KAGRA:2021duu}) for the $z$ distribution.
-Here we assumed the $m_1$ distribution does not evolve with $z$.
-We then transformed the $m_1$ and $z$ samples to $m^z_1$ and $d_L$ samples.
-Here we chose the $\Lambda$CDM model with parameters consistent with the results in \cite{Planck:2018vyg} as the cosmological model for the transformation of the samples.
+To prepare the mock data, we first generate \variable{output/n_true_samples.txt} sets of $m_1$ and $z$ samples from a \textsc{power law + peak} distribution (see, e.g., Eq.~B4 in \cite{KAGRA:2021duu}) for the $m_1$ distribution and uniform distribution over comoving volume and source-frame time (see, e.g., Eq.~8 in \cite{KAGRA:2021duu}) for the $z$ distribution.
+Here we assume the $m_1$ distribution does not evolve with $z$.
+We then transform the $m_1$ and $z$ samples to $m^z_1$ and $d_L$ samples.
+Here we choose the $\Lambda$CDM model with parameters consistent with the results in \cite{Planck:2018vyg} as the cosmological model for the transformation of the samples.
 
-After the transformation, we applied a selection function to the $m^z_1$ and $d_L$ samples.
-We used the physically modelled selection function implemented in \cite{Lorenzo-Medina:2024opt}.
-Note that we marginalized over $q$ for the selection function, similar to the transformation in Sec.~\ref{sec:transformation}.
+After the transformation, we apply a selection function to the $m^z_1$ and $d_L$ samples.
+We use the physically modelled selection function implemented in \cite{Lorenzo-Medina:2024opt}.
+We marginalize over $q$ in the selection function, to be consistent with the approach described in Sec.~\ref{sec:transformation}.
 After applying the selection function, which filters out the non-detectable events based on their $m^z_1$ and $d_L$, we retained \variable{output/n_obs_samples.txt} sets of samples for further analysis.
 
-We then generated the $m^z_1$ posterior probability distribution samples for each true $m^z_1$ sample by
+We then generate the $m^z_1$ posterior probability distribution samples for each true $m^z_1$ sample by
 \begin{gather*}
@@
 
-We first reconstructed the observed detector-frame mass population distribution with \textsc{figaro} as described in Sec.~\ref{sec:reconstruction}.
+We first reconstruct the observed detector-frame mass population distribution with \textsc{figaro} (see Sec.~\ref{sec:reconstruction}).
 The result of the reconstruction is shown in Fig.~\ref{fig:simulation_reconstruction}.
@@
     \end{aligned}
-    \label{eq:transformation_H0}
 \end{equation}
@@
         Top: source-frame $m_1$ population distribution.
-        Bottom: detector-frame $m^z_1$ observed distribution for different $H_0$ values. The low-mass peak is suppressed due to the selection function.
+        Bottom: detector-frame $m^z_1$ observed distribution for different $H_0$ values.
+        The low-mass peak is suppressed due to the selection function.
     }
@@
 
-We then applied our optimization scheme to each reconstructed distribution to infer the optimized $H_0$.
-In this simplified case, since the parameter space is one-dimensional, instead of using the optimization algorithm, we relied on a grid-based search to find the optimized $H_0$ between $5$ and $150$ $\mathrm{km}\,\mathrm{s}^{-1}\,\mathrm{Mpc}^{-1}$.
+We then apply our optimization scheme to each reconstructed distribution to infer the optimized $H_0$.
+In this case, since the parameter space is one-dimensional, instead of using the optimization algorithm, we rely on a grid-based search to find the optimized $H_0$ between $5$ and $150$ $\mathrm{km}\,\mathrm{s}^{-1}\,\mathrm{Mpc}^{-1}$.
 For each $H_0$, we computed the $d_\mathrm{JS}$ between the two populations and found the optimized $H_0$ that minimizes $d_\mathrm{JS}$.
@@
 Similar to the previous subsection, we assume perfect knowledge of both models, the selection function and all the parameters that are not objects of the inference.
-Other than $H_0$, we included the power law index $\alpha$, the peak mass $\mu$, and the peak width $\sigma$ of the \textsc{power law + peak} model as the free parameters in the analysis.
+Other than $H_0$, we include the power-law index $\alpha$, the peak mass $\mu$, and the peak width $\sigma$ of the \textsc{power law + peak} model as the free parameters in the analysis.
 The minimum mass $m_\mathrm{min}$ and maximum mass $m_\mathrm{max}$, and the range of mass tapering at the low mass end $\delta$ are fixed.
@@
 In principle, more parameters can be inferred with our method, for example, the peak weight $w$ in the mass population model and the present mass density $\Omega_m$ in the $\Lambda$CDM model.
-However, the optimization procedure becomes more computationally expensive as the number of parameters increases, therefore, we only chose a subset of the parameters to demonstrate the idea of our method.
+However, the optimization procedure becomes more computationally expensive as the number of parameters increases, therefore, we choose a subset of the parameters to demonstrate the idea of our method.
+By performing the population inference together with the cosmological inference, we account for parameter degeneracies between the cosmological and population models.
 
@@
 As long as the events included in the analysis do not change, the reconstruction can be reused.
-With more and more events in the near future, the standard spectral-siren method may not be feasible to apply, and this feature of our method can be advantageous.
+With more and more events in the near future, the standard spectral-siren method may become computationally expensive, and this feature of our method can become advantageous.
 
@@
         \midrule
-        $H_0$ & $[10, 300]$ \\
-        $\alpha$ & $[1.01, 10]$ \\
-        $\mu$ & $[10, 70]$ \\
-        $\sigma$ & $[0.01, 10]$ \\
+        $H_0$ & $[1, 350]$ \\
+        $\alpha$ & $[1.01, 15]$ \\
+        $\mu$ & $[0.01, 70]$ \\
+        $\sigma$ & $[0.01, 60]$ \\
         \botrule
@@
     \caption{
-        Optimized parameter distribution for the mock data inference. Simulated values are marked in red.
+        Optimized parameter distribution for the mock data inference.
+        Simulated values are marked in red.
     }
@@
 
-This is an expected result, as the mock data is generated with the same population model and cosmological model we used in the analysis.
+The consistency between the simulated and inferred parameters is expected, as the mock data is generated with the same population model and cosmological model we used in the analysis.
 We can also see that the uncertainty in $H_0$ is larger compared to the analysis in Sec.~\ref{sec:inference_H0}, as more parameters are inferred simultaneously.
@@
 In this section, we apply our method to the \ac{GW} events released by the \ac{LVK} as part of the GWTC-2.1 \cite{LIGOScientific:2021usb, ligo_scientific_collaboration_and_virgo_2022_6513631} and GWTC-3 \cite{KAGRA:2021vkt, ligo_scientific_collaboration_and_virgo_2023_8177023} catalogs.
-We included all $70$ events labeled as BBH in Table~I of \cite{KAGRA:2021duu}, these correspond to all the events with $\mathrm{FAR_{min}} < 1\, \mathrm{yr}^{-1}$, where $\mathrm{FAR_{min}}$ is the smallest \ac{FAR} among all search pipelines.
+We include all $70$ events labeled as BBH in Table~I of \cite{KAGRA:2021duu}, these correspond to all the events with $\mathrm{FAR_{min}} < 1\, \mathrm{yr}^{-1}$, where $\mathrm{FAR_{min}}$ is the smallest \ac{FAR} among all search pipelines.
 These events are different from the events used in \cite{LIGOScientific:2021aug}, where the authors use a \ac{FAR} threshold of $\mathrm{FAR_{min}} < 0.25\, \mathrm{yr}^{-1}$, therefore the results we present here might be slightly different from the ones in \cite{LIGOScientific:2021aug}.
@@
 
-Similar to Sec.~\ref{sec:mock_data}, we first reconstructed the observed $m^z_1$ population distribution.
+Similar to Sec.~\ref{sec:mock_data}, we first reconstruct the observed $m^z_1$ population distribution.
 We use the posterior probability distribution samples of the \ac{GW} events from \cite{LIGOScientific:2019lzm, KAGRA:2023pio}.
@@
 For the cosmological model, we fixed the parameters to the results from \cite{Planck:2018vyg} except $H_0$.
-For the population model, we fixed $\sigma$, $w$, $m_\mathrm{min}$, $m_\mathrm{max}$, and $\delta$ to the results from \cite{LIGOScientific:2021aug}.
+For the population model, we fixed $\alpha$, $w$, $m_\mathrm{min}$, $m_\mathrm{max}$, and $\delta$ to the results from \cite{LIGOScientific:2021aug}.
 Note that the fixed parameters are not necessarily the true values, but they can be considered as a reasonable guess for the parameters.
@@
          & & $w$ & $0.024$ \\
-        $\Omega_m$ & $0.315$ & $\sigma$ & $3.8\ \mathrm{M}_\odot$\\
+        $\Omega_m$ & $0.315$ & $\alpha$ & $4.2\ \mathrm{M}_\odot$\\
         $\Omega_\Lambda$ & $0.685$ & $m_\mathrm{min}$ & $5\ \mathrm{M}_\odot$ \\
@@
 We then perform the remaining steps similar to Sec.~\ref{sec:inference_multi}.
-We set the bounds for $\kappa$ to $[-10, 10]$, and the bounds for other free parameters are the same as in Table~\ref{tab:bounds}.
+We set the bounds for the power-law index of the redshift population model $\kappa$ to $[-100, 100]$, and the bounds for other free parameters are the same as in Table~\ref{tab:bounds}.
 The result of the remapping is shown in Fig.~\ref{fig:real_result_multi}, and it is consistent with the posterior distribution reported in \cite{KAGRA:2021duu}.
@@
 The result implies that \variable{output/real_H0.txt}, with symmetric $90\%$-credible intervals around the median.
-The uncertainty is much smaller than the one reported in \cite{KAGRA:2021duu} ($H_0 = 50^{+37}_{-30} \, \mathrm{km/s/Mpc}$), mainly because it only includes the uncertainty from the reconstruction, which underestimates the true uncertainty which also includes the mismatch between the population model and the observed population distribution.
+The uncertainty is smaller than the one reported in \cite{KAGRA:2021duu} ($H_0 = 50^{+37}_{-30} \, \mathrm{km/s/Mpc}$), mainly because it only includes the uncertainty from the reconstruction, which underestimates the true uncertainty which also includes the mismatch between the population model and the observed population distribution.
 Figure \ref{fig:real_result_multi} shows a peak in the marginalized posterior distribution of $H_0$ around $70\, \mathrm{km/s/Mpc}$.
@@
 
-This analysis we report here is a proof of concept of the method presented in this paper, and we simply aim to demonstrate that our method can be applied to real data with relatively little effort.
-To perform a detailed analysis, a more comprehensive configuration that considers different population models, cosmological models, and selection functions is required, as well as a more robust optimization scheme to include more astrophysical parameters.
+% This analysis we report here is a proof of concept of the method presented in this paper, and we simply aim to demonstrate that our method can be applied to real data with relatively little effort.
+% To perform a detailed analysis, a more comprehensive configuration that considers different population models, cosmological models, and selection functions is required, as well as a more robust optimization scheme to include more astrophysical parameters.
 

### 2024-12-16 src/scripts/get_real_H0.py
Update get_real_H0.py

@@ -1,7 +1,7 @@
 import numpy as np
 import paths
 
-f = np.load(paths.data / 'real/multi' / '4d_Powell.npz')
+f = np.load(paths.data / 'real/multi' / '4c_Powell.npz')
 result = f['result']
 result = result[(result[:, 0] > 40) & (result[:, 0] < 100)]
 H0 = result[:, 0]

### 2024-12-16 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 Next, we chose the \textsc{power law + peak} model as the source-frame population model $\Lambda$ and the $\Lambda$CDM model as the cosmological model $\Omega$.
-For the redshift population model, instead of the uniform distribution over comoving volume and source-frame time, we used the \textsc{power law} redshift evolution model described in \cite{KAGRA:2021duu}.
+For the redshift population model, instead of the uniform distribution over comoving volume and source-frame time, we used the \textsc{power law} redshift evolution model described in Eq.~8 of \cite{KAGRA:2021duu}.
 Similar to Sec.~\ref{sec:inference_multi}, we fixed some parameters of the population model and the cosmological model.
@@
 We set the bounds for the power-law index of the redshift population model $\kappa$ to $[-100, 100]$, and the bounds for other free parameters are the same as in Table~\ref{tab:bounds}.
-The result of the remapping is shown in Fig.~\ref{fig:real_result_multi}, and it is consistent with the posterior distribution reported in \cite{KAGRA:2021duu}.
+The result of the remapping is shown in Fig.~\ref{fig:real_result_multi}, and it is consistent with the posterior distribution reported in \cite{LIGOScientific:2021aug}.
 \begin{figure}
@@
 The result implies that \variable{output/real_H0.txt}, with symmetric $90\%$-credible intervals around the median.
-The uncertainty is smaller than the one reported in \cite{KAGRA:2021duu} ($H_0 = 50^{+37}_{-30} \, \mathrm{km/s/Mpc}$), mainly because it only includes the uncertainty from the reconstruction, which underestimates the true uncertainty which also includes the mismatch between the population model and the observed population distribution.
-Figure \ref{fig:real_result_multi} shows a peak in the marginalized posterior distribution of $H_0$ around $70\, \mathrm{km/s/Mpc}$.
-This means a significant fraction of the reconstructed observed population distribution has an optimized $H_0$ around $70\, \mathrm{km/s/Mpc}$, which again does not include the uncertainty from the mismatch between the population model and the observed population distribution.
-The peak should be interpreted as an artifact of the optimization procedure.
-
-% This analysis we report here is a proof of concept of the method presented in this paper, and we simply aim to demonstrate that our method can be applied to real data with relatively little effort.
-% To perform a detailed analysis, a more comprehensive configuration that considers different population models, cosmological models, and selection functions is required, as well as a more robust optimization scheme to include more astrophysical parameters.
+The uncertainty is slightly smaller than the one reported in \cite{LIGOScientific:2021aug} ($H_0 = 50^{+37}_{-30} \, \mathrm{km/s/Mpc}$), mainly because it only includes the uncertainty from the reconstruction, which underestimates the true uncertainty which also includes the mismatch between the population model and the observed population distribution.
+Figure \ref{fig:real_result_multi} shows a peak in the marginalized posterior distribution of $H_0$ around $140\, \mathrm{km/s/Mpc}$.
+This means a significant fraction of the reconstructed observed population distribution has an optimized $H_0$ around $140\, \mathrm{km/s/Mpc}$, which again does not include the uncertainty from the mismatch between the population model and the observed population distribution.
 
@@
 With the numbers of observed \ac{GW} events that are expected to grow larger and larger, potentially hitting the thousand mark by the end of the fifth observing run, performing a complete analysis from scratch with different astrophysical models will soon become too computationally expensive.
-The method we propose in this paper can help mitigate this issue by providing a fast and efficient way to explore different models before performing the full hierarchical Bayesian inference, especially keeping in mind that, if a wrong population model is chosen, the inference result will be biased \citep{Mukherjee:2021rtw, Mastrogiovanni:2021wsd, Pierra:2023deu, LIGOScientific:2020kqk, KAGRA:2021duu}.
+The method we propose in this paper can help mitigate this issue by providing a fast and efficient way to explore different models before performing the full hierarchical Bayesian inference, especially keeping in mind that, if a wrong population model is chosen, the inference result will be biased \citep{Mukherjee:2021rtw, Mastrogiovanni:2021wsd, Pierra:2023deu, LIGOScientific:2020kqk, KAGRA:2021duu, LIGOScientific:2021aug}.
 This makes our method a powerful diagnostic tool for model selection.

### 2024-12-16 src/scripts/get_real_H0.py
Update plots

@@ -3,7 +3,6 @@ import paths
 
 f = np.load(paths.data / 'real/multi' / '4c_Powell.npz')
 result = f['result']
-result = result[(result[:, 0] > 40) & (result[:, 0] < 100)]
 H0 = result[:, 0]
 med = np.median(H0)
 low = np.percentile(H0, 10)

### 2024-12-16 src/scripts/plot_real_result_multi.py

@@ -8,5 +8,5 @@ result = f['result']
 result = result[(result[:, 3] > -5) & (result[:, 3] < 15)] # plot limits
 parameters = ["$H_0$", "$\\mu$", "$\\sigma$", "$\\kappa$"]
 
-fig = corner(result, labels=parameters, color='steelblue', truth_color='red', levels = [0.3935, 0.90], plot_datapoints=False, plot_density=False, smooth=1, fill_contours=True, contourf_kwargs={'colors': ['white', 'darkturquoise', 'mediumturquoise'], 'alpha': [1, 0.2, 0.5]}, contour_kwargs={'linewidths': 1})
+fig = corner(result, labels=parameters, color='steelblue', truth_color='red', levels = [0.3935, 0.90], plot_datapoints=False, plot_density=False, smooth=1, fill_contours=True, contourf_kwargs={'colors': ['white', 'darkturquoise', 'mediumturquoise'], 'alpha': [1, 0.2, 0.5]}, contour_kwargs={'linewidths': 1}, show_titles=True)
 fig.savefig(paths.figures / "real_result_multi.pdf", bbox_inches='tight')

### 2024-12-16 src/scripts/plot_simulation_result_multi.py

@@ -9,5 +9,5 @@ result = f['result']
 simulated_truth = {'$H_0$': Planck18.h*100, '$\\alpha$': 3.5, '$\\mu$': 35, '$\\sigma$': 5, '$\\delta$': 5, '$w$': 0.2, '$m_\mathrm{min}$': 5, '$m_\mathrm{max}$': 90}
 parameters = ["$H_0$", "$\\alpha$", "$\\mu$", "$\\sigma$"]
 
-fig = corner(result, labels=parameters, truths=[simulated_truth[param] for param in parameters], color='steelblue', truth_color='red', levels = [0.3935, 0.95], plot_datapoints=False, plot_density=False, smooth=1, fill_contours=True, contourf_kwargs={'colors': ['white', 'darkturquoise', 'mediumturquoise'], 'alpha': [1, 0.2, 0.5]}, contour_kwargs={'linewidths': 1})
+fig = corner(result, labels=parameters, truths=[simulated_truth[param] for param in parameters], color='steelblue', truth_color='red', levels = [0.3935, 0.95], plot_datapoints=False, plot_density=False, smooth=1, fill_contours=True, contourf_kwargs={'colors': ['white', 'darkturquoise', 'mediumturquoise'], 'alpha': [1, 0.2, 0.5]}, contour_kwargs={'linewidths': 1}, show_titles=True)
 fig.savefig(paths.figures / "simulation_result_multi.pdf", bbox_inches='tight')

### 2024-12-16 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 However, the features present in the currently inferred mass distribution are not so well characterized yet, and their astrophysical origin is a debated topic \citep{Zevin:2017evb, Mapelli:2020vfa, Zevin:2020gbd, Mandel:2018hfr, Marchant:2023wno}.
-In this case, the spectral-siren method depends on the choice of the population model, which can introduce biases in the inference \citep{Mastrogiovanni:2021wsd, Mukherjee:2021rtw, Pierra:2023deu}.
+In this case, the spectral-siren method depends on the choice of the population model, which can introduce biases in the inference \citep{Mukherjee:2021rtw, Mastrogiovanni:2021wsd, Pierra:2023deu, LIGOScientific:2020kqk, KAGRA:2021duu, LIGOScientific:2021aug}.
 Furthermore, the intrinsic mass distribution is degenerate with cosmology.
@@
 We can also see that the uncertainty in $H_0$ is larger compared to the analysis in Sec.~\ref{sec:inference_H0}, as more parameters are inferred simultaneously.
-The uncertainty in $\sigma$ is relatively large, this may be due to the large uncertainty in the reconstruction at the gap between the low-mass and high-mass peaks, which can be seen in 
-Fig.~\ref{fig:simulation_reconstruction}.
 
@@
 The result implies that \variable{output/real_H0.txt}, with symmetric $90\%$-credible intervals around the median.
-The uncertainty is slightly smaller than the one reported in \cite{LIGOScientific:2021aug} ($H_0 = 50^{+37}_{-30} \, \mathrm{km/s/Mpc}$), mainly because it only includes the uncertainty from the reconstruction, which underestimates the true uncertainty which also includes the mismatch between the population model and the observed population distribution.
+The uncertainty is larger than the one reported in \cite{LIGOScientific:2021aug} ($H_0 = 50^{+37}_{-30} \, \mathrm{km/s/Mpc}$), .
 Figure \ref{fig:real_result_multi} shows a peak in the marginalized posterior distribution of $H_0$ around $140\, \mathrm{km/s/Mpc}$.
-This means a significant fraction of the reconstructed observed population distribution has an optimized $H_0$ around $140\, \mathrm{km/s/Mpc}$, which again does not include the uncertainty from the mismatch between the population model and the observed population distribution.
+This means a significant fraction of the reconstructed observed population distribution has an optimized $H_0$ around $140\, \mathrm{km/s/Mpc}$, which does not include the uncertainty from the mismatch between the population model and the observed population distribution.
+The reason for the peak differing from the one reported in \cite{LIGOScientific:2021aug}, which is around $50\, \mathrm{km/s/Mpc}$, .
 

### 2024-12-18 src/scripts/get_real_H0.py
Use 1 sigma instead of 90%

@@ -5,8 +5,8 @@ f = np.load(paths.data / 'real/multi' / '4c_Powell.npz')
 result = f['result']
 H0 = result[:, 0]
 med = np.median(H0)
-low = np.percentile(H0, 10)
-high = np.percentile(H0, 90)
+low = np.percentile(H0, 50-68/2)
+high = np.percentile(H0, 50+68/2)
 
 with open(paths.output/'real_H0.txt', 'w') as f:
     f.write(f'$H_0 = {med:.2f}^{{+{high - med:.2f}}}_{{-{med - low:.2f}}} \, \mathrm{{km/s/Mpc}}$')

### 2024-12-19 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 We include all $70$ events labeled as BBH in Table~I of \cite{KAGRA:2021duu}, these correspond to all the events with $\mathrm{FAR_{min}} < 1\, \mathrm{yr}^{-1}$, where $\mathrm{FAR_{min}}$ is the smallest \ac{FAR} among all search pipelines.
-These events are different from the events used in \cite{LIGOScientific:2021aug}, where the authors use a \ac{FAR} threshold of $\mathrm{FAR_{min}} < 0.25\, \mathrm{yr}^{-1}$, therefore the results we present here might be slightly different from the ones in \cite{LIGOScientific:2021aug}.
+The cut used to select these events is different from the one used in \cite{LIGOScientific:2021aug}, against which we will compare our results: there, the authors use a \ac{FAR} threshold of $\mathrm{FAR_{min}} < 0.25\, \mathrm{yr}^{-1}$, therefore the results we present here will be, in principle, different.
 We opted for a larger \ac{FAR} threshold -- consistent, however, with the one used in population studies such as \cite{KAGRA:2021duu} -- to have a larger number of events to include in the non-parametric analysis.
@@
 
-We then perform the remaining steps similar to Sec.~\ref{sec:inference_multi}.
+We then perform the remaining steps in the same way as Sec.~\ref{sec:inference_multi}.
 We set the bounds for the power-law index of the redshift population model $\kappa$ to $[-100, 100]$, and the bounds for other free parameters are the same as in Table~\ref{tab:bounds}.
@@
 \end{figure}
-The result implies that \variable{output/real_H0.txt}, with symmetric $90\%$-credible intervals around the median.
-The uncertainty is larger than the one reported in \cite{LIGOScientific:2021aug} ($H_0 = 50^{+37}_{-30} \, \mathrm{km/s/Mpc}$), .
-Figure \ref{fig:real_result_multi} shows a peak in the marginalized posterior distribution of $H_0$ around $140\, \mathrm{km/s/Mpc}$.
-This means a significant fraction of the reconstructed observed population distribution has an optimized $H_0$ around $140\, \mathrm{km/s/Mpc}$, which does not include the uncertainty from the mismatch between the population model and the observed population distribution.
-The reason for the peak differing from the one reported in \cite{LIGOScientific:2021aug}, which is around $50\, \mathrm{km/s/Mpc}$, .
+we find \variable{output/real_H0.txt}, (68\% credible interval around the median).
+Nonetheless, Figure~\ref{fig:real_result_multi} shows a peak in the marginalized posterior distribution of $H_0$ around $140\, \mathrm{km/s/Mpc}$, with a width of around $30-35\, \mathrm{km/s/Mpc}$ roughly consistent with the uncertainty reported in \cite{LIGOScientific:2021aug} ($H_0 = 50^{+37}_{-30} \, \mathrm{km/s/Mpc}$).
+%This means a significant fraction of the reconstructed observed population distribution has an optimized $H_0$ around $140\, \mathrm{km/s/Mpc}$, which does not include the uncertainty from the mismatch between the population model and the observed population distribution.
+
+The reason for the peak differing from the one reported in \cite{LIGOScientific:2021aug}, which is around $50\, \mathrm{km/s/Mpc}$, can be several, and most likely stemming from the various approximations made in this work.
+First of all, the choice of the Jensen-Shannon distance as a measure of ``closeness'' between the non-parametric reconstruction and the predicted population is an arbitrary choice and, despite being a commonly used metric, does not stem from a specific derivation and/or assumption on the process that generated the observations.
+As we stated in Section~\ref{sec:remapping}, the output of our method is not a posterior distribution in the strict sense due to this choice: the derivation of a well-grounded metric to connect non-parametric and parametric distributions will be discussed in a future paper.
+Another possibility is the approximation made while including the selection effects: as discussed in \cite{Essick:2023upv}, this choice can lead, to some extent, to a bias in the reconstruction, and thus be the reason why the peak of the posterior distribution for $H_0$ is shifted.
+Accounting for selection effects while reconstructing the observed distribution is in principle possible -- although not explored in the currently available literature -- and will be, as well, investigated in the future.
 

### 2025-01-09 src/tex/ms.tex
[showyourwork] overleaf sync

@@
 \usepackage{orcidlink}
+\usepackage{tabularx}
 
@@
 
-\begin{table}[htbp]
+\begin{table}
     \caption{Notation used in this paper}
@@
 The optimization method requires bounds for the free parameters, which we show in Table~\ref{tab:bounds}.
-\begin{table}[htbp]
+\begin{table}
     \caption{Free parameter bounds for the joint inference presented in Section~\ref{sec:inference_multi}.}
-    \begin{tabular}{cc}
+    \begin{tabularx}{\linewidth}{>{\centering\arraybackslash}X >{\centering\arraybackslash}X}
         \toprule
@@
         \botrule
-    \end{tabular}
+    \end{tabularx}
     \label{tab:bounds}
@@
         Simulated values are marked in red.
+        The shaded regions represent the 39.35\% and 90\% credible intervals.
     }
@@
 The fixed parameters of the population model are shown in Table~\ref{tab:fixed_parameters}.
-\begin{table}[htbp]
+\begin{table}
     \caption{Fixed parameters of the $\Lambda$CDM model (left) and the \textsc{power law + peak} model (right).}
-    \begin{tabular}{cc|cc}
+    \begin{tabularx}{\linewidth}{>{\centering\arraybackslash}X >{\centering\arraybackslash}X | >{\centering\arraybackslash}X >{\centering\arraybackslash}X}
         \toprule
@@
          & & $w$ & $0.024$ \\
-        $\Omega_m$ & $0.315$ & $\alpha$ & $4.2\ \mathrm{M}_\odot$\\
+        $\Omega_m$ & $0.315$ & $\alpha$ & $4.2\ \mathrm{M}_\odot$ \\
         $\Omega_\Lambda$ & $0.685$ & $m_\mathrm{min}$ & $5\ \mathrm{M}_\odot$ \\
         $w$ & $-1$ & $m_\mathrm{max}$ & $109\ \mathrm{M}_\odot$ \\
-         & & $\delta$ & $4.9\ \mathrm{M}_\odot$\\
+         & & $\delta$ & $4.9\ \mathrm{M}_\odot$ \\
         \botrule
-    \end{tabular}
+    \end{tabularx}
     \label{tab:fixed_parameters}
@@
         Optimized parameter distribution for the real data inference.
+        The shaded regions represent the 39.35\% and 90\% credible intervals.
     }
@@
 \end{figure}
-we find \variable{output/real_H0.txt}, (68\% credible interval around the median).
+We find \variable{output/real_H0.txt}, (68\% credible interval around the median).
 Nonetheless, Figure~\ref{fig:real_result_multi} shows a peak in the marginalized posterior distribution of $H_0$ around $140\, \mathrm{km/s/Mpc}$, with a width of around $30-35\, \mathrm{km/s/Mpc}$ roughly consistent with the uncertainty reported in \cite{LIGOScientific:2021aug} ($H_0 = 50^{+37}_{-30} \, \mathrm{km/s/Mpc}$).
-%This means a significant fraction of the reconstructed observed population distribution has an optimized $H_0$ around $140\, \mathrm{km/s/Mpc}$, which does not include the uncertainty from the mismatch between the population model and the observed population distribution.
 
-The reason for the peak differing from the one reported in \cite{LIGOScientific:2021aug}, which is around $50\, \mathrm{km/s/Mpc}$, can be several, and most likely stemming from the various approximations made in this work.
-First of all, the choice of the Jensen-Shannon distance as a measure of ``closeness'' between the non-parametric reconstruction and the predicted population is an arbitrary choice and, despite being a commonly used metric, does not stem from a specific derivation and/or assumption on the process that generated the observations.
+The reason for the peak differing from the one reported in \cite{LIGOScientific:2021aug}, which is around $50\, \mathrm{km/s/Mpc}$, can be attributed to two main factors.
+First, the choice of the Jensen-Shannon distance as a measure of ``closeness'' between the non-parametric reconstruction and the predicted population is an arbitrary choice and, despite being a commonly used metric, does not stem from a specific derivation and/or assumption on the process that generated the observations.
 As we stated in Section~\ref{sec:remapping}, the output of our method is not a posterior distribution in the strict sense due to this choice: the derivation of a well-grounded metric to connect non-parametric and parametric distributions will be discussed in a future paper.
@@
 Accounting for selection effects while reconstructing the observed distribution is in principle possible -- although not explored in the currently available literature -- and will be, as well, investigated in the future.
+Despite these challenges, our method provides a foundation for exploring different models efficiently and can significantly aid in the analysis of large datasets expected in the near future.
 
